### 2022-11-15 use_case/analysis_vasseur_fox.Rmd
Finish with McCann / Brose

---
title: "Vasseur \& Fox (2007) replication"
author: Alain Danet
date: \today
output:
  bookdown::pdf_document2:
    fig_caption: true 
    keep_tex: true
    toc: true
    fontsize: 12pt
    geometry: margin=2cm
    number_sections: true
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = TRUE,
  fig.path="figures/"
)
```

```{r, results = FALSE}
library(tidyverse)
library(magrittr)
library(cowplot)
library(here)
require(rgl)
options(rgl.printRglwidget = TRUE)

mytheme <- theme_cowplot() +
background_grid()

theme_set(mytheme)
options(knitr.table.format = "latex")
```

```{r}
data <- read_csv(here("res", "vasseur_fox_res.csv"))
colnames(data)[1:2] <- c("rho", "sigma") 
```
```{r}
avg  <- data %>%
  group_by(sigma, rho) %>%
  summarise(across(everything(), mean))
```

```{r}
surface3d(avg$rho, avg$stab_pred, avg$sigma)
map(c("rho", "stab_pred", ))
```

```{r}
cons_sync <- avg %>%
  ggplot(aes(rho, sigma, z = sync_cons)) +
  geom_contour_filled() +
  labs(x = "Environmental correlation", y = "Fluctuation strength") +
  labs(fill = "Consumer synchrony")

pred_stab <- ggplot(avg, aes(rho, sigma, z = stab_pred)) +
  geom_contour_filled() +
  labs(x = "Environmental correlation", y = "Fluctuation strength") +
  labs(fill = "Predator stability")

cons1_stab <- avg %>%
  ggplot(aes(rho, sigma, z = stab_cons1)) +
  geom_contour_filled() +
  labs(x = "Environmental correlation", y = "Fluctuation strength") +
  labs(fill = "Consumer 1 stability")

cons2_stab <- avg %>%
  ggplot(aes(rho, sigma, z = stab_cons2)) +
  geom_contour_filled() +
  labs(x = "Environmental correlation", y = "Fluctuation strength") +
  labs(fill = "Consumer 2 stability")
```

```{r}
pp <- plot_grid(
  cons_sync, pred_stab, cons1_stab, cons2_stab,
  labels = "auto")
save_plot(
  filename = here("res", "vasseurfox2007fig2.pdf"),
  pp, ncol = 2, nrow = 2)
```