### 2022-11-18 report/project_definition.Rmd
Clarify differences between McCann and Brose

---
title: "Post-doc project: A complex food-web theory of stability against perturbation"
author: "Alain Danet"
date: \today
output:
  bookdown::pdf_document2
bibliography: references.bib
---

# Rationale

## Perturbation ecology

- Multiple perturbation affect simultaneously mortality rates of species
- We can assume that those multiple stressors can be summed into environmental
  stochasticity

## Correlation among pertubations

- Depending on the nature of the stressors and niche of species, we can expect
  different scenarios
- Perturbations such as fishing will affect primarily some species (e.g.
  top-trophic level)
- Perturbations in habitat quality will affect species more globally
  (however some might benefit).
- Perturbation in temperature will affect differently species according to their thermal niche

All the above scenarios can result in quite complex combinations of situation
that can be tricky to generalize. Another way to tackle this situation is to
test more general theories of ecological communities.

I see four different scenarios of perturbation correlations among species that I
name as the following homogeneous, null, guilds, hidden niches. The homogeneous
scenario considers that all species are affected in the same way, which would
correspond to a general perturbation such as a general improvement/depletion of
habitat quality. The null scenario considers that all species respond randomly
to stress, i.e. they do not share perturbation, i.e. their perturbations are
uncorrelated. The guilds scenario assumes that intra-guild perturbations are
positively correlated and that interguild perturbations are not correlated, it
is a scenario that assumes that species sharing the same trophic level share the
same niche, i.e. there is no intraguild response diversity. The hidden niche
scenario comes the theory of the same name[@scheffer_self-organized_2006;
@barabas_emergent_2013; @martinez-blancas_species_nodate]. It assumes that
cluster of species having similar traits (here food-web trophic position) would
have hidden niche dimensions that stabilize their coexistence, such as species
in the same guild will have response diversity (portfolio effect),
anti-correlated perturbations.

While all the above scenario can seem abstract at a first glance, it is possible
to draw realistic situations where they can apply.

The homogeneous scenario can apply in situations where stressors select for
homogeneous response to stressors as for example, as it might be the case in
altered environments where we can end up with the same functions (i.e. the same
food-web) but where the diversity of other niche dimensions have been shrunken
(lost of diversity hypothesis). The guild scenario supposes a full niche
conservatism where species on the same feeding guild are likely to share the
same evolutionary history and such their respond in the same way to
environmental perturbation. The hidden niche is the reverse of the former,
intraguild species are anti-correlated, such as there is maximal response
diversity (insurance hypothesis), which is the mechanism by which biodiversity
is thought to maintain long-term stability of ecosystem functions.

While response diversity has been well studied in small food-web modules and
competitive communities, it has been overlooked in more complex food-webs. The
complex relationships among species across food-web might dampens (compensate),
propagate or just scale with food-chain and small food-web modules.

## Stability in food-webs

The question at 1bn pounds is "what is driving stability?", i.e. the propagation
of perturbations across food-webs. Early studies focused on the effect of
removing one species either by looking at the topology [e.g.,
@dunne_network_2002] or dynamic [@stouffer_compartmentalization_2011]. In that
sense, it would be interesting to see if environmental stochasticity predicts the same
pattern that species extinction (i.e. pertubation for one species only).

The big other ingredient of food-web stability is the strength of species
interactions, as demonstrated by previous studies [@mccann_weak_1998;
@brose_allometric_2006]. @mccann_weak_1998 showed that food-web having strong
interactions can persist if the strong interactions are coupled with weak ones.
Interestingly, a strong interaction needs to be coupled in two ways, by a
feeding and a competitive interactions. @brose_allometric_2006 showed the
strength of interactions are driven by body size, such larger body sized animals
produce lower interaction strengths (because of their lower metabolic rates).

Interestingly, I see two schools on the definition of interaction strengths in
the Bioenergetic model. McCann defines it by modulating consumer preference
($\omega$) such as $I_{ki} = \omega_{ik}* x_i * y_i / B_{0i}$ while
@brose_allometric_2006 modules interaction strength by modulating predator/prey
mass ratios (Z) such as Z decrease the metabolism of predators
relative to prey (i.e. $x_i$ and $y_i$). Thus increasing Z decreases $I_{ki}$.
A common feature of both approaches is that they both consider that the strength
of interactions decrease as the trophic level increases.

A main difference between Brose and McCann approaches is that the Brose
approach of modulating interactions also modulates metabolic losses ($x_iy_iB_i$).
Then the approach of McCann has the advantage to only change feeding rates but
it is rather phenomenological modelling while Brose approach is more mechanistic
(fully metabolic) but it changes two things at the same time.

Although in complex food-webs, Brose method can create strong asymetry in
interaction strength by coupling specialists and generalists, as they
respectively generate strong and weak interactions. Such mechanism is
triggered by the fact that consumer preference is equally shared among preys
($\omega_ki = 1 \ \sum_k 1 $, such as $\omega_{ki} = 1$ or .1 for a consumer
having one prey or 10 preys).

### Stability in small modules

- @mccann_diversitystability_2000, @mccann_weak_1998,
  @vasseur_environmental_2007, @ripa_food_2003

### Stability in food-chains

- @barbier_pyramids_2019, @shanafelt_stability_2018

### Stability in complex food-web

- @thebault_stability_2010, @brose_allometric_2006

## Stochasticity in food-webs

@vasseur_environmental_2007 found that Stochasticity and response diversity can
stabilize food-web in a McCann model with strong asymetry in species interactions.
Is it valid also with a Brose model without this asymetry? Is it also valid with
different functional responses?

# Research questions

## Response diversity

- How does response diversity drive food-web stability?
- How does response diversity affect coupling between predator and preys?
- How does response diversity effect on stability is modulated by interaction strength
  distribution?

# Methods

## Simulation framework:

- Start analysis with really robust food-webs (high Z and type III functional
  response) to avoid species extinction
- Really general propagation: one species pertubated at the time (are there
  differences with extinction pattern ?)
- Modulating interaction strength distribution: generate random asymetric
  consumer preference ($\omega_ij$) but with constraints on predator
  ($\sum\omega_i = 1$) and preys (prey should be involved in one strong
  interaction at the time may be).

# References {-}