### 2022-12-06 report/mccann_brose.Rmd
Analysis Vasseur & Fox version Brose

---
title: "Brose for Vasseur and Fox"
author: "Alain Danet"
date: \today
output:
  bookdown::pdf_document2:
    fig_caption: true
    keep_tex: true
    toc: true
    fontsize: 12pt
    geometry: margin=2cm
    number_sections: true
bibliography: references.bib
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = TRUE,
  fig.path="figures/"
)
```

```{r, results = FALSE}
library(tidyverse)
library(magrittr)
library(cowplot)
library(here)

mytheme <- theme_cowplot(font_size = 8) +
background_grid() +
theme(legend.position = "bottom")

theme_set(mytheme)
```

```{r}
sim_brut <- read_csv(here("res", "vasseur_fox__brose_res.csv"))

sim <- sim_brut %>%
  rename_with(~str_replace_all(., c("ρ" = "rho", "σₑ" = "sigma"))) %>%
  group_by(Z, fr, rho, sigma) %>%
  summarise(across(everything(), median)) %>%
  select(-rep, -sim_timing)
```

```{r}
sim %>%
  filter(stab_com > 10)
```

```{r}
sim %>%
  ggplot(aes(y = stab_com, x = fr)) +
  geom_boxplot() +
  scale_y_log10()
```

```{r}
cons_sync <- sim %>%
  ggplot(aes(rho, sigma, z = sync_cons)) +
  geom_contour_filled() +
  facet_grid(rows = vars(fr), cols = vars(Z)) +
  labs(x = "Environmental correlation", y = "Fluctuation strength") +
  labs(fill = "Consumer synchrony")

pred_stab <- ggplot(sim, aes(rho, sigma, z = log10(stab_pred))) +
  geom_contour_filled() +
  facet_grid(rows = vars(fr), cols = vars(Z)) +
  labs(x = "Environmental correlation", y = "Fluctuation strength") +
  labs(fill = "Predator stability")

cons1_stab <- sim %>%
  ggplot(aes(rho, sigma, z = log10(stab_cons1))) +
  geom_contour_filled() +
  facet_grid(rows = vars(fr), cols = vars(Z)) +
  labs(x = "Environmental correlation", y = "Fluctuation strength") +
  labs(fill = "Consumer 1 stability")

cons2_stab <- sim %>%
  ggplot(aes(rho, sigma, z = log10(stab_cons2))) +
  geom_contour_filled() +
  facet_grid(rows = vars(fr), cols = vars(Z)) +
  labs(x = "Environmental correlation", y = "Fluctuation strength") +
  labs(fill = "Consumer 2 stability")
```

# Vasseur/Fox vs Brose

```{r}
cons_sync_cap <- "Column facets represent increasing Z, row ones respectively
Predator interference (PI, h = 1, c = 1), type II (h = 1, c = 0) and type III (h
  = 2, c = 0)"
```

```{r cs, fig.cap=cons_sync_cap}
cons_sync
```

We find that the strength of environmental fluctuations desynchronises consumer
fluctuations for very low strength of environmental fluctuations, and
has mostly no effects for higher strength of environmental fluctuations (Fig.
\@ref(fig:cs)). This
result constrasts with previous results that showed that the strength of
environmental fluctuations synchronises consumer fluctuations
@vasseur_environmental_2007. Furthermore, we find that it is environmental
correlation that drives consumer synchrony, negative correlations promoting
synchrony while positive correlations promote synchrony. Again it contrasts with
@vasseur_environmental_2007, which found that environmental correlations had
little effects on consumer synchrony.

```{r ps, fig.cap="Stability as 1/CV. See fig. 1 for facets"}
pred_stab
```

In turn, we find that predator stability decreases with increasing environmental
fluctuations , but is little affected by environmental
synchrony of consumer (Fig. \@ref(fig:ps)). It contrasts with
@vasseur_environmental_2007 which found an island of high predator stability for
intermediation strength of environmental fluctuation and high positive
correlation of environmental fluctuations between consumers. Overall, the
stability are really high, between 1 and $2.10^5$ (median: 90). It is surely
because both the metabolism and the interaction strenghs are too weak, so
fluctuations are really tiny and fluctuations does not propagate much . I should
try lower Z.

```{r}
stab <- ggplot(sim, aes(rho, sigma, z = log10(stab_com))) +
  geom_contour_filled() +
  facet_grid(rows = vars(fr), cols = vars(Z)) +
  labs(x = "Environmental correlation", y = "Fluctuation strength") +
  labs(fill = "Community stability")

sp_stab <- ggplot(sim, aes(rho, sigma, z = log10(1 / avg_cv_sp))) +
  geom_contour_filled() +
  facet_grid(rows = vars(fr), cols = vars(Z)) +
  labs(x = "Environmental correlation", y = "Fluctuation strength") +
  labs(fill = "Average species stab")

sync <- sim %>%
  ggplot(aes(rho, sigma, z = sync)) +
  geom_contour_filled() +
  facet_grid(rows = vars(fr), cols = vars(Z)) +
  labs(x = "Environmental correlation", y = "Fluctuation strength") +
  labs(fill = "Community synchrony")
```

# References {-}