### 2023-01-14 use_case/module_analysis.qmd

---
title: Module stability
author: Alain Danet
output: html
---

```{r}
library(tidyverse)
library(magrittr)
library(arrow)
```

```{r}
df_brut <- read_csv(here::here("res", "cs_sim.csv"))
df <- df_brut %>%
  select(where(is.double))
```

```{r}
summary(df)

df %>%
  filter(stab_com <= 1000) %>%
  ggplot(aes(x = as.factor(richness), y = stab_com, color = as.factor(ct))) +
  geom_boxplot()
```

```{r}
corrplot::corrplot(
  cor(df %>% select(-rep, -Z, -env_stoch, -sim_timing), method = "spearman"),
  diag = FALSE, type = "upper"
)
```