### 2023-01-31 report/module_analysis.Rmd
Add varying Z for various interaction strengths

---
title: Module analysis
author: alain
date: 2022-09-12
output:
  bookdown::pdf_document2
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = TRUE,
  fig.path="figures/"
)
```

```{r, results = FALSE}
library(tidyverse)
library(magrittr)
library(cowplot)
library(here)
library(arrow)

mytheme <- theme_cowplot(font_size = 8) +
background_grid() +
theme(legend.position = "bottom")

theme_set(mytheme)
```

# Stochastic natural death

## Competition

```{r}
comp_brut <- open_dataset(here::here("res", "sim_module_competition.arrow"), format = "arrow")
comp <- comp_brut %>%
  collect() %>%
  select(where(is.double))
colnames(comp)
```

```{r, eval = FALSE}
unique(comp$env_stoch)
summary(comp$total_biomass)
```

```{r}
comp %>%
  filter(env_stoch == 0.9) %>%
  pivot_longer(c(stab_com, sync, avg_cv_sp), names_to = "stab", values_to =
  "values") %>%
  ggplot(aes(x = as.factor(richness), y = values, color = as.factor(alpha_ij))) +
  geom_boxplot() +
  facet_grid(
    rows = vars(stab), cols = vars(rho),
    scales = "free_y") +
  labs(x = "Species richness", y = "Stability component", color = "Interspecific competition")
```

- Main results:
  - Species richness increases stability
  - Interspecific competition does not affect stability expect when
    interspecific competition equals to intraspecific competition (decrease the effect of species richness).
  - Interspecific competition desynchronises primary producers but increases
    varibility
  - The decrease of response diversity decreases the effect of species richness on stability
  - The decrease of response diversity increase synchrony but decrease
    population variability

## Modules

```{r}
mod_brut <- open_dataset(here::here("res", "sim_module_d.arrow"), format = "arrow")
mod <- mod_brut %>%
  collect() %>%
  select(where(is.numeric) | starts_with("module")) %>%
  filter(richness / S == 1.0)
colnames(mod)
```

```{r, eval = FALSE}
unique(mod$env_stoch)
summary(as_tibble(mod))
```

```{r}
mod %>%
  filter(rho == 0, env_stoch == .9) %>%
  pivot_longer(c(stab_com, sync, avg_cv_sp), names_to = "stab", values_to =
  "values") %>%
  ggplot(aes(x = module_name, y = values, color = as.factor(alpha_ij))) +
  geom_boxplot() +
  facet_grid(rows = vars(stab), scales = "free_y")
```

```{r}
mod %>%
  filter(rho == 0.5, env_stoch == .9) %>%
  pivot_longer(c(stab_com, sync, avg_cv_sp), names_to = "stab", values_to =
  "values") %>%
  ggplot(aes(x = module_name, y = values, color = as.factor(alpha_ij))) +
  geom_boxplot() +
  facet_grid(rows = vars(stab), scales = "free_y")

```

```{r}
mod %>%
  filter(rho == 1, env_stoch == .9) %>%
  pivot_longer(c(stab_com, sync, avg_cv_sp), names_to = "stab", values_to =
  "values") %>%
  ggplot(aes(x = module_name, y = values, color = as.factor(alpha_ij))) +
  geom_boxplot() +
  facet_grid(rows = vars(stab), scales = "free_y")
```