### 2023-05-13 report/cs_analysis.Rmd
Update report

---
title: Simulation connectance richness
author: Alain Danet
output:
  bookdown::pdf_document2
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = TRUE,
  fig.path="figures/"
)
```

```{r, results = FALSE}
library(tidyverse)
library(magrittr)
library(cowplot)
library(here)
library(arrow)
library(piecewiseSEM)
library(semEff)
library(easystats)
library(glmmTMB)

mytheme <- theme_cowplot(font_size = 8) +
background_grid() +
theme(legend.position = "bottom")
theme_set(mytheme)
source(here::here("R", "network_metrics.R"))
source(here::here("R", "string_replacements.R"))
```

```{r}
sim_files <- list.files(here("res", "simCSZenrich"))
sim_list <- map(sim_files, function (x) {
  open_dataset(here::here("res", "simCSZenrich", x),
    format = "arrow") %>%
  collect()
})
```

```{r}
# TODO set extinction threshold in simulations
simZ_prep <- map_dfr(sim_list,
  function (x) {
    x %>%
      # Remove simulation that have failed to run
      filter(!is.na(richness)) %>%
      # Transform back interaction strength into matrices
      mutate(across(where(is.list), as.list)) %>%
      mutate(across(c(int_strength, max_int),
          ~map2(.x, S,
            function(y, rich) {
              matrix(y, nrow = rich)
            }
          )
      )
        ) %>%
      # Get matrix of alive species
      mutate(max_int_alive  = map2(max_int, bm_sp,
          function(mat, bm) {
            mask <- bm > 10^-6
            mat[mask, mask, drop = FALSE]
          }
        )
        )
  }
)
# Compute metrics of interest
simZ <- simZ_prep %>%
  mutate(
    persistence = richness / S,
    async = 1 / sync,
    stab_pop = 1 / avg_cv_sp,
    ct_alive = map_dbl(max_int_alive,
      ~ifelse(sum(.x) == 0, 0,
      sum(.x > 0) / ((ncol(.x))^2))),
    bm_sp_alive = map(bm_sp, ~.x[.x > 10^-6]),
    tlvl_alive = map2(tlvl, bm_sp, ~.x[.y > 10^-6]),
    max_tlvl_alive = map_dbl(tlvl_alive, max),
    w_avg_tlvl_alive = map2_dbl(tlvl_alive, bm_sp_alive, ~sum(.x * (.y / sum(.y)))),
    max_max_int_alive = map_dbl(max_int_alive,
      ~ifelse(length(.x[.x > 0]) == 0, 0, max(.x[.x > 0]))
      ),
    avg_max_int_alive = map_dbl(max_int_alive,
      ~ifelse(length(.x[.x > 0]) == 0, 0, mean(.x[.x > 0]))
      ),
    gini_max_int_alive = map_dbl(max_int_alive, ~gini(.x))
      )
simZ$tlvl_alive[[3]]
simZ$max_int_alive[[3]]
```

```{r, eval = FALSE}
colnames(sim)
unique(simZ$env_stoch)
summary(simZ$total_biomass)
summary(simZ$persistence)
summary(simZ$ct_alive)
summary(simZ$gini_max_int_alive)
```

```{r}
simZ %>%
  filter(rho == 0) %>%
  pivot_longer(
    c(stab_com, stab_pop, sync, total_biomass),
    names_to = "metric",
    values_to = "stab") %>%
  ggplot(aes(x = richness, y = stab, color = factor(env_stoch))) +
  geom_point() +
  facet_grid(rows = vars(metric), cols = vars(productivity), scales = "free_y") +
  labs(
    color = "Evironmental stochasticity",
    x = "Species richness",
    y = "Stability component"
  )
```

```{r}
simZ %>%
  filter(rho == 1) %>%
  mutate(stab_pop = 1/ avg_cv_sp) %>%
  pivot_longer(
    c(stab_com, stab_pop, sync, total_biomass),
    names_to = "metric",
    values_to = "stab") %>%
  ggplot(aes(x = richness, y = stab, color = factor(env_stoch))) +
  geom_point() +
  facet_grid(rows = vars(metric), cols = vars(productivity), scales = "free_y") +
  labs(
    color = "Evironmental stochasticity",
    x = "Species richness",
    y = "Stability component"
  )
```

```{r}
p_basic <- ~corrplot::corrplot(
  cor(simZ %>%
    filter(productivity == "enrichment") %>%
    select(c(S, richness, ct, Z, ct_alive, total_biomass, avg_max_int_alive,
        w_avg_tlvl_alive, max_tlvl_alive, rho, env_stoch, stab_com, avg_cv_sp,
        sync)),
    method = "spearman"),
  diag = FALSE, type = "upper"
)
p_enrich <- ~corrplot::corrplot(
  cor(simZ %>%
    filter(productivity == "basic") %>%
    select(c(S, richness, ct, Z, ct_alive, total_biomass, avg_max_int_alive,
        w_avg_tlvl_alive, max_tlvl_alive, rho, env_stoch, stab_com, avg_cv_sp,
        sync)),
    method = "spearman"),
  diag = FALSE, type = "upper"
)
plot_grid(
  ggdraw(p_basic),
  ggdraw(p_enrich),
  nrow = 2
)
```

- Main results:
  - Stability increases with richness but not connectance (no overall effect)
  - Richness increases stability by increasing maximum and average trophic
    level, but also by decreasing interaction strength and increasing asymetry
    in trophic interaction, and finally by increasing total biomass (I need to control for that).
  - Synchrony decreases with species richness, max and average trophic level but increases with
    average interaction strength and symetry of interaction distributions.
  - Population varibility decreases with species richness, average and maximum
    trophic level, asymetry of interaction strength but it is increased by
    average interaction strength
  - Same effects of food-web structure on both Synchrony and Population variability
    - Surpringly, as in Danet et al. (2021), connectance has little effect on
      temporal stability and its components!

# Covariation

```{r}
compute_rotated_pca <- function(.data = NULL, naxis = 2) {
  .data <- na.omit(.data)
  pca <- ade4::dudi.pca(as.data.frame(.data),
    scannf = FALSE, nf = naxis, center = TRUE, scale = TRUE
  )
  pca_rotated <- psych::principal(.data,
    rotate = "varimax", nfactors = naxis,
    scores = TRUE
  )

  return(list(normal = pca, rotated = pca_rotated))
}
```

```{r}
sim_sem <- simZ %>%
  filter(rho == 0) %>%
  mutate(across(c(async, sync, avg_cv_sp, stab_com, richness, avg_max_int_alive, gini_max_int), log, , .names = "log_{.col}")) %>%
  mutate(
    log1_avg_max_int_alive = log(avg_max_int_alive + 1),
    productivity_f = factor(productivity)

    ) %>%
  as.data.frame() %>%
  select(c(
      ct_alive, log_richness, max_tlvl_alive, Z,
      gini_max_int, avg_max_int_alive, log1_avg_max_int_alive, env_stoch,
      log_stab_com, log_sync, log_async, log_avg_cv_sp, productivity, productivity_f)
  )
scaled_sim_sem <- sim_sem %>%
  mutate(across(where(is.double), ~scale(.x)[,1]))
```

```{r}
pred_may <- "log1_avg_max_int_alive*log_richness*ct_alive*productivity + env_stoch"
rand <- "(1 | Z)"

mod_may_single <- lm(formula(str_replace_all(paste0("log_stab_com ~ ", pred_may), "\\*", "+")),
  scaled_sim_sem)
performance::check_collinearity(mod_may_single)
```

```{r}
mod_may <- glmmTMB(formula(paste0("log_stab_com ~ ", pred_may, " + ", rand)), scaled_sim_sem)

r2_nakagawa(mod_may)
par_mod_may <- parameters(mod_may)
plot(parameters(mod_may))
```

```{r}
mod_may_sync <- glmmTMB(formula(paste0("log_sync ~ ", pred_may, " + ", rand)),
  scaled_sim_sem)
r2_nakagawa(mod_may_sync)
plot(parameters(mod_may_sync))
```

```{r}
mod_may_cv_sp <- glmmTMB(formula(paste0("log_avg_cv_sp ~ ", pred_may, " + ", rand)),
  scaled_sim_sem)
r2_nakagawa(mod_may_cv_sp)
plot(parameters(mod_may_cv_sp))
```

## Rho model

- Should be include also the monotrophic communities ? I.e. Communities
  containing only plants?

```{r}
sim_rho <- simZ %>%
  #filter(ct_alive != 0) %>%
  mutate(across(c(async, sync, stab_pop, avg_cv_sp, stab_com, richness, avg_max_int_alive, gini_max_int), log, , .names = "log_{.col}")) %>%
  mutate(
    log1_avg_max_int_alive = log(avg_max_int_alive + 1),
    productivity_f = factor(productivity)

    ) %>%
  select(c(
      ct_alive, log_richness, max_tlvl_alive, Z, rho,
      gini_max_int, avg_max_int_alive, log1_avg_max_int_alive, env_stoch,
      log_stab_com, log_sync, log_async, log_stab_pop, log_avg_cv_sp, productivity, productivity_f,
      persistence)
  )
sim_rho_scaled <- sim_rho %>%
  mutate(across(where(is.double), ~scale(.x)[,1]))
```

```{r}
sim_rho %>%
  filter(ct_alive == 0, env_stoch == 0.5) %>%
  ggplot(aes(x = log_richness, y = log_stab_com)) +
  geom_point() +
  geom_smooth() +
  facet_grid(cols = vars(rho), rows = vars(productivity_f))
```

```{r}
pred_rho <- "log_richness +
ct_alive +
log1_avg_max_int_alive +
productivity +
rho +
I(rho^2) +
env_stoch +

log_richness:ct_alive +
log_richness:log1_avg_max_int_alive +
log_richness:productivity +
log_richness:rho +

log1_avg_max_int_alive:ct_alive +
log1_avg_max_int_alive:productivity +
log1_avg_max_int_alive:rho +

ct_alive:rho +
ct_alive:productivity +
rho:productivity +
I(rho^2):productivity +

log_richness:log1_avg_max_int_alive:ct_alive +
log_richness:log1_avg_max_int_alive:productivity +
log_richness:log1_avg_max_int_alive:rho +
log_richness:ct_alive:productivity +
log_richness:ct_alive:rho +
log_richness:rho:productivity +

log1_avg_max_int_alive:ct_alive:rho +
log1_avg_max_int_alive:ct_alive:productivity +
log1_avg_max_int_alive:rho:productivity +

ct_alive:rho:productivity"
rand <- "(1 | Z)"

mod_rho_single <- lm(formula(str_replace_all(paste0("log_stab_com ~ ", pred_rho), "(\\*|:)", "+")),
  sim_rho_scaled)
performance::check_collinearity(mod_rho_single)
```

```{r}
model2tibble <- function(x, response = NULL) {
  out <- parameters(x) %>%
    filter(Effects == "fixed", Parameter != "(Intercept)") %>%
    as_tibble() %>%
    janitor::clean_names() %>%
    mutate(
      response = var_replacement()[as.character(formula(x)[2])],
      parameter = str_replace_all(parameter, var_replacement()),
      count_int = str_count(parameter, ":"),
      effect_type = case_when(
        count_int == 0 ~ "single",
        count_int == 1 ~ "double",
        count_int == 2 ~ "triple",
        TRUE ~ "other"

      )
    ) %>%
    select(response, parameter:ci_high, effect_type)
    out
}
plot_effect_type <- function(x) {
  effect <- c("single", "double", "triple")
  p <- map(effect,
    function(y) {
      x %>%
        filter(effect_type == y) %>%
        arrange(coefficient) %>%
        ggplot(aes(
            y = reorder(parameter, -coefficient),
            x = coefficient,xmin = ci_low, xmax = ci_high,
            color = response
            )) +
        geom_vline(xintercept = 0, color = "black", linetype = "dashed", linewidth = 1.5) +
        geom_pointrange(position = position_dodge(width = 0.5)) +
        labs(x = "Standardized Coefficients")
    }
  )
  names(p) <- effect
  p
}
```

```{r}
mod_rho_scaled <- map(setNames(
    c("log_stab_com", "log_stab_pop", "log_async"),
    c("log_stab_com", "log_stab_pop", "log_async")
    ),
  function (x) {
    glmmTMB(formula(paste0(x, " ~ ", pred_rho, " + ", rand)),
      sim_rho_scaled)
  }
)
coef_mod_rho_scaled <- map_dfr(mod_rho_scaled, ~model2tibble(.x))
ti <- map_dfr(mod_rho_scaled, ~as.data.frame(r2_nakagawa(.x)), .id = "response") %>%
  mutate(response = var_replacement()[response])
p_mod_rho_scaled <- plot_effect_type(coef_mod_rho_scaled)
plot_grid(plotlist = p_mod_rho_scaled)
```

```{r}
mod_rho <- map(setNames(
    c("log_stab_com", "log_stab_pop", "log_async"),
    c("log_stab_com", "log_stab_pop", "log_async")
    ),
  function (x) {
    glmmTMB(formula(paste0(x, " ~ ", pred_rho, " + ", rand)),
      sim_rho)
  }
)
coef_mod_rho <- map_dfr(mod_rho, ~model2tibble(.x))
ti <- map_dfr(mod_rho, ~as.data.frame(r2_nakagawa(.x)), .id = "response") %>%
  mutate(response = var_replacement()[response])
p_mod_rho <- plot_effect_type(coef_mod_rho)
plot_grid(plotlist = p_mod_rho)
```

```{r, eval = FALSE}
sim_rho
```

```{r}
pred_list <- list(
  log_richness = c(log(1), log(5), log(10), log(20), log(30), log(50)),
  ct_alive = c(0, .10, .20),
  rho = c(0, .5, 1),
  log1_avg_max_int_alive = c(log(0 + 1), round(log(.81 + 1), 1), round(log(5.65 + 1), 1)),
  env_stoch = .5,
  productivity = c("basic", "enrichment"),
  Z = 100
)
pred_rho <- expand.grid(pred_list) %>%
  filter(
    !(log1_avg_max_int_alive == 0 & ct_alive != 0 | ct_alive == 0 & log1_avg_max_int_alive != 0)
  )

pred_rho$log_stab_com <- predict(mod_rho[["log_stab_com"]], newdata = pred_rho)

pred_rho %>%
  ggplot(aes(
      y = log_stab_com,
      x = log_richness,
      color = factor(ct_alive),
      linewidth = factor(log1_avg_max_int_alive)
      )) +
  geom_line() +
  facet_grid(
    rows = vars(rho),
    cols = vars(productivity)
  )
```

```{r, eval = FALSE}
mod_rho_tot <- glmmTMB(
  log_stab_com ~ log1_avg_max_int_alive*log_richness*ct_alive*rho*productivity +
    env_stoch + (1 | Z),
  sim_rho_scaled)
coef_mod_rho_tot <- model2tibble(mod_rho_tot)
coef_mod_rho_tot %>%
  ggplot(aes(y = parameter, x = coefficient, xmin = ci_low, xmax = ci_high)) +
  geom_pointrange() +
  labs(x = "Standardized Coefficients")
```

```{r}
sim_rho_pers <- simZ %>%
  mutate(across(c(S, ct, avg_max_int, gini_max_int), log, , .names = "log_{.col}")) %>%
  mutate(
    log1_avg_max_int = log(avg_max_int + 1),
    productivity_f = factor(productivity)
    ) %>%
  as.data.frame() %>%
  select(c(
      log_S, ct, max_tlvl, Z, rho,
      gini_max_int, avg_max_int, log1_avg_max_int, env_stoch,
      productivity, productivity_f,
      persistence)
  )
sim_rho_pers_scaled <- sim_rho_pers %>%
  mutate(across(where(is.double), ~scale(.x)[,1]))

pred_rho_pers <- "log1_avg_max_int * log_S * rho +
log1_avg_max_int * log_S * productivity +
log1_avg_max_int * log_S * ct +
log1_avg_max_int * ct * rho +
log1_avg_max_int * ct * rho +
log1_avg_max_int * ct * productivity +
log_S * rho * productivity +
log_S * ct * productivity +
log_S * ct * rho +
env_stoch"
rand <- "(1 | Z)"

mod_rho_pers <- glmmTMB(formula(paste0("persistence ~ ", pred_rho_pers, " + ", rand)),
  sim_rho_pers_scaled)

r2_nakagawa(mod_rho_pers)
plot(parameters(mod_rho_pers))
```

## Regression tree

```{r}
```

```{r, eval = FALSE}
library(rpart)
library(rpart.plot)
set.seed(123)
m1 <- rpart(
  formula = log_stab_com ~ .,
  data    = sim_rho %>%
    select(log_stab_com, log_richness, ct_alive, rho, env_stoch,
      gini_max_int, max_tlvl_alive, log1_avg_max_int_alive, productivity),
  method  = "anova"
  )
rpart.plot(m1)
```

```{r, eval=FALSE}
bagged_m1 <- bagging(
  formula = log_stab_com ~ .,
  data    = sim_rho %>%
    select(log_stab_com, log_richness, ct_alive, rho, env_stoch,
      gini_max_int, max_tlvl_alive, log1_avg_max_int_alive, productivity),
  coob    = TRUE
)
```

```{r}
library(ipred)
ctrl <- trainControl(method = "cv",  number = 10)

# CV bagged model
bagged_cv <- train(
  Sale_Price ~ .,
  data = ames_train,
  method = "treebag",
  trControl = ctrl,
  importance = TRUE
  )
```

# Effect of species richness

```{r}
sim_rich_prod <- simZ %>%
  filter(
    Z == 100,
    env_stoch == .5,
    rho %in% c(0, 1)
  )
```

```{r}
p_sim_rich_prod <- sim_rich_prod %>%
  mutate(stab_pop = 1/ avg_cv_sp) %>%
  pivot_longer(
    c(stab_com, stab_pop, sync),
    names_to = "metric",
    values_to = "stab") %>%
  mutate(
    metric = var_replacement()[metric],
    rho = rho_replacement()[as.character(rho)]
  ) %>%
  ggplot(aes(x = richness, y = stab, color = factor(productivity))) +
  geom_point(alpha = .1, shape = 20) +
  geom_smooth() +
  facet_grid(rows = vars(metric), cols = vars(rho), scales = "free_y") +
  labs(
    color = "Productivity",
    x = "Species richness",
    y = "Stability component"
  ) +
  scale_x_continuous(
    sec.axis = sec_axis(~ . ,
      name = "Response diversity",
      breaks = NULL,
      labels = NULL)
  )
ggsave(
  filename = here::here("fig", "stability_rich.pdf"),
  p_sim_rich_prod,
  scale = 1.8,
  width = 88,
  height = 88 * .6,
  units = "mm"
)
```