### 2023-05-14 report/cs_analysis.Rmd
Add targets for result analysis

---
title: Simulation connectance richness
author: Alain Danet
output:
  bookdown::pdf_document2
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = TRUE,
  fig.path="figures/"
)
```

```{r, results = FALSE}
library(tidyverse)
library(magrittr)
library(ggcorrplot)
library(cowplot)
library(here)
library(arrow)
library(piecewiseSEM)
library(semEff)
library(easystats)
library(glmmTMB)
library(targets)

library(rpart)
library(rpart.plot)
library(caret)
library(ipred)

mytheme <- theme_cowplot(font_size = 8) +
background_grid() +
theme(legend.position = "bottom")
theme_set(mytheme)
source(here::here("R", "network_metrics.R"))
source(here::here("R", "string_replacements.R"))
```

```{r, eval = FALSE}
colnames(sim)
unique(simZ$env_stoch)
summary(simZ$total_biomass)
summary(simZ$persistence)
summary(simZ$ct_alive)
summary(simZ$gini_max_int_alive)
```

# Introduction

- Response diversity is key in driving stability in ecological communities
- Response diversity control asynchrony with richness
- Mainly done in plants
- Food-webs structure can also modulate synchrony by predator-prey relationships
- Interaction strength, connectance, richness and enrichment can also drive more

# Research questions

##  What are the mecanisms driving stability in food-webs ?

- How structure of food-webs (connectance, species richness, prey-predator
     size ratio) gives emergence to food-web structure (max trophic level,
     weighted avg trophic level, omnivory) ?
- How food-web structure influence biomass fluxes (avg interaction
     strength, sd of interaction strength, asymetry)?
- How food-web structure and biomass fluxes affect stability components?
     total biomass, species variability, statistical averaging, evenness,
     compensatory dynamics (interaction and environment)?

# What is the effects of species richness on stability facing multiple stressors in competitive and food-webs communities?

- How do species richness drive average population stability and asynchrony
  in competitive and food-webs communities ?
- How is it affected by enrichment and response diversity?

# Material & Methods

## Bioenergetic model

TO Complete

- Important bits:
  - Niche model (Connectance highly constrained by richness)
  - Type III functional response
  - No predator interference
  - Standardize carrying capacity? (Remove the effect of species richness on
    total biomass, i.e. remove the effect of niche complementarity on biomass
    produced by primary producers)

## Variables

- S: 5 to 60
- C: .02 to .3
- K: 1 and 20
- $\rho$: 0 to 1
- Z: 10 to 1000

```{r}
tar_load(simZ)
tar_load(sim_rho)
tar_load(sim_rho_scaled)
```

```{r}
simZ %>%
  filter(rho == 0) %>%
  pivot_longer(
    c(stab_com, stab_pop, sync, total_biomass),
    names_to = "metric",
    values_to = "stab") %>%
  ggplot(aes(x = richness, y = stab, color = factor(env_stoch))) +
  geom_point() +
  facet_grid(rows = vars(metric), cols = vars(productivity), scales = "free_y") +
  labs(
    color = "Evironmental stochasticity",
    x = "Species richness",
    y = "Stability component"
  )
```

```{r}
simZ %>%
  filter(rho == 1) %>%
  mutate(stab_pop = 1/ avg_cv_sp) %>%
  pivot_longer(
    c(stab_com, stab_pop, sync, total_biomass),
    names_to = "metric",
    values_to = "stab") %>%
  ggplot(aes(x = richness, y = stab, color = factor(env_stoch))) +
  geom_point() +
  facet_grid(rows = vars(metric), cols = vars(productivity), scales = "free_y") +
  labs(
    color = "Evironmental stochasticity",
    x = "Species richness",
    y = "Stability component"
  )
```

```{r}
cor_enrich <- cor(simZ %>%
    filter(productivity == "enrichment") %>%
    select(c(S, richness, ct, Z, ct_alive, total_biomass, avg_max_int_alive,
        w_avg_tlvl_alive, max_tlvl_alive, rho, env_stoch, stab_com, avg_cv_sp,
        sync)),
    method = "spearman")
cor_basic <- cor(simZ %>%
    filter(productivity == "basic") %>%
    select(c(S, richness, ct, Z, ct_alive, total_biomass, avg_max_int_alive,
        w_avg_tlvl_alive, max_tlvl_alive, rho, env_stoch, stab_com, avg_cv_sp,
        sync)),
    method = "spearman")
p_cor_basic <- ggcorrplot(cor_basic, type = "lower",
   lab = TRUE)
p_cor_enrich <- ggcorrplot(cor_enrich, type = "lower",
   lab = TRUE)
p_cor <- plot_grid(
  p_cor_basic,
  p_cor_enrich,
  ncol = 2
)
ggsave(here("report", "figures", "p_cor.pdf"),
  plot = p_cor,
  device = NULL,
  path = NULL,
  scale = 3.0,
  units = "mm",
  width = 140,
  height = 140 * .5,
  dpi = 300,
  limitsize = TRUE,
  bg = NULL
)
```

- Main results:
  - Stability increases with richness but not connectance (no overall effect)
  - Richness increases stability by increasing maximum and average trophic
    level, but also by decreasing interaction strength and increasing asymetry
    in trophic interaction, and finally by increasing total biomass (I need to control for that).
  - Synchrony decreases with species richness, max and average trophic level but increases with
    average interaction strength and symetry of interaction distributions.
  - Population varibility decreases with species richness, average and maximum
    trophic level, asymetry of interaction strength but it is increased by
    average interaction strength
  - Same effects of food-web structure on both Synchrony and Population variability
    - Surpringly, as in Danet et al. (2021), connectance has little effect on
      temporal stability and its components!

# Effect of species richness

```{r}
sim_rich_prod <- simZ %>%
  filter(
    Z == 100,
    env_stoch == .5,
    rho %in% c(0, 1)
  )
```

```{r}
p_sim_rich_prod <- sim_rich_prod %>%
  mutate(stab_pop = 1/ avg_cv_sp) %>%
  pivot_longer(
    c(stab_com, stab_pop, sync),
    names_to = "metric",
    values_to = "stab") %>%
  mutate(
    metric = var_replacement()[metric],
    rho = rho_replacement()[as.character(rho)]
  ) %>%
  ggplot(aes(x = richness, y = stab, color = factor(productivity))) +
  geom_point(alpha = .1, shape = 20) +
  geom_smooth() +
  facet_grid(rows = vars(metric), cols = vars(rho), scales = "free_y") +
  labs(
    color = "Productivity",
    x = "Species richness",
    y = "Stability component"
  ) +
  scale_x_continuous(
    sec.axis = sec_axis(~ . ,
      name = "Response diversity",
      breaks = NULL,
      labels = NULL)
  )
ggsave(
  filename = here::here("fig", "stability_rich.pdf"),
  p_sim_rich_prod,
  scale = 1.8,
  width = 88,
  height = 88 * .6,
  units = "mm"
)
```

```{r}
p_sim_rich_prod <- sim_rich_prod %>%
  filter(ct_alive != 0) %>%
  mutate(stab_pop = 1/ avg_cv_sp) %>%
  pivot_longer(
    c(stab_com, stab_pop, sync),
    names_to = "metric",
    values_to = "stab") %>%
  mutate(
    metric = var_replacement()[metric],
    rho = rho_replacement()[as.character(rho)]
  ) %>%
  ggplot(aes(x = richness, y = stab, color = factor(productivity))) +
  geom_point(alpha = .1, shape = 20) +
  geom_smooth() +
  facet_grid(rows = vars(metric), cols = vars(rho), scales = "free_y") +
  labs(
    color = "Productivity",
    x = "Species richness",
    y = "Stability component"
  ) +
  scale_x_continuous(
    sec.axis = sec_axis(~ . ,
      name = "Response diversity",
      breaks = NULL,
      labels = NULL)
  )
ggsave(
  filename = here::here("fig", "stability_rich_fw_only.pdf"),
  p_sim_rich_prod,
  scale = 1.8,
  width = 88,
  height = 88 * .6,
  units = "mm"
)
```

```{r}
p_sim_rich_prod_p_only <- sim_rich_prod %>%
  filter(ct_alive == 0) %>%
  mutate(stab_pop = 1/ avg_cv_sp) %>%
  pivot_longer(
    c(stab_com, stab_pop, sync),
    names_to = "metric",
    values_to = "stab") %>%
  mutate(
    metric = var_replacement()[metric],
    rho = rho_replacement()[as.character(rho)]
  ) %>%
  ggplot(aes(x = richness, y = stab, color = factor(productivity))) +
  geom_point(alpha = .1, shape = 20) +
  geom_smooth() +
  facet_grid(rows = vars(metric), cols = vars(rho), scales = "free_y") +
  labs(
    color = "Productivity",
    x = "Species richness",
    y = "Stability component"
  ) +
  scale_x_continuous(
    sec.axis = sec_axis(~ .,
      name = "Response diversity",
      breaks = NULL,
      labels = NULL)
  )
ggsave(
  filename = here::here("fig", "stability_rich_p_only.pdf"),
  p_sim_rich_prod_p_only,
  scale = 1.8,
  width = 88,
  height = 88 * .6,
  units = "mm"
)
```

# Covariation

```{r}
compute_rotated_pca <- function(.data = NULL, naxis = 2) {
  .data <- na.omit(.data)
  pca <- ade4::dudi.pca(as.data.frame(.data),
    scannf = FALSE, nf = naxis, center = TRUE, scale = TRUE
  )
  pca_rotated <- psych::principal(.data,
    rotate = "varimax", nfactors = naxis,
    scores = TRUE
  )

  return(list(normal = pca, rotated = pca_rotated))
}
```

```{r}
sim_sem <- simZ %>%
  filter(rho == 0) %>%
  mutate(across(c(async, sync, avg_cv_sp, stab_com, richness, avg_max_int_alive, gini_max_int), log, , .names = "log_{.col}")) %>%
  mutate(
    log1_avg_max_int_alive = log(avg_max_int_alive + 1),
    productivity_f = factor(productivity)

    ) %>%
  as.data.frame() %>%
  select(c(
      ct_alive, log_richness, max_tlvl_alive, Z,
      gini_max_int, avg_max_int_alive, log1_avg_max_int_alive, env_stoch,
      log_stab_com, log_sync, log_async, log_avg_cv_sp, productivity, productivity_f)
  )
scaled_sim_sem <- sim_sem %>%
  mutate(across(where(is.double), ~scale(.x)[,1]))
```

```{r}
pred_may <- "log1_avg_max_int_alive*log_richness*ct_alive*productivity + env_stoch"
rand <- "(1 | Z)"

mod_may_single <- lm(formula(str_replace_all(paste0("log_stab_com ~ ", pred_may), "\\*", "+")),
  scaled_sim_sem)
performance::check_collinearity(mod_may_single)
```

```{r}
mod_may <- glmmTMB(formula(paste0("log_stab_com ~ ", pred_may, " + ", rand)), scaled_sim_sem)

r2_nakagawa(mod_may)
par_mod_may <- parameters(mod_may)
plot(parameters(mod_may))
```

```{r}
mod_may_sync <- glmmTMB(formula(paste0("log_sync ~ ", pred_may, " + ", rand)),
  scaled_sim_sem)
r2_nakagawa(mod_may_sync)
plot(parameters(mod_may_sync))
```

```{r}
mod_may_cv_sp <- glmmTMB(formula(paste0("log_avg_cv_sp ~ ", pred_may, " + ", rand)),
  scaled_sim_sem)
r2_nakagawa(mod_may_cv_sp)
plot(parameters(mod_may_cv_sp))
```

## Rho model

- Should be include also the monotrophic communities ? I.e. Communities
  containing only plants?

```{r}
sim_rho %>%
  filter(ct_alive == 0, env_stoch == 0.5) %>%
  ggplot(aes(x = log_richness, y = log_stab_com)) +
  geom_point() +
  geom_smooth() +
  facet_grid(cols = vars(rho), rows = vars(productivity_f))
```

```{r}
pred_rho <- "log_richness +
ct_alive +
log1_avg_max_int_alive +
productivity +
rho +
I(rho^2) +
env_stoch +

log_richness:ct_alive +
log_richness:log1_avg_max_int_alive +
log_richness:productivity +
log_richness:rho +

log1_avg_max_int_alive:ct_alive +
log1_avg_max_int_alive:productivity +
log1_avg_max_int_alive:rho +

ct_alive:rho +
ct_alive:productivity +
rho:productivity +
I(rho^2):productivity +

log_richness:log1_avg_max_int_alive:ct_alive +
log_richness:log1_avg_max_int_alive:productivity +
log_richness:log1_avg_max_int_alive:rho +
log_richness:ct_alive:productivity +
log_richness:ct_alive:rho +
log_richness:rho:productivity +

log1_avg_max_int_alive:ct_alive:rho +
log1_avg_max_int_alive:ct_alive:productivity +
log1_avg_max_int_alive:rho:productivity +

ct_alive:rho:productivity"
rand <- "(1 | Z)"

mod_rho_single <- lm(formula(str_replace_all(paste0("log_stab_com ~ ", pred_rho), "(\\*|:)", "+")),
  sim_rho_scaled)
performance::check_collinearity(mod_rho_single)
```

```{r}
model2tibble <- function(x, response = NULL) {
  out <- parameters(x) %>%
    filter(Effects == "fixed", Parameter != "(Intercept)") %>%
    as_tibble() %>%
    janitor::clean_names() %>%
    mutate(
      response = var_replacement()[as.character(formula(x)[2])],
      parameter = str_replace_all(parameter, var_replacement()),
      count_int = str_count(parameter, ":"),
      effect_type = case_when(
        count_int == 0 ~ "single",
        count_int == 1 ~ "double",
        count_int == 2 ~ "triple",
        TRUE ~ "other"

      )
    ) %>%
    select(response, parameter:ci_high, effect_type)
    out
}
plot_effect_type <- function(x) {
  effect <- c("single", "double", "triple")
  p <- map(effect,
    function(y) {
      x %>%
        filter(effect_type == y) %>%
        arrange(coefficient) %>%
        ggplot(aes(
            y = reorder(parameter, -coefficient),
            x = coefficient,xmin = ci_low, xmax = ci_high,
            color = response
            )) +
        geom_vline(xintercept = 0, color = "black", linetype = "dashed", linewidth = 1.5) +
        geom_pointrange(position = position_dodge(width = 0.5)) +
        labs(x = "Standardized Coefficients")
    }
  )
  names(p) <- effect
  p
}
```

```{r}
mod_rho_scaled <- map(setNames(
    c("log_stab_com", "log_stab_pop", "log_async"),
    c("log_stab_com", "log_stab_pop", "log_async")
    ),
  function (x) {
    glmmTMB(formula(paste0(x, " ~ ", pred_rho, " + ", rand)),
      sim_rho_scaled)
  }
)
coef_mod_rho_scaled <- map_dfr(mod_rho_scaled, ~model2tibble(.x))
ti <- map_dfr(mod_rho_scaled, ~as.data.frame(r2_nakagawa(.x)), .id = "response") %>%
  mutate(response = var_replacement()[response])
p_mod_rho_scaled <- plot_effect_type(coef_mod_rho_scaled)
plot_grid(plotlist = p_mod_rho_scaled)
```

```{r}
mod_rho <- map(setNames(
    c("log_stab_com", "log_stab_pop", "log_async"),
    c("log_stab_com", "log_stab_pop", "log_async")
    ),
  function (x) {
    glmmTMB(formula(paste0(x, " ~ ", pred_rho, " + ", rand)),
      sim_rho)
  }
)
coef_mod_rho <- map_dfr(mod_rho, ~model2tibble(.x))
ti <- map_dfr(mod_rho, ~as.data.frame(r2_nakagawa(.x)), .id = "response") %>%
  mutate(response = var_replacement()[response])
p_mod_rho <- plot_effect_type(coef_mod_rho)
plot_grid(plotlist = p_mod_rho)
```

```{r, eval = FALSE}
sim_rho
```

```{r}
pred_list <- list(
  log_richness = c(log(1), log(5), log(10), log(20), log(30), log(50)),
  ct_alive = c(0, .10, .20),
  rho = c(0, .5, 1),
  log1_avg_max_int_alive = c(log(0 + 1), round(log(.81 + 1), 1), round(log(5.65 + 1), 1)),
  env_stoch = .5,
  productivity = c("basic", "enrichment"),
  Z = 100
)
pred_rho <- expand.grid(pred_list) %>%
  filter(
    !(log1_avg_max_int_alive == 0 & ct_alive != 0 | ct_alive == 0 & log1_avg_max_int_alive != 0)
  )

pred_rho$log_stab_com <- predict(mod_rho[["log_stab_com"]], newdata = pred_rho)

pred_rho %>%
  ggplot(aes(
      y = log_stab_com,
      x = log_richness,
      color = factor(ct_alive),
      linewidth = factor(log1_avg_max_int_alive)
      )) +
  geom_line() +
  facet_grid(
    rows = vars(rho),
    cols = vars(productivity)
  )
```

```{r, eval = FALSE}
mod_rho_tot <- glmmTMB(
  log_stab_com ~ log1_avg_max_int_alive*log_richness*ct_alive*rho*productivity +
    env_stoch + (1 | Z),
  sim_rho_scaled)
coef_mod_rho_tot <- model2tibble(mod_rho_tot)
coef_mod_rho_tot %>%
  ggplot(aes(y = parameter, x = coefficient, xmin = ci_low, xmax = ci_high)) +
  geom_pointrange() +
  labs(x = "Standardized Coefficients")
```

```{r}
sim_rho_pers <- simZ %>%
  mutate(across(c(S, ct, avg_max_int, gini_max_int), log, , .names = "log_{.col}")) %>%
  mutate(
    log1_avg_max_int = log(avg_max_int + 1),
    productivity_f = factor(productivity)
    ) %>%
  as.data.frame() %>%
  select(c(
      log_S, ct, max_tlvl, Z, rho,
      gini_max_int, avg_max_int, log1_avg_max_int, env_stoch,
      productivity, productivity_f,
      persistence)
  )
sim_rho_pers_scaled <- sim_rho_pers %>%
  mutate(across(where(is.double), ~scale(.x)[,1]))

pred_rho_pers <- "log1_avg_max_int * log_S * rho +
log1_avg_max_int * log_S * productivity +
log1_avg_max_int * log_S * ct +
log1_avg_max_int * ct * rho +
log1_avg_max_int * ct * rho +
log1_avg_max_int * ct * productivity +
log_S * rho * productivity +
log_S * ct * productivity +
log_S * ct * rho +
env_stoch"
rand <- "(1 | Z)"

mod_rho_pers <- glmmTMB(formula(paste0("persistence ~ ", pred_rho_pers, " + ", rand)),
  sim_rho_pers_scaled)

r2_nakagawa(mod_rho_pers)
plot(parameters(mod_rho_pers))
```

## Regression tree

```{r}
```

```{r, eval = FALSE}
library(rpart)
library(rpart.plot)
set.seed(123)
m1 <- rpart(
  formula = log_stab_com ~ .,
  data    = sim_rho %>%
    select(log_stab_com, log_richness, ct_alive, rho, env_stoch,
      gini_max_int, max_tlvl_alive, log1_avg_max_int_alive, productivity),
  method  = "anova"
  )
rpart.plot(m1)
```

```{r, eval=FALSE}
bagged_m1 <- bagging(
  formula = log_stab_com ~ .,
  data    = sim_rho %>%
    select(log_stab_com, log_richness, ct_alive, rho, env_stoch,
      gini_max_int, max_tlvl_alive, log1_avg_max_int_alive, productivity),
  coob    = TRUE
)
```

```{r}
library(ipred)
ctrl <- trainControl(method = "cv",  number = 10)

# CV bagged model
bagged_cv <- train(
  Sale_Price ~ .,
  data = ames_train,
  method = "treebag",
  trControl = ctrl,
  importance = TRUE
  )
```

# Effect of species richness

```{r}
ctrl <- trainControl(method = "cv",  number = 10)

# CV bagged model
bagged_cv <- train(
  stab_com ~ richness + ct_alive + rho + productivity + max_tlvl_alive +
    avg_max_int_alive,
  data = simZ,
  cp = .2,
  maxdepth = 6,
  method = "treebag",
  trControl = ctrl,
  importance = TRUE
  )
plot(varImp(bagged_cv), 20)
plot(bagged_cv)
```

```{r}
m1 <- rpart(
  formula = stab_com ~ richness + ct_alive + rho + productivity + max_tlvl_alive +
    avg_max_int_alive,
  data    = simZ,
  method  = "anova"
  )
rpart.plot(m1)
```