### 2023-05-19 report/cs_analysis.Rmd
Sim with higher persistence hopefully: low Z and K gradient

---
title: Simulation connectance richness
author: Alain Danet
output:
  bookdown::pdf_document2
always_allow_html: true
bibliography: references.bib
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.pos = "H",  # pdf mode
  #fig.dim = c(7, 7),
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = TRUE,
  fig.path="figures/"
)
```

```{r, results = FALSE}
library(targets)
library(tidyverse)
library(magrittr)
library(ggcorrplot)
library(cowplot)
library(here)
library(arrow)
library(piecewiseSEM)
library(semEff)
library(easystats)
library(glmmTMB)

library(rpart)
library(rpart.plot)
library(caret)
library(ipred)
library(kableExtra)

mytheme <- theme_cowplot(font_size = 8) +
background_grid() +
theme(legend.position = "bottom")
theme_set(mytheme)
source(here::here("R", "network_metrics.R"))
source(here::here("R", "string_replacements.R"))
```

# Introduction

Understanding the drivers of temporal stability is a key question in Ecology.
BEF theory shows that species richness is a key driver to understand the
temporal stability of community biomass.

Species richness was demonstrated to increase community stability by driving higher
asynchrony in community in various assemblages and ecosystems. In particular,
high response diversity, which drives independant species fluctuations, seems to a key
factor driving the positive effect of species richness of stability
[@loreau_species_2008; @thibaut_understanding_2013; @zhao_biodiversity_2022]. However,
those studies were often limited to monotrophic assemblages. As natural
communities are often composed of multiple trophic level, our understanding of
community stability needs to incorporate food-web structure.

Food-web structure has been shown to have tremendous effect on community
stability. The degree of generalism of predator was shown to modulate
stability-richness relationships [@thebault_trophic_2005]. The strength and distribution of
interaction strength was also shown to be strong driver of community stability
[@mccann_weak_1998]. Omnivory was also shown to drive higher stability
[@mccann_reevaluating_1997]. However, it is quite unkown how those food-web
structure elements would affect stability-species richness relationships.

The effect of environmental stressors can also strongly affect community
variability. Higher environmental stress and variability may dampens the effect
of richness on stability [@loreau_biodiversity_2001]. Lastly, productivity of a
system, as the quantity of biomass input in the ecosystem, can have tremendous
effect on both community structure and stabilty [@lindeman_trophic-dynamic_1942;
@binzer_dynamics_2012; @binzer_interactive_2016]. Productivity can increase
species richness in the food-web by increasing the level of energy input
[@lindeman_trophic-dynamic_1942] but also destabilises community biomass by
increasing population variability, a behavior coined as the paradox of
enrichment [@rosenzweig_paradox_1971]. A better understanding of the effect of
species richness on community stability should then include environmental
gradient and stressors.

Community stability can itself be partitioned in several components, i.e.
population stability and population asynchrony [@thibaut_understanding_2013;
@zhao_biodiversity_2022]. Population stability can be itself partioned in total
biomass and population variability effect. Population asynchrony can further be
partioned in statistical averaging effect and compensatory dynamics effects
[@zhao_biodiversity_2022]. Partitioning stability allows to better understand
how food-web structure and external factors affect stability. For exemple,
environmental stochasticity is expected to primarily affect population
stability, while response diversity should affect primarily compensatory
dynamics.

```{r}
concept_cap <- "Food-web structure, response diversity and enviromental gradient
can modulate stability-richness relationships. (a & b) Food-web structure may
modulate stability-richness relationships additively (a) or interactively (b).
(c) Decrease in response diversity and increase in environmental variability
should dampen stability-richness relationships. While the loss of response
diversity should dampen the effect of species richness on asynchrony, the
increase in environmental variability should decrease primarily population
stability. (d) Increase in productivity should increase species richness
(trophic energy hypothesis) but also dampens the relationship between stability
and species richness (paradox of enrichment). (e) Summary of the framework, how
food-web structure and external perturbations modulate the effects of species
richness on community stability. More precisely, how they affect the different
components of stability: total biomass, population variability, statistical
averaging effects and compensatory dynamics."
```

```{r, fig.cap=concept_cap}
knitr::include_graphics(here("report", "figures", "conceptual_figures.pdf"))
```

```{r}
tribble(
  ~Type, ~`Food-web structure`, ~`Effect on stability-richness`,
  "Food-web structure", "Connectance", "Strengthen (Miele et al., 2016) / No effect (Thébault & Loreau, 2005)",
  "Food-web structure", "Omnivory", "Strengthen (Mccann et al., 1997)",
  "Food-web structure", "Interaction strength", "Destabilising (Mccann et al., 1998) / Stabilising in
  productive (Binzer et al., 2012; Binzer et al., 2016; Tabi et al., 2019)",
  "Response diversity", "", "Dampens",
  "Productivity", "", "Dampens (Binzer et al., 2012; Binzer et al., 2016; Tabi et al., 2019)",
  "Environmental variability", "", "Dampens",
)
```

```{r, eval = FALSE}
tar_load(sim_prep)
# test for disconnected species
ti <- sim_prep %>%
  filter(max_tlvl > 1) %>%
  mutate(
    disconnected_prod = map_lgl(max_int_alive,
      # which species has no predator (colsum) is not a predator
      ~any(colSums(.x) == 0 & rowSums(.x) == 0)
    )
  )

ti %>%
  ggplot(aes(x = as.factor(disconnected_prod), y = stab_com)) +
  geom_boxplot()
```

```{r}
tar_load(sim_stab)
sim_stab %<>%
  mutate(
    sae_even = map_dbl(sae, ~.x["even"]),
    evenness_sae = map_dbl(sae, ~.x["eveness"]),
    sae_total = map_dbl(sae, ~.x["total"])
  )
tar_load(sim)
sim_ok <- sim %>%
  select(-stab_com, -async) %>%
  left_join(sim_stab) %>%
  mutate(inv_sum_pop_var = pop_stab / total_biomass)
```

```{r}
sim_ok %>%
  filter(ct_alive != 0, rho == 0, env_stoch == .1)
sim_sem <- sim_ok %>%
  filter(ct_alive != 0) %>%
  mutate(across(c(stab_com, async, pop_stab, total_biomass, inv_sum_pop_var, sae_even, evenness_sae, cpe_env, cpe_int), log, , .names = "log_{.col}")) %>%
  na.omit()
```

```{r}
sem_list_basic <- list(
  w_avg_tlvl_alive = lm(w_avg_tlvl_alive ~ richness + ct_alive, sim_sem),
#  avg_omnivory = lm(avg_omnivory ~ richness + ct_alive + w_avg_tlvl_alive, sim_sem),
  sd_max_int_alive = lm(sd_max_int_alive ~ richness + ct_alive + w_avg_tlvl_alive + Z, sim_sem),
  avg_max_int_alive = lm(avg_max_int_alive ~ richness + ct_alive + w_avg_tlvl_alive + Z, sim_sem),
  log_inv_sum_pop_var = lm(log_inv_sum_pop_var ~ richness + avg_max_int_alive + sd_max_int_alive + w_avg_tlvl_alive + ct_alive + env_stoch + rho, sim_sem),
  log_total_biomass = lm(log_total_biomass ~ richness + avg_max_int_alive + sd_max_int_alive + w_avg_tlvl_alive + ct_alive + env_stoch + rho, sim_sem),
  log_sae_even = lm(log_sae_even ~ richness, sim_sem),
  log_evenness_sae = lm(log_evenness_sae ~ richness + avg_max_int_alive + sd_max_int_alive + w_avg_tlvl_alive + ct_alive, sim_sem),
  log_cpe_env = lm(log_cpe_env ~ richness + avg_max_int_alive + sd_max_int_alive + w_avg_tlvl_alive + ct_alive + rho + env_stoch, sim_sem),
  log_cpe_int = lm(log_cpe_int ~ richness + avg_max_int_alive + sd_max_int_alive + w_avg_tlvl_alive + ct_alive, sim_sem),
  log_async = lm(log_async ~ log_sae_even + log_evenness_sae + log_cpe_env + log_cpe_int, sim_sem),
  #log_stab_pop = lm(log_pop_stab ~ log_total_biomass + log_inv_sum_pop_var, sim_sem),
  log_stab_com = lm(log_stab_com ~ log_total_biomass + log_inv_sum_pop_var + log_async, sim_sem)
)
sem_basic <- as.psem(sem_list_basic)
ti <- summary(sem_basic)$coefficients[, c(1, 2, 8, 9)]
colnames(ti)[4] <- "Signif"
ti
```
```{r, eval = FALSE}
lm(
  log_stab_com ~ log_total_biomass + log_inv_sum_pop_var + log_async,
  sim_ok %>%
    filter(ct_alive == 0) %>%
    mutate(across(c(stab_com, async, pop_stab, total_biomass, inv_sum_pop_var, sae_even, evenness_sae, cpe_env, cpe_int), log, , .names = "log_{.col}"))%>%
    mutate(across(where(is.double), ~scale(.x)[,1]))
)
```

```{r}
sim_stab %>%
  ggplot(aes(x = as.factor(rho), y = cpe_env)) +
  geom_boxplot() +
  geom_hline(yintercept = 1)
sim_stab %>%
  ggplot(aes(x = as.factor(env_stoch), y = cpe)) +
  geom_boxplot() +
  geom_hline(yintercept = 1)

sim_stab %>%
  ggplot(aes(x = as.factor(env_stoch), y = stab_com)) +
  geom_boxplot() +
  geom_hline(yintercept = 1)
sim_stab %>%
  ggplot(aes(x = as.factor(rho), y = stab_com)) +
  geom_boxplot() +
  geom_hline(yintercept = 1)
sim_stab %>%
  ggplot(aes(x = as.factor(rho), y = evenness_sae)) +
  geom_boxplot() +
  geom_hline(yintercept = 1)
sim_stab %>%
  ggplot(aes(x = as.factor(rho), y = sae_total)) +
  geom_boxplot() +
  geom_hline(yintercept = 1)

sim_stab %>%
  ggplot(aes(x = as.factor(env_stoch), y = sae_total)) +
  geom_boxplot() +
  geom_hline(yintercept = 1)
sim_stab %>%
  ggplot(aes(x = as.factor(env_stoch), y = evenness_sae)) +
  geom_boxplot() +
  geom_hline(yintercept = 1)

sim_stab %>%
  ggplot(aes(x = as.factor(env_stoch), y = pop_stab)) +
  geom_boxplot() +
  geom_hline(yintercept = 1)
sim_stab %>%
  ggplot(aes(x = as.factor(rho), y = pop_stab)) +
  geom_boxplot() +
  geom_hline(yintercept = 1)
```

```{r}
sim %>%
  filter(ct_alive != 0) %>%
  ggplot(aes(y = avg_omnivory, x = richness)) +
  geom_point()

sim %>%
  filter(ct_alive != 0) %>%
  ggplot(aes(y = avg_omnivory, x = ct_alive)) +
  geom_point()

sim %>%
  filter(ct_alive != 0) %>%
  ggplot(aes(x = avg_omnivory, y = 1/avg_cv_sp)) +
  geom_point()

sim %>%
  filter(ct_alive != 0) %>%
  ggplot(aes(x = avg_omnivory, y = sync)) +
  geom_point()

sim %>%
  filter(ct_alive != 0) %>%
  ggplot(aes(x = Z, y = max_tlvl_alive)) +
  geom_point()
```

```{r}
p_sim_sync_rich <- sim %>%
  filter(ct_alive != 0) %>%
  ggplot(aes(x = richness, y = sync, color = as.factor(env_stoch))) +
  geom_point() +
  geom_smooth() +
  facet_grid(cols = vars(productivity), rows = vars(rho))

p_sim_pop_rich <- sim %>%
  filter(ct_alive != 0) %>%
  ggplot(aes(x = richness, y = 1/avg_cv_sp, color = as.factor(env_stoch))) +
  geom_point() +
  geom_smooth() +
  facet_grid(cols = vars(productivity), rows = vars(rho))
```

```{r}
p_sim_stab_rich <- sim %>%
  filter(ct_alive != 0) %>%
  ggplot(aes(x = richness, y = stab_com, color = as.factor(env_stoch))) +
  geom_point() +
  geom_smooth() +
  facet_grid(cols = vars(productivity), rows = vars(rho))
ggsave(
  filename = here("report", "figures", "sim_stab_richnness_rho_stoch.pdf"),
  p_sim_stab_rich,
  width = 80,
  height = 80 * 4,
  units = "mm",
  scale = 2.4
)
```

```{r}
knitr::include_graphics(here::here("report", "figures", "sim_stab_richnness_rho_stoch.pdf"))
```

# Annexes

# Research questions

For me, there are two ways to see the question.

##  What are the mecanisms driving stability in food-webs ?

- How structure of food-webs (connectance, species richness, prey-predator
     size ratio) gives emergence to food-web structure (max trophic level,
     weighted avg trophic level, omnivory) ?
- How food-web structure influence biomass fluxes (avg interaction
     strength, sd of interaction strength, asymetry)?
- How food-web structure and biomass fluxes affect stability components?
     total biomass, species variability, statistical averaging, evenness,
     compensatory dynamics (interaction and environment)?

## What is the effects of species richness on stability facing multiple stressors in competitive and food-webs communities?

- How do species richness drive average population stability and asynchrony
  in competitive and food-webs communities ?
- How is it affected by enrichment and response diversity?

# Material & Methods

## Bioenergetic model

TO Complete

- Important bits:
  - Niche model (Connectance highly constrained by richness)
  - Type III functional response
  - No predator interference
  - Enrichment: K = 1 or 20
  - Standardized carrying capacity (Remove the effect of species richness on
    total biomass, i.e. remove the effect of niche complementarity on biomass
    produced by primary producers)

- Brose started with 5 basal species and K = 1, so his minimum total K is 5

## Variables

- S: 5 to 60
- C: .02 to .3
- K: 1 and 20
- $\rho$: 0 to 1
- Z: 5 to 1000

```{r}
tar_load(simZ)
tar_load(sim_rho)
tar_load(sim_rho_scaled)
```

```{r}
cor_enrich <- cor(simZ %>%
    filter(productivity == "enrichment") %>%
    select(c(S, richness, ct, Z, ct_alive, total_biomass, avg_max_int_alive,
        sd_max_int_alive, cv_max_int_alive,
        gini_max_int_alive,
        w_avg_tlvl_alive, max_tlvl_alive, rho, env_stoch, stab_com, avg_cv_sp,
        sync)) %>% na.omit(),
    method = "spearman")
cor_basic <- cor(simZ %>%
    filter(productivity == "basic") %>%
select(c(S, richness, ct, Z, ct_alive, total_biomass, avg_max_int_alive,
        sd_max_int_alive, cv_max_int_alive,
        gini_max_int_alive,
        w_avg_tlvl_alive, max_tlvl_alive, rho, env_stoch, stab_com, avg_cv_sp,
        sync)) %>% na.omit(),
    method = "spearman")
p_cor_basic <- ggcorrplot(cor_basic, type = "lower",
   lab = TRUE)
p_cor_enrich <- ggcorrplot(cor_enrich, type = "lower",
   lab = TRUE)
p_cor <- plot_grid(
  p_cor_basic,
  p_cor_enrich,
  ncol = 2,
  labels = "auto"
)
ggsave(here("report", "figures", "p_cor.pdf"),
  plot = p_cor,
  device = NULL,
  path = NULL,
  scale = 3.0,
  units = "mm",
  width = 140,
  height = 140 * .5,
  dpi = 300,
  limitsize = TRUE,
  bg = NULL
)
```

```{r}
cor_cap <- "Correlation among foodweb structure and stability for basic (a) and
enriched (b) environments."
```

```{r, fig.cap=cor_cap}
knitr::include_graphics(here::here("report", "figures", "p_cor.pdf"))
```

# Effect of species richness

```{r}
sim_rich_prod <- simZ %>%
  filter(
    env_stoch == .5
  )
```

```{r}
p_sim_rich_prod <- sim_rich_prod %>%
  mutate(stab_pop = 1/ avg_cv_sp) %>%
  pivot_longer(
    c(stab_com, stab_pop, sync),
    names_to = "metric",
    values_to = "stab") %>%
  mutate(
    metric = var_replacement()[metric],
    rho = rho_replacement()[as.character(rho)]
  ) %>%
  ggplot(aes(x = richness, y = stab, color = factor(productivity))) +
  geom_point(alpha = .1, shape = 20) +
  geom_smooth() +
  facet_grid(rows = vars(metric), cols = vars(rho), scales = "free_y") +
  labs(
    color = "Productivity",
    x = "Species richness",
    y = "Stability component"
  ) +
  scale_x_continuous(
    sec.axis = sec_axis(~ . ,
      name = "Response diversity",
      breaks = NULL,
      labels = NULL)
  )
ggsave(
  filename = here::here("report", "figures", "stability_rich.pdf"),
  p_sim_rich_prod,
  scale = 1.8,
  width = 88,
  height = 88 * .6,
  units = "mm"
)
```

```{r}
p_sim_rich_prod_fw_only <- sim_rich_prod %>%
  filter(ct_alive != 0, avg_max_int_alive != 0) %>%
  mutate(stab_pop = 1/ avg_cv_sp) %>%
  pivot_longer(
    c(stab_com, stab_pop, sync),
    names_to = "metric",
    values_to = "stab") %>%
  mutate(
    metric = var_replacement()[metric],
    rho = rho_replacement()[as.character(rho)]
  ) %>%
  ggplot(aes(x = log(richness), y = log(stab), color = factor(productivity))) +
  geom_point(alpha = .1, shape = 20) +
  geom_smooth(method = "lm") +
  facet_grid(rows = vars(metric), cols = vars(rho), scales = "free_y") +
  labs(
    color = "Productivity",
    x = "Species richness",
    y = "Stability component"
  ) +
  scale_x_continuous(
    sec.axis = sec_axis(~ . ,
      name = "Response diversity",
      breaks = NULL,
      labels = NULL)
  )
ggsave(
  filename = here("report", "figures", "p_sim_rich_prod_fw_only.pdf"),
  p_sim_rich_prod_fw_only,
  scale = 1.8,
  width = 88,
  height = 88 * .6,
  units = "mm"
)
```

```{r}
p_sim_rich_prod_p_only <- sim_rich_prod %>%
  filter(ct_alive == 0) %>%
  mutate(stab_pop = 1/ avg_cv_sp) %>%
  pivot_longer(
    c(stab_com, stab_pop, sync),
    names_to = "metric",
    values_to = "stab") %>%
  mutate(
    metric = var_replacement()[metric],
    rho = rho_replacement()[as.character(rho)]
  ) %>%
  ggplot(aes(x = richness, y = stab, color = factor(productivity))) +
  geom_point(alpha = .1, shape = 20) +
  geom_smooth() +
  facet_grid(rows = vars(metric), cols = vars(rho), scales = "free_y") +
  labs(
    color = "Productivity",
    x = "Species richness",
    y = "Stability component"
  ) +
  scale_x_continuous(
    sec.axis = sec_axis(~ .,
      name = "Response diversity",
      breaks = NULL,
      labels = NULL)
  )
ggsave(
  filename = here("report", "figures", "p_sim_rich_prod_p_only.pdf"),
  p_sim_rich_prod_p_only,
  scale = 1.8,
  width = 88,
  height = 88 * .6,
  units = "mm"
)
```

```{r, fig.cap = "Food-Web only."}
knitr::include_graphics(here("report", "figures", "p_sim_rich_prod_fw_only.pdf"))
```

```{r, fig.cap = "Primary producers only."}
knitr::include_graphics(here::here("report", "figures", "p_sim_rich_prod_p_only.pdf"))
```

- Enrichment decreases the effect of response diversity of stability
- Enrichment increases synchrony
- Richness decreases population stability, but increases it in absence of response
  diversity
- Food-web maintains some asynchrony despite the absence of response
  diversity but synchrony increase with richness in basic environment

```{r}
simZ %>%
  filter(rho == 1) %>%
  mutate(stab_pop = 1/ avg_cv_sp) %>%
  pivot_longer(
    c(stab_com, stab_pop, sync, total_biomass),
    names_to = "metric",
    values_to = "stab") %>%
  ggplot(aes(x = richness, y = stab, color = factor(env_stoch))) +
  geom_point() +
  facet_grid(rows = vars(metric), cols = vars(productivity), scales = "free_y") +
  labs(
    color = "Evironmental stochasticity",
    x = "Species richness",
    y = "Stability component"
  )
```

# Food-web structure and interaction strength drive stability

```{r}
compute_rotated_pca <- function(.data = NULL, naxis = 2) {
  .data <- na.omit(.data)
  pca <- ade4::dudi.pca(as.data.frame(.data),
    scannf = FALSE, nf = naxis, center = TRUE, scale = TRUE
  )
  pca_rotated <- psych::principal(.data,
    rotate = "varimax", nfactors = naxis,
    scores = TRUE
  )

  return(list(normal = pca, rotated = pca_rotated))
}
```

```{r}
sim_sem <- simZ %>%
  filter(ct_alive != 0, env_stoch == 0.5, rho == 0) %>%
  mutate(across(c(async, stab_pop, stab_com, richness, ct_alive, sd_max_int_alive, avg_max_int_alive), log, , .names = "log_{.col}")) %>%
  mutate(
    log1_avg_max_int_alive = log(avg_max_int_alive + 1),
    log1_sd_max_int_alive = log(sd_max_int_alive + 1),
    productivity_f = factor(productivity)
    ) %>%
  as.data.frame()
scaled_sim_sem <- sim_sem %>%
  mutate(across(where(is.double), ~scale(.x)[,1]))
```

```{r}
var_fw_struct <- c("richness", "ct_alive", "Z")
var_fw_emergent_struct <- c("max_tlvl_alive", "w_avg_tlvl_alive")
var_fw_flux <- c("gini_max_int_alive", "avg_max_int_alive", "sd_max_int_alive",
  "cv_max_int_alive")
var_stab <- c("async", "stab_com", "stab_pop")
```

```{r}
#colnames(simZ)
pairs(sim_sem %>%
  filter(env_stoch == .5) %>%
  select(c(var_fw_struct, var_fw_emergent_struct, var_fw_flux, stab_pop)))
```

```{r, eval = FALSE}
performance::check_collinearity(lm(w_avg_tlvl_alive ~ max_tlvl_alive,sim_sem))
performance::check_collinearity(lm(w_avg_tlvl_alive ~ ct_alive + richness,sim_sem))
performance::check_collinearity(lm(w_avg_tlvl_alive ~ ct_alive + richness,sim_sem))
performance::check_collinearity(lm(avg_max_int_alive ~ w_avg_tlvl_alive + ct_alive + richness,sim_sem))
performance::check_collinearity(lm(avg_max_int_alive ~ max_tlvl_alive + ct_alive,sim_sem))
performance::check_collinearity(lm(stab_com ~ log_richness + log1_avg_max_int_alive + log1_sd_max_int_alive + w_avg_tlvl_alive + ct_alive,sim_sem))
performance::check_collinearity(lm(stab_com ~ richness + avg_max_int_alive + sd_max_int_alive + w_avg_tlvl_alive + ct_alive,sim_sem))
```

```{r}
sem_list_basic <- list(
  w_avg_tlvl_alive = lm(w_avg_tlvl_alive ~ richness + ct_alive, sim_sem),
  sd_max_int_alive = lm(sd_max_int_alive ~ richness + ct_alive + w_avg_tlvl_alive, sim_sem),
  avg_max_int_alive = lm(avg_max_int_alive ~ richness + ct_alive + w_avg_tlvl_alive, sim_sem),
  log_async = lm(log_async ~ richness + avg_max_int_alive + sd_max_int_alive + w_avg_tlvl_alive + ct_alive, sim_sem),
  log_stab_pop = lm(log_stab_pop ~ richness + avg_max_int_alive + sd_max_int_alive + w_avg_tlvl_alive + ct_alive, sim_sem),
  log_stab_com = lm(log_stab_com ~ log_async + log_stab_pop, sim_sem)
)
sem_basic <- as.psem(sem_list_basic)
```

```{r}
ti <- summary(sem_basic)$coefficients[, c(1, 2, 8, 9)]
colnames(ti)[4] <- "Signif"
ti %>%
  mutate(
    Response = var_replacement()[Response],
    Predictor = var_replacement()[Predictor]
  ) %>%
  kable(caption = "Standardized estimates for the Structural Equation Models.") %>%
  collapse_rows(1, valign = "middle") %>%
  kable_styling()
```

```{r, eval = FALSE}
ti <- map_dfr(setNames(sem_list_basic, names(sem_list_basic)),
  performance::check_collinearity,
  .id = "response") %>%
  as_tibble
ti %>%
  mutate(
    response = var_replacement()[response],
    Term = var_replacement()[Term]
  ) %>%
  kable(caption = "Variance Inflation Factors for the Structural Equation Model.") %>%
  kable_styling()
```

```{r, eval = FALSE}
map_dfr(sem_list_basic, rsquared) %>%
  mutate(Response = var_replacement()[Response]) %>%
  select(Response, R.squared) %>%
  kable(caption = "Rsquared for the Structural Equation Model.") %>%
  kable_styling()
```

```{r}
sim_sem_nodivresp <- simZ %>%
  filter(ct_alive != 0, env_stoch == 0.5, rho == 1) %>%
  mutate(across(c(async, stab_pop, stab_com, richness, ct_alive, sd_max_int_alive, avg_max_int_alive), log, , .names = "log_{.col}")) %>%
  mutate(
    log1_avg_max_int_alive = log(avg_max_int_alive + 1),
    log1_sd_max_int_alive = log(sd_max_int_alive + 1),
    productivity_f = factor(productivity)
    ) %>%
  as.data.frame()
```

```{r}
sem_list_nodivresp <- list(
  w_avg_tlvl_alive = lm(w_avg_tlvl_alive ~ richness + ct_alive, sim_sem_nodivresp),
  sd_max_int_alive = lm(sd_max_int_alive ~ richness + ct_alive + w_avg_tlvl_alive, sim_sem_nodivresp),
  avg_max_int_alive = lm(avg_max_int_alive ~ richness + ct_alive + w_avg_tlvl_alive, sim_sem_nodivresp),
  log_async = lm(log_async ~ richness + avg_max_int_alive + sd_max_int_alive + w_avg_tlvl_alive + ct_alive, sim_sem_nodivresp),
  log_stab_pop = lm(log_stab_pop ~ richness + avg_max_int_alive + sd_max_int_alive + w_avg_tlvl_alive + ct_alive, sim_sem_nodivresp),
  log_stab_com = lm(log_stab_com ~ log_async + log_stab_pop, sim_sem_nodivresp)
)
sem_nodivresp <- as.psem(sem_list_nodivresp)
```

```{r}
coef_sem_nodivresp <- summary(sem_nodivresp)$coefficients[, c(1, 2, 8, 9)]
colnames(coef_sem_nodivresp)[4] <- "Signif"
coef_sem_nodivresp %>%
  mutate(
    Response = var_replacement()[Response],
    Predictor = var_replacement()[Predictor]
  ) %>%
  kable(caption = "Standardized estimates for the Structural Equation Models with no response diversity.") %>%
  collapse_rows(1, valign = "middle") %>%
  kable_styling()
```

# Effect of multiple stressors on stability in monotrophic and food-webs

```{r}
tar_load(p_mod_rho_scaled)
p_mod_rho_scaled_tot <- plot_grid(plotlist = p_mod_rho_scaled)
ggsave(here("report", "figures", "p_mod_rho_scaled_tot.pdf"),
  plot = p_mod_rho_scaled_tot,
  device = NULL,
  path = NULL,
  scale = 2.4,
  units = "mm",
  width = 140,
  height = 140 * .5,
  dpi = 300,
  limitsize = TRUE,
  bg = NULL
)
```

```{r, eval = FALSE}
tar_load(p_mod_rho)
plot_grid(plotlist = p_mod_rho)
```

```{r, eval = FALSE}
sim_rho
```

```{r}
tar_load(pred_rho)
p_pred_rho <- pred_rho %>%
  mutate(rho = rho_replacement()[as.character(rho)]) %>%
  ggplot(aes(
      y = log_stab_com,
      x = log_richness,
      color = factor(ct_alive),
      linewidth = factor(log1_avg_max_int_alive)
      )) +
  geom_line() +
  labs(
    color = "Connectance",
    linewidth = "Avg interaction strength",
    y = "Community stability",
    x = "Species richness (log)"
    ) +
  facet_grid(
    rows = vars(rho),
    cols = vars(productivity)
  )
ggsave(here("report", "figures", "p_pred_rho.pdf"),
  plot = p_pred_rho,
  device = NULL,
  path = NULL,
  scale = 2.4,
  units = "mm",
  width = 140,
  height = 140 * .5,
  dpi = 300,
  limitsize = TRUE,
  bg = NULL
)
```

```{r}
predS_cap <- "Predictions for the model of drivers of community stability in
basic and enriched environments (columns) and across varying response diversity
(rows)."
```

```{r predS, fig.cap = predS_cap}
knitr::include_graphics(here("report", "figures", "p_pred_rho.pdf"))
```

```{r}
coef_rho_cap <- "Coefficients for the model of drivers of community stability,
population stability and asynchrony. See predictions for a more simple overview."
```

```{r coef-rho, fig.cap = coef_rho_cap}
knitr::include_graphics(here("report", "figures", "p_mod_rho_scaled_tot.pdf"))
```

- From the predictions of the model:
  - Stong interaction among interaction strength, connectance and species
    richness, enrichment and response diversity = *Food-web structure effects on
    stability is context dependant*
  - High connectance and high interaction strength is generally destabilizing,
    driving negative relationships between stability and species richness in
    enrichment context (high $K$)
  - At the reverse, low connectance and high interaction strength is stabilizing
  - In basic environment, high interaction strength stabilise the community
    biomass in the absence of response diversity. At the opposite, low
    interaction strength generate a negative stability richness relationship.
  - In enrichment context, response diversity has little effect on stability!
  - In enrichment context, High interaction strength is generally stabilizing!

```{r, eval = FALSE}
mod_rho_tot <- glmmTMB(
  log_stab_com ~ log1_avg_max_int_alive*log_richness*ct_alive*rho*productivity +
    env_stoch + (1 | Z),
  sim_rho_scaled)
coef_mod_rho_tot <- model2tibble(mod_rho_tot)
coef_mod_rho_tot %>%
  ggplot(aes(y = parameter, x = coefficient, xmin = ci_low, xmax = ci_high)) +
  geom_pointrange() +
  labs(x = "Standardized Coefficients")
```

```{r, eval = FALSE}
sim_rho_pers <- simZ %>%
  mutate(across(c(S, ct, avg_max_int, gini_max_int), log, , .names = "log_{.col}")) %>%
  mutate(
    log1_avg_max_int = log(avg_max_int + 1),
    productivity_f = factor(productivity)
    ) %>%
  as.data.frame() %>%
  select(c(
      log_S, ct, max_tlvl, Z, rho,
      gini_max_int, avg_max_int, log1_avg_max_int, env_stoch,
      productivity, productivity_f,
      persistence)
  )
sim_rho_pers_scaled <- sim_rho_pers %>%
  mutate(across(where(is.double), ~scale(.x)[,1]))

pred_rho_pers <- "log1_avg_max_int * log_S * rho +
log1_avg_max_int * log_S * productivity +
log1_avg_max_int * log_S * ct +
log1_avg_max_int * ct * rho +
log1_avg_max_int * ct * rho +
log1_avg_max_int * ct * productivity +
log_S * rho * productivity +
log_S * ct * productivity +
log_S * ct * rho +
env_stoch"
rand <- "(1 | Z)"

mod_rho_pers <- glmmTMB(formula(paste0("persistence ~ ", pred_rho_pers, " + ", rand)),
  sim_rho_pers_scaled)

r2_nakagawa(mod_rho_pers)
plot(parameters(mod_rho_pers))
```

## Regression tree

```{r, eval = FALSE}
library(rpart)
library(rpart.plot)
set.seed(123)
m1 <- rpart(
  formula = log_stab_com ~ .,
  data    = sim_rho %>%
    select(log_stab_com, log_richness, ct_alive, rho, env_stoch,
      gini_max_int, max_tlvl_alive, log1_avg_max_int_alive, productivity),
  method  = "anova"
  )
rpart.plot(m1)
```

```{r, eval=FALSE}
bagged_m1 <- bagging(
  formula = log_stab_com ~ .,
  data    = sim_rho %>%
    select(log_stab_com, log_richness, ct_alive, rho, env_stoch,
      gini_max_int, max_tlvl_alive, log1_avg_max_int_alive, productivity),
  coob    = TRUE
)
```

```{r, eval = FALSE}
library(ipred)
ctrl <- trainControl(method = "cv",  number = 10)

# CV bagged model
bagged_cv <- train(
  Sale_Price ~ .,
  data = ames_train,
  method = "treebag",
  trControl = ctrl,
  importance = TRUE
  )
```

```{r, fig.cap = "Variable importance from bagged regression tree."}
ctrl <- trainControl(method = "cv",  number = 10)

# CV bagged model
bagged_cv <- train(
  stab_com ~ richness + ct_alive + max_tlvl_alive +
    w_avg_tlvl_alive + avg_max_int_alive + sd_max_int_alive + rho + productivity,
  data = simZ %>% filter(ct_alive != 0, avg_max_int_alive != 0),
  cp = .2,
  maxdepth = 6,
  method = "treebag",
  trControl = ctrl,
  importance = TRUE
  )
plot(varImp(bagged_cv), 8)
```

```{r, fig.cap="Regression tree "}
m1 <- rpart(
  formula = stab_com ~ richness + ct_alive + rho + productivity + w_avg_tlvl_alive + max_tlvl_alive +
    avg_max_int_alive + sd_max_int_alive,
  data    = simZ %>%
    filter(ct_alive != 0, avg_max_int_alive != 0),
  method  = "anova"
  )
rpart.plot(m1)
```

# Appendix

```{r}
sim_ok %>%
  ggplot(aes(x = as.factor(Z), y = max_tlvl_alive, color = as.factor(S))) +
  geom_boxplot() +
  facet_grid(cols = vars(productivity))
sim_ok %>%
  ggplot(aes(x = as.factor(Z), y = max_tlvl_alive, color = as.factor(S))) +
  geom_boxplot() +
  facet_grid(cols = vars(env_stoch), rows = vars(productivity))
sim_ok %>%
  ggplot(aes(x = as.factor(Z), y = avg_max_int_alive, color = as.factor(S))) +
  geom_boxplot() +
  facet_grid(cols = vars(productivity))

sim_ok %>%
  ggplot(aes(x = as.factor(Z), y = persistence, color = as.factor(S))) +
  geom_boxplot() +
  facet_grid(cols = vars(env_stoch), rows = vars(productivity))
sim_ok %>%
  ggplot(aes(x = as.factor(Z), y = pop_stab, color = as.factor(S))) +
  geom_boxplot() +
  facet_grid(cols = vars(env_stoch), rows = vars(productivity))
sim_ok %>%
  ggplot(aes(x = ct_alive, y = persistence, color = as.factor(S))) +
  geom_point()
```

- K = 1 is too low for meaningful food-webs, I should start with K = 5 as in
  Brose et al. (2006)