### 2023-05-19 use_case/cpe_decomposition.Rmd

---
title: "Test CPE decomposition"
author: Alain Danet
date: \today
output:
  bookdown::pdf_document2:
    fig_caption: true
    keep_tex: true
    toc: true
    fontsize: 12pt
    geometry: margin=2cm
    number_sections: true
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = TRUE,
  fig.path="figures/"
)
```

```{r, results = FALSE}
library(tidyverse)
library(magrittr)
library(cowplot)
library(here)
library(arrow)

source(here::here("R", "stability_decomposition.R"))
source(here::here("R", "cross_correlated_error.R"))

mytheme <- theme_cowplot(font_size = 12) +
  background_grid() +
  theme(legend.position = "bottom")
theme_set(mytheme)
options(knitr.table.format = "latex")
```

# Toy examples

```{r}
# No interaction, common environment response, CPE inf to 1
mat <- matrix(
  c(1, 1,
    0, 0,
    1, 1,
    0, 0), byrow = T, ncol = 2)
compensatory_effect(mat)
```

```{r}
# No interaction, opposite environment response, CPE sup to 1
mat <- matrix(
  c(1, 0,
    1, 1,
    1, 0,
    0, 1), byrow = T, ncol = 2)
compensatory_effect(mat)
```

# In simulation without interactions

```{r}
# Common environmental response + no interaction
ti <- cross_correlated_response(
  rho = 1,
  evt_sd = .6, n = 10, time_len = 1000,
  return_tbl = TRUE) %>%
mutate(
  bm_brut = rnorm(length(error), mean = 100, sd = 1),
  stoch_d = .1 * bm_brut * exp(error),
  bm = bm_brut - stoch_d
)
mat <- ti %>%
  pivot_wider(c(time), names_from = "species", values_from = "bm") %>%
  select(-time) %>%
  as.matrix()

mat_stoch <- ti %>%
  pivot_wider(c(time), names_from = "species", values_from = "stoch_d") %>%
  select(-time) %>%
  as.matrix()
# Due to interaction, should be 1
compensatory_effect(mat) / compensatory_effect(mat_stoch)
# Due to common environmental response, should be less than 1, destabilizing:
compensatory_effect(mat_stoch)
```

```{r}
# Independent fluctuations + no interaction
ti <- cross_correlated_response(
  rho = 0,
  evt_sd = .6, n = 10, time_len = 1000,
  return_tbl = TRUE) %>%
mutate(
  bm_brut = rnorm(length(error), mean = 100, sd = 1),
  stoch_d = .1 * bm_brut * exp(error),
  bm = bm_brut - stoch_d
)
mat <- ti %>%
  pivot_wider(c(time), names_from = "species", values_from = "bm") %>%
  select(-time) %>%
  as.matrix()

mat_stoch <- ti %>%
  pivot_wider(c(time), names_from = "species", values_from = "stoch_d") %>%
  select(-time) %>%
  as.matrix()
# Due to interaction, should be 1
compensatory_effect(mat) / compensatory_effect(mat_stoch)
# Due to Independent environmental response, should be 1, i.e:
compensatory_effect(mat_stoch)
```

```{r}
data_raw <- open_dataset(here::here("res", "simCSZenrich2", "sim_toy_compensatory_dyn.arrow"),
          format = "arrow") %>%
collect() %>%
filter(!is.na(richness)) %>%
mutate(
  stoch_d = map2(species, stoch, function(bm, d_stoch) {
    .1 * bm * exp(d_stoch)
    }),
  species = map(species, ~matrix(.x, nrow = 100)),
  stoch = map(stoch, ~matrix(.x, nrow = 100)),
  stoch_d = map(stoch_d, ~matrix(.x, nrow = 100)),
  cpe = map_dbl(species, compensatory_effect),
  cpe_env = map_dbl(stoch_d, compensatory_effect),
  cpe_int = cpe / cpe_env
)
```

```{r}
data_raw %>%
  ggplot(aes(x = as.factor(rho), y = cpe_env)) +
  geom_boxplot() +
  geom_hline(yintercept = 1) +
  geom_hline(yintercept = .5, linetype = "dashed")

data_raw %>%
  ggplot(aes(x = as.factor(rho), y = cpe)) +
  geom_boxplot() +
  geom_hline(yintercept = 1) +
  geom_hline(yintercept = 2, linetype = "dotted") +
  geom_hline(yintercept = .5, linetype = "dashed")

data_raw %>%
  ggplot(aes(x = as.factor(rho), y = cpe_int)) +
  geom_boxplot() +
  geom_hline(yintercept = 1) +
  geom_hline(yintercept = 2, linetype = "dashed")
```