### 2023-11-24 use_case/cpe_decomposition.Rmd

---
title: "Test CPE decomposition"
author: Alain Danet
date: \today
output:
  bookdown::pdf_document2:
    fig_caption: true
    keep_tex: true
    toc: true
    fontsize: 12pt
    geometry: margin=2cm
    number_sections: true
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  #fig.dim = c(7, 7),
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = TRUE,
  fig.path="figures/"
)
```

```{r, results = FALSE}
library(tidyverse)
library(magrittr)
library(cowplot)
library(here)
library(arrow)

source(here::here("R", "stability_decomposition.R"))
source(here::here("R", "cross_correlated_error.R"))

mytheme <- theme_cowplot(font_size = 12) +
  background_grid() +
  theme(legend.position = "bottom")
theme_set(mytheme)
options(knitr.table.format = "latex")
```

# Toy examples

```{r}
# No interaction, common environment response, CPE inf to 1
mat <- matrix(
  c(1, 1,
    0, 0,
    1, 1,
    0, 0), byrow = T, ncol = 2)
compensatory_effect(mat)
```

```{r}
# No interaction, opposite environment response, CPE sup to 1
mat <- matrix(
  c(1, 0,
    1, 1,
    1, 0,
    0, 1), byrow = T, ncol = 2)
compensatory_effect(mat)
```

# In simulation without interactions

```{r}
# Common environmental response + no interaction
ti <- cross_correlated_response(
  rho = 1,
  evt_sd = .6, n = 10, time_len = 1000,
  return_tbl = TRUE) %>%
mutate(
  bm_brut = rnorm(length(error), mean = 100, sd = 1),
  stoch_d = .1 * bm_brut * exp(error),
  bm = bm_brut - stoch_d
)
mat <- ti %>%
  pivot_wider(c(time), names_from = "species", values_from = "bm") %>%
  select(-time) %>%
  as.matrix()

mat_stoch <- ti %>%
  pivot_wider(c(time), names_from = "species", values_from = "stoch_d") %>%
  select(-time) %>%
  as.matrix()
# Due to interaction, should be 1
compensatory_effect(mat) / compensatory_effect(mat_stoch)
# Due to common environmental response, should be less than 1, destabilizing:
compensatory_effect(mat_stoch)
```

```{r}
# Independent fluctuations + no interaction
ti <- cross_correlated_response(
  rho = 0,
  evt_sd = .6, n = 10, time_len = 1000,
  return_tbl = TRUE) %>%
mutate(
  bm_brut = rnorm(length(error), mean = 100, sd = 1),
  stoch_d = .1 * bm_brut * exp(error),
  bm = bm_brut - stoch_d
)
mat <- ti %>%
  pivot_wider(c(time), names_from = "species", values_from = "bm") %>%
  select(-time) %>%
  as.matrix()

mat_stoch <- ti %>%
  pivot_wider(c(time), names_from = "species", values_from = "stoch_d") %>%
  select(-time) %>%
  as.matrix()
# Due to interaction, should be 1
compensatory_effect(mat) / compensatory_effect(mat_stoch)
# Due to Independent environmental response, should be 1, i.e:
compensatory_effect(mat_stoch)
```

```{r}
data_raw <- open_dataset(
  here::here("use_case", "sim_toy_compensatory_dyn.arrow"),
  format = "arrow") %>%
collect() %>%
filter(!is.na(richness)) %>%
mutate(
  stoch_d = map2(species, stoch, function(bm, d_stoch) {
    .1 * bm * exp(d_stoch)
    }),
  species = map(species, ~matrix(.x, nrow = 100)),
  stoch = map(stoch, ~matrix(.x, nrow = 100)),
  stoch_d = map(stoch_d, ~matrix(.x, nrow = 100)),
  cpe = map_dbl(species, compensatory_effect),
  cpe_env = map_dbl(stoch_d, compensatory_effect),
  cpe_int = cpe / cpe_env
)
```

```{r}
data_raw %>%
  ggplot(aes(x = as.factor(rho), y = cpe_env)) +
  geom_boxplot() +
  geom_hline(yintercept = 1) +
  geom_hline(yintercept = .5, linetype = "dashed")

data_raw %>%
  ggplot(aes(x = as.factor(rho), y = cpe)) +
  geom_boxplot() +
  geom_hline(yintercept = 1) +
  geom_hline(yintercept = 2, linetype = "dotted") +
  geom_hline(yintercept = .5, linetype = "dashed")

data_raw %>%
  ggplot(aes(x = as.factor(rho), y = cpe_int)) +
  geom_boxplot() +
  geom_hline(yintercept = 1) +
  geom_hline(yintercept = 2, linetype = "dashed")
```

```{r}
source(here::here("R", "plotting_methods.R"))
```

```{r}
ti <- data_raw %>%
  filter(rho == 0, Z == 10, env_stoch == .6)
nice_ts <- ti$species[[2]] %>%
  .[, colSums(.) > 0] %>%
  as_tibble() %>%
  mutate(time = 1:n()) %>%
  pivot_longer(-time, names_to = "species", values_to = "biomass")
p_nice_ts_sp <- nice_ts %>%
  ggplot(aes(y = biomass, x = time, color = species)) +
  geom_line() +
  labs(y = "Biomass", x = "Time") +
  theme_half_open() +
  theme(
    legend.position = "none",
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
ggsave_multiple(
  c("p_nice_ts_sp.png", "p_nice_ts_sp.pdf"),
  plot = p_nice_ts_sp,
  path = here::here("use_case", "figures"),
  scale = 1.5,
  dpi = 600,
  width = 88,
  height = 88 * .5,
  units = "mm"
)

sum_nice_ts <- nice_ts %>%
  group_by(time) %>%
  summarise(biomass = sum(biomass))

p_sum_nice_ts <- sum_nice_ts %>%
  ggplot(aes(y = biomass, x = time)) +
  geom_line(color = "black") +
  labs(y = "Biomass", x = "Time") +
  theme_half_open() +
  theme(
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )
ggsave_multiple(
  c("p_sum_nice_ts.png", "p_sum_nice_ts.pdf"),
  plot = p_sum_nice_ts,
  path = here::here("use_case", "figures"),
  scale = 1.5,
  dpi = 600,
  width = 88,
  height = 88 * .5,
  units = "mm"
)

p_hist_biomass <- sum_nice_ts %>%
  ggplot(aes(x = biomass)) +
  geom_histogram(bins = 50, fill = "black") +
  labs(x = "Biomass", y = "") +
  theme_void() +
  theme(
    axis.ticks = element_blank(),
    axis.text = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank()
  )

ggsave_multiple(
  c("p_hist_biomass.png", "p_hist_biomass.pdf"),
  plot = p_hist_biomass,
  path = here::here("use_case", "figures"),
  scale = 1,
  dpi = 600,
  width = 88,
  height = 88 * 1,
  units = "mm"
)
```