### 2024-01-15 paper/manuscript.Rmd
Cleaning for sending to coauthors

---
title: Food-web structure, stochasticity and response diversity drive temporal community stability in complex communities
author: "Alain Danet, (Sonia Kéfi, Thomas F Johnson), Andrew P. Beckerman"
date: ""
output:
  officedown::rdocx_document:
    number_sections: false
  bookdown::pdf_document2:
    latex_engine: xelatex
    keep_tex: true
    spacing: double
    number_sections: false
always_allow_html: true
bibliography: references.bib
link-citations: true
csl: ecology-letters.csl
toc: false
linestretch: 2.0
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{hyperref}
  - \hypersetup{backref=true}
  - \usepackage{setspace}\doublespacing
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.pos = "H",  # pdf mode
  #fig.dim = c(7, 7),
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = TRUE,
  fig.path="figures/"
)
```

```{r, results = FALSE}
sapply(c("targets", "tidyverse", "magrittr", "cowplot", "here", "piecewiseSEM",
    "semEff", "ggcorrplot", "kableExtra", "rmarkdown", "officedown"),
  require,
  character.only = TRUE)
mytheme <- theme_cowplot(font_size = 8) +
background_grid() +
theme(legend.position = "bottom")
theme_set(mytheme)
sapply(list.files(here("R"), full.names = TRUE), source)
```

---
## IDEA, meeting with Sonia
#- Simplifying SEM in main text, move details to appendix
#- State diagram for stability - richness relationships along two food-web
#  metrics and panels varying response diversity.
#- Look at the french data for portefolio and compensatory dynamics effects.
#- Mid-October
---

# Abstract

Current biotic homogenization and increasing frequency of extreme events are
projected to dramatically destabilize ecological communities. Understanding how
complex communities such as food-webs respond to such threats is crucial to take
adequate conservation measures. We then investigated how response diversity,
environmental stochasticity, and food-web structure drive temporal stability of
biomass in complex food-webs, using a stochastic bioenergetic food-web model. We
found that higher average trophic level, omnivory degree and generalism
increased  average population stability while species richness had little
effects. At the opposite higher average trophic level, omnivory degree and
generalism decreased asynchrony and compensatory dynamics raising trophic
interactions. We further found that that high interaction strength drive
negative stability-species richness when response diversity is low but that high
response diversity always drive positive stability-richness relationship. Our
study reveals that response diversity and food-web structure drive temporal
stability and further improve our understanding of biodiversity-stability
relationships in complex trophic communities.

# Introduction

Global environmental change constitutes a suite of major threats to biodiversity
and ecosystem functioning [@lee_ipcc_2023; @simmons_refocusing_2021]. A central
feature of these threats is that they are predicted to dramatically increase
variability in the environment that species face [@lee_ipcc_2023] with potential
impacts on the stability of communities and associated impacts on biodiversity
and functioning in these communities. Assessing the impacts of this variability
on communities is challenging for several reasons.

First, the stability of communities is often evaluated using temporal
variability the metric of interest (e.g. abundance or biomass), because most
natural communities face persistent perturbations so that other stability
metrics such as resilience and resistance are not always applicable
[@arnoldi_resilience_2016]. Understanding temporal community variability,
measured as the inverse of the coefficient of variation
[@tilman_biodiversity_2006; @roscher_identifying_2011;
@thibaut_understanding_2013; @zhao_biodiversity_2022], requires scaling down
the focus to populations. Community stability can be mathematically partitioned
into average population stability and asynchrony [@loreau_species_2008;
@thibaut_understanding_2013; @zhao_biodiversity_2022], asynchrony measuring how
much each species fluctuations are compensating each other at the community
level. Population stability and asynchrony depend both on the number of species
and their dominance [@loreau_species_2008; @thibaut_understanding_2013;
@zhao_biodiversity_2022], the interactions among them [@thebault_trophic_2005;
@thebault_relationship_2006; @loreau_species_2008; @danet_species_2021;
@eschenbrenner_diversity_2023] and on the way that species respond to
environmental perturbations [response diversity, @ives_stability_2000;
@elmqvist_response_2003; @ross_how_2023]. Therefore, accounting for the effects
on community structure on population stability and asynchrony is crucial to
improve our understanding of community stability in face of environmental
change.

Average population stability is decreased by external forcing, such as the
intensity and the frequency of perturbations [@loreau_species_2008]. Species richness can
decrease [@tilman_biodiversity_2006; @loreau_species_2008], increase
[@houlahan_negative_2018], or have little effects on population stability. This
effect of species richness on population stability depends on the evenness of
the community such as dominant species contribute more the average of the
community than subordinate species. At the core of ecological theory,
population stability depends on internal forcing, here species interactions. The
topological structure and the distribution of interaction strength have tremendous
effects on population stability, such as strong interaction strength are
generally destabilising [@mccann_weak_1998; @mccann_reevaluating_1997;
@thebault_trophic_2005]. In particular for food-webs, top-down and bottom-up
controls can increase population stability.

Asynchrony in turn can be partitioned into portfolio effects and compensatory
dynamics and is mostly determined by internal factors. Portfolio effects stems
from the statistical averaging of independent population fluctuations
[@doak_statistical_1998; @zhao_biodiversity_2022], thereby increasing
asynchrony. Compensatory dynamics in turn captures population fluctuations that
are more synchronous or asynchronous than expected for independent fluctuations,
and can stem from species interactions and species responses to perturbations
[@zhao_biodiversity_2022]. Response diversity captures how among species’
differences in their response to the environmental variation, can underpin
asynchronous dynamics if their responses are different but synchronous if their
responses are similar. Theory and empirical data emerging around response
diversity suggests it is  “core ecological mechanism underpinning many of the
statistical explanations of the diversity–stability relationships”
[@ives_stability_2000; @elmqvist_response_2003; @ross_how_2023]. In turn,
the distribution and strength of species interactions can also modulate
asynchrony through compensatory dynamics such as a strong asymmetry of consumer
preference can generate asynchrony in resource fluctuations [@mccann_weak_1998;
@vasseur_environmental_2007], and a generalist consumer can in turn generate
synchrony among its prey [@raimondo_interspecific_2004]. Competition can also
have synchronising effects in strong species growth rates regime
[@loreau_species_2008]. Finally, portfolio and compensatory effects can also be
affected by the population evenness such as fluctuations of a dominant
species will hardly be compensated by a subordinate species. Improving our
understanding of community stability facing perturbations therefore requires to
articulate the effect response diversity, species interactions, species
richness and evenness on the different components of stability.

The stability of ecological communities facing perturbations stems from two
parallel bodies of work: species interactions and response diversity. These two
bodies of work share the common thread of community structure acting through a
filter to drive stability. However, the filter is different in each space; in
the first, the filter is the topology and the strength of species interactions
and in the second variation in species’ responses. Moreover, the theory has been
developed largely in theoretical or empirical communities characterised by a
limited array of interactions (e.g. often lacking in consumers) and by limited
species diversity/richness. Thus, ecology lacks a body of theory integrating
across structure and response diversity frameworks and reflecting species rich
communities. Here, we integrate these two perspectives on diversity - stability
to evaluate the combined effects of community structure and response diversity on
temporal stability in species rich, complex communities characterised by
competition and predation.

The ability to do this is supported by a key statistical advance: it is now
possible to partition asynchrony into two component sources, the statistical
averaging effect (SAE) and compensatory dynamics effect (CPE)
[@zhao_biodiversity_2022]. SAE quantifies the portfolio effect emerging from
independent species fluctuations, while CPE describes compensatory dynamics
emerging from species interactions and differential responses to environmental
stochasticity. This allows an unprecedented opportunity to obtain mechanistic
insight about the effects of stressors and food-web structure on stability in
complex, species rich communities.

Our primary objective is to derive theory that predicts how the components of
community stability, that can differentially drive patterns of stability, deliver novel
insights into the source and patterns of diversity-stability relationships in
complex communities facing environmental stochasticity. Our approach delivers
revised insight into diversity-productivity relationships in species rich
communities when both community structure per se, and species
responses can vary. In the following sections, we begin by introducing the
multi-species model and the features of the model that allow us to integrate
across variation in community structure and response diversity. We
then introduce assumptions we used to implement environmental stochasticity,
response diversity, variation in community structure and variation in the
distribution of interaction types and strengths. Next we define our in-silico
experimental design that specifically aligns with three core questions and
associated analyses: (1) What are the direct effects of food-web structure and
interaction strength on population stability, compensatory dynamics,
and statistical averaging effects (i.e. portfolio effects)? (2) What are the
total effects of food-web structure, response diversity, environmental
stochasticity on community stability. Finally, merging the two parallel bodies of
work, we asked (3) How do response diversity and food-web structure interact to
drive the sign of the relation between community stability and
species richness?

---
#Here we deliver inference from modelling of complex biodiverse communities on
#how food-web structure (connectance, average interaction strength, average
#trophic level), response diversity, and environmental stochasticity combine to
#shape temporal stability of community biomass, as well as its partition into
#asynchrony and population stability. We then assess how food-web complexity and
#response diversity shape stability-richness relationships.

#Environmental forcing (such as food availability and
#warming) and land-use intensity was reported to synchronise population
#fluctuations [@benton_population_2001; @bluthgen_land_2016; @song_reduced_2015;
#@liu_aridity_2023] while habitat degradation tends to decrease population
#stability [@olivier_urbanization_2020]. Understanding the effects of global
#change on community temporal stability therefore require to account for their
#simultaneous effects on community and population scales [@kefi_advancing_2019;
#@simmons_refocusing_2021].

#Increasing species richness has been reported to decrease population stability
#in experimental grasslands [@tilman_biodiversity_2006;
#@roscher_identifying_2011], while increasing population stability in natural
#settings [@houlahan_negative_2018] [add mechanisms?]. In turn, higher species
#richness often results in higher asynchrony, through portfolio effects (i.e.
#statistical averaging effects due to independent species fluctuations) and
#compensatory dynamics (due to species interactions and response diversity of
#species due to the environment) [@loreau_species_2008;
#@thibaut_understanding_2013]. Systematic increase in asynchrony with species
#richness is the mechanism that drives reported positive relationships between
#community stability and species richness [ref]. Anthropogenic pressures, such as
#eutrophication, has been reported to weakens the relationship between species
#richness and asynchrony in grasslands [@hautier_eutrophication_2014] but such
#empirical evidence remains scarce. However, stressors such as biotic
#homogenization [@buisson_toward_2013; @muthukrishnan_invasive_2020;
#@saladin_rapid_2020] and environmental forcing that decrease species response
#diversity should dampens the relationships between asynchrony and species
#richness [@ives_stability_2000; @thibaut_understanding_2013] due to a decrease
#in portfolio effects and compensatory effects. [Talk about the effects of
#evenness in temporal variability and evenness in abundance? e.g.,
#@zhao_biodiversity_2022; @thibaut_understanding_2013]. Because stressors can
#affect both community stability per se and how community stability scales with
#species richness, improving our understanding require to study both
#[@ives_stability_2007; @hautier_anthropogenic_2015].

#including levels of generalism, the distribution of interaction strengths and
#the prevalence of omnivory, can also influence stability. The degree of
#generalism of predator was shown to modulate stability-richness relationships
#[@thebault_trophic_2005], and generalist predators can synchronise the dynamics
#of their preys[@raimondo_interspecific_2004]. The strength and distribution of
#interaction strength was also shown to be strong driver of community stability,
#such as stronger interaction strengths can be destabilizing [@may_will_1972;
#@mccann_weak_1998]. Omnivory was also shown to drive higher population stability
#[@mccann_reevaluating_1997]. The effects of food-web structure on ecosystem
#stability has been mostly assessed with small modules [but see
#@eschenbrenner_diversity_nodate], and we know very little how food-web structure
#affects the relationships between species richness and temporal stability in
#complex food-webs. So understanding the effects of stressors on community
#stability and on community stability-richness relationships necessitate to
#account for food-web structure.
---

# Methods

## Bioenergenetic model

We simulated the dynamics of complex food-webs using the allometric bioenergetic
model [@yodzis_body_1992; @brose_allometric_2006; @delmas_simulations_2017].
This model is widely used to simulate the dynamics of complex ecological
communities because of the simplicity of its parametrisation and that its
parameters are rooted in the metabolic theory of ecology [@brown_toward_2004].
The model describes species biomass dynamics over time with two master equations (eq. \@ref(eq:dbdt)), one for the
primary producers (eq. \@ref(eq:prod)) and one for the consumers (eq. \@ref(eq:prod)):

\begin{subequations}
    \label{eq:dbdt}

    \begin{equation}
        \label{eq:cons}
        \frac{dB_i}{dt} =
        \sum_{j \in \{\mathrm{res}\}_i} x_i y_i B_k F_{ij}
        - \sum_{j \in \{\mathrm{cons}\}_i} \frac{x_j y_j B_j F_{ji}}{e_{ij}}
        - x_i B_i - d_i B_i
    \end{equation}

    \begin{equation}
    	\label{eq:prod}
    	\frac{dB_i}{dt} =
	r_i B_i G_i
	- \sum_{j \in \{\mathrm{cons}\}_i} \frac{x_j y_j B_j F_{ji}}{e_{ij}}
        - d_i B_i
    \end{equation}
\end{subequations}

The consumers gain biomass by consuming resources (i.e. their preys) at a rate
that depends on their mass specific metabolic rate ($x_i$),  maximum consumption
rate ($y_i$) and on the functional response ($F_ij$) describing how the rate of
consumption of a consumer $i$ on a resource $j$ vary with the biomass of this resource.
Consumers then loses biomass by being consumed, $e_ij$ being the assimilation
efficiency of the consumer $j$ on the resource $i$. Consumers also lose biomass
over time through metabolic losses ($-x_iB_i$) and by natural mortality due to
environment ($-d_iB_i$). The primary producers gain biomass over time with a
growth rate ($r_i$, eq. \@ref(eq:prod)) and a functional response ($G_i$)
describing how the growth of the producers vary with their biomass. They lose
biomass by being consumed by consumers and by natural mortality.

\begin{subequations}
    \label{eq:func}
    \begin{equation}
        \label{eq:growth}
        G_i = (1 - \dfrac{\sum_{j \in \{\mathrm{prod}\}} \alpha_{ij} B_j}{K_i})
    \end{equation}

    \begin{equation}
    	\label{eq:holling}
        F_{ij} = \frac{\omega_{ij} B_j^h}
        {B_0^h + c_i B_i
        	+ \sum_{k \in \{\mathrm{res}\}_i} \omega_{ik} B_k^h}
    \end{equation}
\end{subequations}

The growth function of the primary producers (eq. \@ref(eq:growth)) has the shape of a logistic growth
where the growth rate is maximal and null when
$\sum_{j \in \{\mathrm{prod}\}} \alpha_{ij} B_j$ are respectively close to 0 and 1. $\alpha_{ij}$ is the
per capita effect of the producer $j$ on the producer $i$, $K_i$ being the
carrying capacity for the producer $j$. The functional response of a consumer
feeding on a resource (eq. \@ref(eq:holling)) depends on the relative preference
of the consumer on the resource ($\omega_ij$) is limited by a half-saturation
rate ($B_0$), intraspecific interference coefficient ($c_i$), and on the
availability of its other resources
($\sum_{k \in \{\mathrm{res}\}_i} \omega_{ik} B_k^h$). Finally, the $h$ exponent
controls the shape of the functional response from a saturating function
($h = 2$, Holling type II) to a sigmoid ($h = 3$, Holling type III).

## Environmental stochasticity and response diversity

We added a stochastic natural mortality rate to the species dynamics,
representing a fluctuating environment. Based on the literature and previous
models [@vasseur_environmental_2007], we assumed that natural mortality rates
scales inversely with the species body mass ($d_i = d_0M_i^{-1/4}$), as all
other physiological parameters. We used a basal natural mortality rate of $d_0 =
.4$, as done in previous studies[@vasseur_environmental_2007].

We introduced environmental stochasticity by
considering stochastic variation of mortality rates. Doing so allows to consider
stochasticity without modifying species interactions themselves
[@vasseur_environmental_2007]. The stochastic part of the mortality rates
($\epsilon_{i, t}$) follow a normal distribution 0 centered
($\epsilon_{i, t} \sim N(0,\sigma^2_e)$) with a variance ($\sigma^2_e$). The mortality rates each
species at the time $t$ equals to $d_{i,t} = d_i e^{\epsilon_{i, t}}$. We did so
to ensure to not have negative mortality rates [@vasseur_environmental_2007]. We
simulated $\epsilon$ using an Ornstein-Uhlenbeck process, which is a modified
Brownian motion where the stochastic values tend to come back to the central
value (i.e. $\epsilon = 0$) such as $\epsilon_t$ varies according to this
stochastic differential equation: $d\epsilon_t = (0 - \epsilon_t)d_t + \sigma_e
dW_t$, where $dW_t$ is a Brownian motion. The strength of environmental
stochasticity was then controlled by $\sigma_e$.

Response diversity was controlled by the correlation among species stochastic
mortality rates ($\rho_{ij, i \neq j}$) such as response diversity is maximal
when species mortality rates are uncorrelated ($\rho = 0$) and null when species
mortality rates are perfectly correlated ($\rho = 1$) [@ripa_food_2003;
@vasseur_environmental_2007; @gouhier_synchrony_2010]. So the species mortality
rates were a product of the stochastic mortality rates generated by the
Ornstein-Uhlenbeck process and a variance-covariance matrix such as:

\begin{equation}
\begin{aligned}
\epsilon_{i, t} =
  \begin{bmatrix}
     \rho_{11}\sigma^2_e & \rho_{12}\sigma^2_e & \cdots & \rho_{1n}\sigma^2_e \\
     \rho_{21}\sigma^2_e & \rho_{22}\sigma^2_e & \cdots & \rho_{2n}\sigma^2_e \\
     \vdots  & \vdots  & \ddots & \vdots  \\
     \rho_{n1}\sigma^2_e & \rho_{n2}\sigma^2_e & \cdots & \rho_{nn}\sigma^2_e
  \end{bmatrix}
  \begin{bmatrix}
     a_{1,t} \\
     a_{2,t} \\
     \vdots  \\
     a_{n,t}
  \end{bmatrix}
S_{com} = S_{pop} \times \phi
(\#eq:rho)
\end{aligned}
\end{equation}

Where $a_{i,t}$ the stochastic value generated by the Ornstein-Uhlenbeck process
at time $t$, $\rho_{ii} = 1.0$ while all $\rho_{ij, i \neq j}$ were equal and
control the level of response diversity.

## Simulation

```{r}
tar_load(sim_param_tab)
spt <- sim_param_tab
get_sim_par_val <- function(x = "S", data = sim_param_tab) {
  x <- data[data$Parameter == x,][c("min", "max", "n")]
  setNames(unlist(x), c("min", "max", "n"))

}
pS <- get_sim_par_val(x = "S")
pct <- round(get_sim_par_val(x = "ct"), 2)
pZ <- get_sim_par_val(x = "Z")
ph <- get_sim_par_val(x = "h")
pe <- get_sim_par_val(x = "env_stoch")
prd <- get_sim_par_val(x = "resp_div")
pr <- get_sim_par_val(x = "rho")
```
\pagebreak

```{r}
met_cap <- "Examples of stochastic food-web simulations. a, Simulations with low
and high response diversity. b, simulations with low and high predator-prey mass
ratio (upper and bottom panels), low and high species richness (left and right
  panels). The food-web are displayed on the side with the node size
being proportional to the average species biomass across the simulation. We
also display the food-web structure and the stability metrics. Across all
simulations, $h = 2$, $\\sigma_e = .3$. Low and high response diversity were
respectively $\\rho = 1$ and $\\rho = 0$. Low and high PPMR were respectively 1 and 100." %>%
  str_replace_all("\\s+", " ")
```

```{r met, fig.cap = met_cap}
knitr::include_graphics(here("report", "figures", "p_methods.pdf"))
```

(ref:eq-rho) (eq. \@ref(eq:rho))

```{r}
tar_read(tab_sim_param) %>%
  kable(
    booktabs = TRUE,
    caption = "Parameter values used in the simulation experiment. Numbers
    between squared brackets display all the values. Two numbers between
    brackets are the minimum and maximal values, N being the number of unique
    values. PPMR: Predator Prey Mass Ratio. We defined response diversity as \\protect$1-\\rho$
    (ref:eq-rho).",
    label = "param") %>%
  kable_styling(latex_options = c("striped", "hold_position"))
```

We generated food-web structure using the niche model [@williams_simple_2000]
with an initial species richness from $`r pS["min"]`$ to $`r pS["max"]`$, and
connectance from $`r pct["min"]`$ to $`r pct["max"]`$ (Table
\@ref(tab:param)). We generated food-webs without cannibalistic links and
discarded food-webs that contained disconnected species.

We varied the strength of environmental stochasticity by varying $\sigma_e$ from
$`r pe["min"]`$ to $`r pe["max"]`$, the range of mortality rates observed in
protists [@vasseur_environmental_2007]. We decreased response diversity by
increasing $\rho$ from $`r pr["min"]`$ to $`r pr["max"]`$ (i.e. response
diversity being $1 - \rho$) (Fig. \@ref(fig:met)a).

To vary interaction strength and mortality rates (i.e. allometrically
scaled mortality rate $d$) in the community, we varied the Predator-Prey
Mass Ratio (PPMR) from $`r pZ["min"]`$ to $`r pZ["max"]`$ (Fig. \@ref(fig:met)b). We further varied
the functional response of consumer from type II to type III (i.e. from
$h = `r ph["min"]`$ to $h = `r ph["max"]`$, eq. \@ref(eq:holling)) as it was shown to
have a tremendous role on stability [@brose_allometric_2006]. We set a global
carrying capacity of $K' = 10$ to ensure that consumers were not limited by the
biomass input from primary producers. We standardised the carrying capacity by
the number of primary producers and by the interspecific competition among
producers to ensure that the effects of species richness on temporal stability
are not driven by the increase in the number of primary producers, thereby
trivially increase biomass input for consumers. We then standardised $K'$ such as
$K_i = \frac{1 + \alpha_{ij} (S - 1)}{S}K'$ [@ives_stability_2000].

We numerically solved the stochastic differential equation systems during 5000
timesteps and collected the last 500 timesteps. We set $d_t = .1$, and $d_t =
.05$ when instability was detected during the simulations due to stiff changes
in biomass. We set extinction threshold to $10^{-6}$ and if an extinction
happened in the last 500 timesteps, we continued to run the simulation for
another 1000 timesteps until there was no extinction events. When we found
disconnected species at the end of the simulation (i.e. a primary producer
without consumer or a consumer without prey), we removed them and re-run the
simulation for another 1000 timesteps.

## Simulation analysis

We measured food-web structure and dynamics that have been shown to be linked
with stability. We collected total biomass, species richness, connectance,
average trophic level weighted by biomass, average omnivory and average per
capita interaction strength. Connectance was computed as $C = L / S(S-1)$, $L$
being the number of trophic links and $S$ being the number of species in the
community. Species trophic level was computed recursively from the bottom of the
food-web, such as the trophic level of a consumer was equal to the average
trophic level of its resources added to 1, the trophic level of primary
producers being set to 1 [@christensen_ecopath_1992]. Average trophic level was then equal to:
$1/\hat{B}_i\sum_{i}\mathrm{TL}_i$, $\mathrm{TL}_i$ being the trophic level and
$\hat{B}_i$ the average biomass of the species $i$. The degree of omnivory of a
consumer was computed  such as the sum of squares of its
resource trophic levels weighted by the relative preference of the consumer for
each resource: $\sum_i (\mathrm{TL}_i - \hat{\mathrm{TL}})^2 \omega_i/ \sum \omega_i$ [@christensen_ecopath_1992].
Interaction strength was computed was quantified as the biomass fluxes going
from a resource to a consumer, such as: $I = x_i y_i B_k F_{ij}$ (eq.
\@ref(eq:cons)), and averaged over time. Per capita interaction strength was
obtained by dividing average interaction strength by the average biomass of the
consumer. Our way of measuring per interaction strength is very similar to
previous metrics used in previous studies, at the difference here that we
average fluxes over time and not the maximal interaction strength
[@mccann_weak_1998]. Finally, the average interaction strength of a community
was computed as the average per capita interaction strength at the exclusion of
null interaction strengths (i.e. excluding absence of trophic interactions).

We assessed the effects of food-web structure, environmental stochasticity and
response diversity on temporal stability of biomass. We measured temporal stability of biomass
as the inverse of the coefficient of variation. We further partitioned stability
into population stability ($S_{pop}$) and asynchrony ($\phi$) such as:

\begin{equation}
\begin{aligned}
S_{com} = S_{pop} \times \phi
(\#eq:scom)
\end{aligned}
\end{equation}

with $S_{com} = \mu_{tot} / \sigma_{tot}$, $S_{pop} = \mu_{tot} / \sum_i \sigma_i$,
$\phi = \sum_i \sigma_i / \sigma_{tot}$.

$\mu_{tot}$ and $\sigma_{tot}$ being respectively average total biomass and standard
deviation of total biomass.

To further get a mechanistic insight about the effects of stressors and food-web
structure on stability, we partitioned asynchrony into statistical averaging
effect (SAE) and compensatory dynamics effect (CPE) [@zhao_biodiversity_2022].
SAE quantities the portfolio effect emerging from independent species
fluctuations, while CPE describes compensatory dynamics emerging from species
interactions and differential responses to environmental stochasticity.

\begin{equation}
\begin{aligned}
S_{com} = S_{pop} \times SAE \times CPE
(\#eq:scom2)
\end{aligned}
\end{equation}

\begin{equation}
\begin{aligned}
SAE = SAE_{even} \times EVN
(\#eq:sae)
\end{aligned}
\end{equation}

Statistical averaging effect is the share of asynchrony assuming that species are
independent, meaning that the variance total of the community is equal to the
sum of species variances only (i.e. species covariances are null). We can
define community stability ($S_{com, IP}$) in that scenario and then derive $SAE$.

\begin{equation}
\begin{aligned}
S_{com, IP} &= \dfrac{\mu_{tot}}{\sqrt{\sum_i \sigma^2_i}} = SAE \times S_{pop} \\
SAE &= \dfrac{S_{com, IP}}{S_{pop}} = \dfrac{\sum_i \sigma_i}{\sqrt{\sum_i \sigma^2_i}}\\
(\#eq:sae2)
\end{aligned}
\end{equation}

As SAE is further dependent on the evenness of species variance, we
partitioned $SAE$ into $SAE_{even}$ and evenness $EVN$, such as  $EVN = SAE / SAE_{even}$.
$SAE_{even}$ corresponds to the $SAE$ when species quantities are perfectly
even such as from eq. \@ref(eq:sae2), we can obtain $SAE_{even}$ from which $EVN$ can be computed:

\begin{equation}
\begin{aligned}
SAE &= \dfrac{\sum_i \sigma_i}{\sqrt{\sum_i \sigma^2_i}} = \sqrt{\dfrac{1}{\sum_i (\sigma_i\sigma^2_i})^2} = \sqrt{\dfrac{1}{\sum_i (p_i)^2}}\\
SAE_{even} &= \sqrt{\dfrac{1}{\sum_i (1/S)^2}}\\
SAE_{even} &= \sqrt{S}\\
(\#eq:sae-even)
\end{aligned}
\end{equation}

Compensatory dynamics in the food-webs is then the remaining part of asynchrony:

\begin{equation}
\begin{aligned}
CPE &= \dfrac{S_{com}}{SAE \times S_{pop}} = \dfrac{\sqrt{\sum_i \sigma^2_i}}{\sigma_{tot}}\\
(\#eq:cpe)
\end{aligned}
\end{equation}

Compensatory dynamics can arise from predator-prey interactions
($CPE_{int}$) or from differential response of species to environmental
stochasticity ($CPE_{env}$). From the deterministic and stochastic parts of
mortality rates, we computed the compensatory dynamics linked to the environment
such as:

\begin{equation}
\begin{aligned}
CPE_{env} = \mathbf{B} \times \mathrm{diag}(d_i) \times e^\mathbf{E}
(\#eq:cpenv)
\end{aligned}
\end{equation}

With $\mathbf{B}$ and $\mathbf{E}$ being the matrices time x species for species biomass
and stochastic mortality rates respectively. Then compensatory dynamics raising
from species interactions were computed as: $CPE_{int} = CPE / CPE_{env}$.

## Statistical analysis

Using a structural equation model, we disentangled the complex effects of
food-web structure, environmental stochasticity and response diversity on the
different facets of temporal stability of biomass (i.e. population stability,
statistical averaging effects and compensatory dynamic effects). We set the
paths of the path of the model according to theoretical expectations. We tested
the effects of initial species richness, initial connectance, Predator-Prey size
ratio on secondary structure of the food-webs, i.e. realized species richness
and connectance, the average trophic level, the average omnivory and the
interaction strength. Further, we expected those food-web structure to alter
population stability and synchrony. Grounded in theoretical ecology, we expected
that higher average trophic level, average omnivory, lower interaction strength,
and type III functional response (relative to type II) increase population
stability [@mccann_weak_1998; @mccann_reevaluating_1997;
@shanafelt_stability_2018]. Higher Predator-Prey size ratio was also expected to
increase population stability [@brose_allometric_2006]. Finally, the effect of
environmental stochasticity was assumed to be mainly associated to population
stability (i.e. by increasing amplitude of population fluctuations).

Regarding asynchrony, we tested the effects of species richness, connectance,
average trophic level and interaction strength on statistical averaging effects.
We expected that those same food-web structural properties affected
compensatory dynamics due to interactions ($CPE_{int}$). Finally, we tested the
effects of species richness and species response diversity in mortality rates on
the compensatory dynamics due to stochastic fluctuations ($CPE_{env}$). We
completed the structural equation model with the mathematical relationship
linking compensatory dynamics and statistical averaging effects to asynchrony,
and from asynchrony and population stability to community stability.

Prior to test the structural equation model, we logged all the stability
components to transform their relation from multiplicative to additive (eq.
\@ref(eq:scom2)).
We ensure that all the linear models composing the structural equation model
presented low to moderate multicollinearity (Variance Inflation Factor < 10, Table SXX). The
structural equation model using the R package `PiecewiseSEM`
[@lefcheck_piecewisesem_2016]. We then bootstrapped (N = 100) the SEM to obtain
confidence interval on the estimated effects and computed the sum of the direct
and indirect effects using `semEff` R package [@murphy_semeff_2021]. In the
main text, we reported standardised coefficients which are obtained by scaling
the coefficients by the standard deviation of the response and predictor
variable (add formula). Finally, because we analysed simulation, we reported
only the direct standardised coefficients that had an absolute value above $0.1$.

In a second analysis, we aimed to test how stability-species richness
relationships were modulated by food-web structure and the two stressors:
environmental stochasticity and response diversity. We then modelled temporal
stability of biomass according to food-web structure: species richness, weighted
average trophic level, connectance, hill exponent, predator-prey mass ratio, and
average interaction strength, as well as environmental stochasticity and species
response diversity. We further included the two-way interactions between species
richness and food-web structure, plus between species richness and stressors.
Finally, we added the three-way interaction between species richness, food-web
structure and stressors. The VIF were inferior to 5, indicating low
multicollinearity. We then used the linear equation of the model and associated
slope coefficients to predict the shape of the relationship between community
stability and species richness according to a gradient of response diversity and
of food-web structure. To do so, we summed all the coefficients involving species
richness in the model, i.e. the one, two, and three-way terms. The values of the
control variables were set to the values indicated in the Fig. \@ref(fig:pred).

# Results

## Drivers of stability

```{r}
sem_cap <- paste0(
  "Structural Equation Model linking initial model parameters to resulting
  temporal stability of biomass. (a) Arrows display positive (continuous) and
  negative (dashed) standardised effects. Initial parameters were initial
  species richness, initial connectance, Prey-Predator Mass Ratio (PPMR), hill
  exponent, variance of stochastic mortality rates (Env. stochasticity), and
  species response diversity. Only the standardised coefficients whose absolute
  values are superior to 0.1 are displayed. A more complete SEM containing
  further partition of Statistical Averaging Effect and Compensatory effects is
  displayed in Fig. SXX. (b) Total effect of food-web structure on community stability,
  asynchrony, and population stability. The total effects were derived from the
  structural equation model. (c) Partial relationships derived from the SEM, points
  displaying the partial standardized residuals while lines displays the linear
  relationship derived from the SEM.")
```

```{r sem, fig.cap=sem_cap, fig.id = "sem"}
knitr::include_graphics(here("report", "figures", "p_sem_fig.pdf"))
```

```{r}
tar_load(c(sem_eff_tab, sem_simplified_eff_tab))

get_eff <- function(
  y = "pop_stab", x = "richness",
  d = 2, effect_type = "direct",
  tab = sem_simplified_eff_tab) {
  mask <- tab$response == y & tab$predictor == x &
    tab$effect_type == effect_type
  round(tab[mask, ]$effect, d)
}

rpt <- get_eff(y = "pop_stab", x = "w_avg_tlvl")
rph <- get_eff(y = "pop_stab", x = "h")
rpe <- get_eff(y = "pop_stab", x = "env_stoch")

rsr <- get_eff(y = "sae_total", x = "richness")
rst <- get_eff(y = "sae_total", x = "w_avg_tlvl")
rsi <- get_eff(y = "sae_total", x = "avg_int_strength")
rsc <- get_eff(y = "sae_total", x = "ct_alive")

rcr <- get_eff(y = "cpe", x = "richness")
rci <- get_eff(y = "cpe", x = "avg_int_strength")
rcd <- get_eff(y = "cpe", x = "resp_div")

# Detailled SEM coefficients
rsr <- get_eff(y = "sae_even", x = "richness", tab = sem_eff_tab)
ret <- get_eff(y = "evenness_sae", x = "w_avg_tlvl", tab = sem_eff_tab)
rer <- get_eff(y = "evenness_sae", x = "richness", tab = sem_eff_tab)
rea <- get_eff(y = "evenness_sae", x = "avg_int_strength", tab = sem_eff_tab)
rec <- get_eff(y = "evenness_sae", x = "ct_alive", tab = sem_eff_tab)
reh <- get_eff(y = "evenness_sae", x = "h", tab = sem_eff_tab)

ria <- get_eff(y = "cpe_int", x = "avg_int_strength", tab = sem_eff_tab)
ric <- get_eff(y = "cpe_int", x = "ct_alive", tab = sem_eff_tab)
rih <- get_eff(y = "cpe_int", x = "h", tab = sem_eff_tab)
rir <- get_eff(y = "cpe_int", x = "richness", tab = sem_eff_tab)

rvr <- get_eff(y = "cpe_env", x = "richness", tab = sem_eff_tab)
rvd <- get_eff(y = "cpe_env", x = "resp_div", tab = sem_eff_tab)
```
```{r, eval = FALSE}
sem_simplified_eff_tab %>%
  filter(abs(effect) > .1, effect_type == "direct")
```

We found that food-web structure had significant effect on temporal population
stability such as average trophic level and type III functional response (i.e.
hill exponent = 3) increase population stability (resp. $r_\delta$= `r rpt` and
`r rph`, $r_\delta$ being the standardised coefficient, Fig. \@ref(fig:sem)a).
In turn, the variance of environmental stochasticity had a strong negative effect
on population stability ($r_\delta$= `r rpe`).

We further found that food-web structure had a strong impact on statistical
averaging effects, which raises from independent species fluctuations weighted
the evenness of their variance. As expected the mathematical relationship (eq.
\@ref(eq:sae-even)), species richness was a main positive driver of statistical
averaging effect ($r_\delta$= `r rsr`, Fig. \@ref(fig:sem)a), thereby increasing
asynchrony. Average trophic level and average interaction strength had strong
negative effects on statistical averaging effects (resp. $r_\delta$= `r rst` and
`r rsi`), thereby downplaying the portfolio effect raised from species richness.
In turn, we found that average interaction strength had a small positive effect
on compensatory dynamics ($r_\delta$= `r rci`, Fig. \@ref(fig:sem)). Species
response diversity had a strong positive effect on compensatory dynamics
($r_\delta$= `r rcd`).

```{r}

rscp <- get_eff(y = "stab_com", x = "pop_stab")
rsca <- get_eff(y = "stab_com", x = "async")
#rscp <- get_eff(y = "stab_com", x = "pop_stab", tab = sem_eff_tab)
#rsca <- get_eff(y = "stab_com", x = "async", tab = sem_eff_tab)

rac <- get_eff(y = "async", x = "cpe")
ras <- get_eff(y = "async", x = "sae_total")
#rac <- get_eff(y = "async", x = "cpe")
#ras <- get_eff(y = "async", x = "sae_total")

#
rav <- get_eff(y = "async", x = "cpe_env", tab = sem_eff_tab)
rae <- get_eff(y = "async", x = "evenness_sae", tab = sem_eff_tab)

rass <- get_eff(y = "async", x = "sae_even", tab = sem_eff_tab)
rasi <- get_eff(y = "async", x = "cpe_int", tab = sem_eff_tab)
```

Population stability had a positive effect twice higher than asynchrony on the
temporal stability of community (resp. $r_\delta$= `r rscp` and `r rsca`). In
turn, asynchrony was equally driven by statistical averaging effects and
compensatory dynamics (resp. $r_\delta$= `r rac` and `r ras`, Fig.
\@ref(fig:sem)a).

```{r, eval = FALSE}
semtot_cap <- "Total effect of food-web structure on community stability,
asynchrony, and population stability. The total effects were derived from the
structural equation model."
```

```{r semtot, eval = FALSE, fig.cap = semtot_cap, fig.id = "semtot"}
knitr::include_graphics(here::here("report", "figures", "p_sem_tot_stab.png"))
```

```{r}
rtsdiv <- get_eff(y = "stab_com", x = "resp_div", effect_type = "total")
rtsenv <- get_eff(y = "stab_com", x = "env_stoch", effect_type = "total")

rtsr <- get_eff(y = "stab_com", x = "richness", effect_type = "total")
rtst <- get_eff(y = "stab_com", x = "w_avg_tlvl", effect_type = "total")
rtsz <- get_eff(y = "stab_com", x = "Z", effect_type = "total")
rtsh <- get_eff(y = "stab_com", x = "h", effect_type = "total")
rtsc <- get_eff(y = "stab_com", x = "ct_alive", effect_type = "total")
rtsi <- get_eff(y = "stab_com", x = "avg_int_strength", effect_type = "total")

rtpr <- get_eff(y = "pop_stab", x = "richness", effect_type = "total")
rtpt <- get_eff(y = "pop_stab", x = "w_avg_tlvl", effect_type = "total")
rtpz <- get_eff(y = "pop_stab", x = "Z", effect_type = "total")
rtph <- get_eff(y = "pop_stab", x = "h", effect_type = "total")
rtpc <- get_eff(y = "pop_stab", x = "ct_alive", effect_type = "total")
rtpi <- get_eff(y = "pop_stab", x = "avg_int_strength", effect_type = "total")

rtar <- get_eff(y = "async", x = "richness", effect_type = "total")
rtat <- get_eff(y = "async", x = "w_avg_tlvl", effect_type = "total")
rtaz <- get_eff(y = "async", x = "Z", effect_type = "total")
rtah <- get_eff(y = "async", x = "h", effect_type = "total")
rtac <- get_eff(y = "async", x = "ct_alive", effect_type = "total")
rtai <- get_eff(y = "async", x = "avg_int_strength", effect_type = "total")
```

Deriving the sum of direct and indirect effects from the structural equation
model, we show that both environmental stochasticity and response diversity among
species had by far the most important effect on community stability (resp.
$r_\delta$= `r rtsenv` and `r rtsdiv`, Fig. \@ref(fig:sem)b). Species richness
had a low overall positive total effect on community stability (`r rtsr`)
because its strong total positive effect on asynchrony (`r rtar`) was downplayed
by its total negative effect on population stability (`r rtpr`) and that the
weight of population stability on community stability is twice higher than the
one of asynchrony. Average trophic level, type III functional response and
predator prey mass ratio had positive effects on community stability in
total (resp. `r rtst`, `r rtsh`, `r rtsz`, Fig. \@ref(fig:sem)b), driven by
their positive effects on population stability that were weakly counterbalanced
by their negative effects on asynchrony. In turn, connectance and average
interaction strength had low total negative effects on community stability (resp.
`r rtsc` and `r rtsi`, Fig. \@ref(fig:sem)b). While connectance had negative
effects both on population stability and asynchrony, average interaction
strength had a low total positive effect on population stability counterbalanced
by a stronger negative effect on asynchrony.

## Stability-richness relationship

```{r}
tar_load(default_variable_values_pred)
pv <- default_variable_values_pred
pv_cap <- paste0(
  "connectance = ", pv["ct_alive"],
  ", average interaction strength = ", pv["avg_int_strength"],
  ", average trophic level = ", pv["w_avg_tlvl"],
  ", hill exponent = ", pv["h"],
  ", Predator-Prey Mass Ration = ", pv["Z"],
  ", response diversity = ", pv["resp_div"],
  ", and environmental stochasticity = ", pv["env_stoch"]

)
pred_cap <- paste0("Relationship between community stability and species
  richness according to food-web structure and response diversity. a,
  Predictions of the statistical model linking the temporal stability of
  community biomass and species richness. The values chosen for high/low
  response diversity and food-web structure are indicated by the black point in
  (b). Lines display the mean predictions of the model, ribbons display the
  confidence intervals at 95\\% (to recheck). b, Diagrams displaying the slope
  coefficients of the effect of species richness on community stability. Default
  parameter values for the predictions: ", pv_cap, ".") %>%
  str_replace_all("\\s+", " ")
```

```{r pred, fig.cap = pred_cap, fig.id = "pred"}
knitr::include_graphics(here("report", "figures", "p_pred_sr.pdf"))
```

Using a more complex linear model (i.e. including interactions), We found strong
interactions between food-web structure and response diversity on the
relationship between community stability and species richness. We found that
high response diversity almost always leads to positive relationships between
community stability and species richness, no matter the values of average
interaction strength, average trophic level or connectance (Fig.
\@ref(fig:pred)). Furthermore, we found that higher response diversity increased
community stability in average (Fig. \@ref(fig:pred)a). Surprisingly, despite
the relatively small effects of food-web structure on overall community
stability (Fig. \@ref(fig:sem)c \@ref(fig:pred)a), we found that response
diversity dramatically change the shape of the community stability-species
richness relationship from positive to negative when response diversity is low
(Fig. \@ref(fig:pred)b).

---
#We found that
#environmental stochasticity decrease overall community stability but has only
#little effect on the shape of the community stability-species richness
#relationship (Fig. \@ref(fig:pred)a)
---

```{r, eval = FALSE}
tar_read(p_int_fw_str)
```

# Discussion

## Food-web structure and stressors affects population stability and asynchrony

Our results show that community stability in complex food-webs is mainly driven
by population stability rather than synchrony, confirming previous results in
empirical fish communities [@danet_species_2021], which contrasts with empirical
data in birds and bats assemblages that reported an equivalent influence of
population stability and asynchrony on community stability
[@olivier_urbanization_2020]. Although the number of empirical and theoretical
studies is still limited to compare the respective contribution of asynchrony and
population stability to community stability in competitive assemblages and
food-webs, we suggest two non-exclusive and complementary explanations that
might drive a lower contribution of asynchrony in food-webs compared to
competitive assemblages. First, the range of bodymasses is much higher in
food-webs, which typically include species that differ in bodymass by several orders of
magnitude [@brose_predator_2019], which in turn creates highly uneven biomass
distribution, thereby creating uneven contributions of species to asynchrony
[@thibaut_understanding_2013] and limits the raise of high portfolio effects and
compensatory dynamics in species rich communities [@zhao_biodiversity_2022].
Second, the trophic interactions characterising food-webs create strong
interdependence among species, which in turn limits the raise of independent
fluctuations (i.e. portfolio effects, but rise compensatory dynamics). The question of
which mechanisms and community features drive differences in the
contribution of asynchrony and population stability in food-webs and competitive
assemblages is still an open question that might be crucial to accurately
address diagnostics and conservation issues. On top of food-web structure,
we found that environmental stochasticity and environmental correlation among
species were the main drivers of community stability, suggesting that stressors
are the main drivers of temporal stability. This result is in accordance with previous
studies showing strong effects on environmental stochasticity
[@vasseur_environmental_2007] and the loss of response diversity decrease
community stability in tri-trophic food-webs [@ives_stability_2000;
@thebault_trophic_2005].

We further found that average trophic level increased average population
stability, confirming previous theoretical [@brose_allometric_2006;
@shanafelt_stability_2018; @eschenbrenner_diversity_2023] and
[@danet_species_2021] empirical results. Higher average trophic level can
stabilise populations because higher predators are themselves more stable due
to lower mortality rates with in turn makes them less sensible stochastic
perturbation and because they control the amplitude of the fluctuations of their
preys [@shanafelt_stability_2018, we can test that?]. In turn, we found that
average trophic level decreases strongly asynchrony through a negative effect on
statistical averaging effect. Further analysis showed that average trophic level
limited statistical averaging effect by decreasing the evenness of biomass
variability. It is linked to the model assumption that species of higher trophic
level have a lower overall mortality rates, similarly to previous studies
[@vasseur_environmental_2007], thereby increasing average population stability and simultaneously
synchronising populations. The fact that Predator Size Mass Ratio displays the same
effect, thereby building food-webs with bigger species having lower mortality
rates, further lends toward this hypothesis. Furthermore, average omnivory was
strongly correlated with average trophic level, suggesting that average omnivory
also stabilise population stability, in line with previous theoretical results
in small modules [@mccann_reevaluating_1997]. Our results further suggest that
higher average trophic level results in lower compensatory dynamics raising from
species interactions. It might be explained by the fact that average trophic
level is strongly associated with average omnivory, which might tamper
synchronized predator-prey oscillations that can raise compensatory dynamics
[@mccann_reevaluating_1997].

We however found no big effect of species richness on population stability,
which has been reported to be negative, positive or neutral in previous studies
[@ruijven_contrasting_2007; @houlahan_negative_2018]. [interpretation?].
However, species richness had a strong positive on asynchrony, by raising
statistical averaging effects. This result is common in empirical systems
[@tilman_biodiversity_2006; @olivier_urbanization_2020], but strongly depends on
the presence of response diversity [@loreau_species_2008;
@thibaut_understanding_2013; @zhao_biodiversity_2022]. Connectance, quantifying
the generalism in food-webs, had negative effects on both asynchrony and population
stability. We further found that higher connectance leaded to less even biomass
variability, thereby limiting the raise of the portfolio effects, and leaded to
lower compensatory dynamics due to species interactions. This result is in
accordance with empirical findings that a generalist predator can indeed
synchronises the dynamics of a large number of preys
[@raimondo_interspecific_2004]. [No idea about connectance decreasing population
stability]. However, the overall effect of connectance on overall stability was
quite low, similarly to previous empirical findings [@danet_species_2021], but
in apparent disagreement with the predominant role of connectance on stability
in theoretical studies [complexity-stability paradox, @may_will_1972].

---
#In turn,
#we found that species richness magnified the effect of the loss of response
#diversity on the compensatory dynamics due to environment. Our interpretation is
#that higher species richness synchronises more species fluctuations due to
#environmental mortality when species are not independent because each added
#species adds (S - 1, S being richness) positive covariance terms in the
#stochastic mortality rates, thereby increasing the overall synchrony in
#mortality rates when cross correlation are not null. This assumption that each
#species are equally correlated to each other seems unlikely to be true but as
#far we know, there is compelling evidence yet for alternative correlation
#structures.
---

Food-web dynamical features, the type of functional response (from II to III) and average
interaction strength, also had an effect on community stability. Our results
confirm the stabilising effect of the functional response of type III on
populations [@uszko_effects_2017; @brose_allometric_2006]. Compared to type II,
functional response of type III models a decrease of the feeding rates of
consumers when their preys have a low biomass which leads to a less dramatic
fluctuation of prey biomass. Furthermore, we found that functional response of
type III decreased asynchrony by decreasing compensatory dynamics due to species
interactions. This effect might be due to prey switching of consumers when one of
them reach low abundance [@mccann_diversitystability_2000]. Average interaction
strength, as connectance, had overall little effect on community stability.
Average interaction strength was found to decrease asynchrony by decreasing the
evenness of biomass variability, thereby decreasing portfolio effects,
despite having a small positive effect on compensatory dynamics. In turn, we
surprisingly found that average interaction strength had little effects on
population stability, which contrasts with the importance of this parameter in
complexity-stability relationship [@may_will_1972].

## Stability-richness relationship

We found that environmental stochasticity does not affect the shape of the
community stability - richness but rather control the overall community
stability (i.e. the intercept of the relation). At the opposite, variation in
response diversity had a strong impact on the relationship between stability and
richness. The reason is that the strength of environmental stochasticity mainly
affect population stability and less asynchrony, the latter being the driver of
species-richness relationships [@ives_stability_2000; @thebault_trophic_2005;
@thibaut_understanding_2013]. Average interaction strength and connectance that
were found to have little effects on overall stability, were found to have a
strong effect on stability-richness relationships. This result is in accordance
with the complexity-stability theory that predicts that community stability
(i.e. measured as the dominant eigen value of the jacobian matrix) should
decrease as the community is species diverse, has higher connectance, and
higher interaction strength [@may_will_1972]. However, we found that the
negative effects of complexity on stability-richness relationship are cancelled
when response diversity is high. An interpretation is that environmental
stochasticity itself does not override the negative complexity-stability
relationships, but it the presence of response diversity and resulting
asynchrony in species fluctuations which overrides the negative effects of
complexity on stability-richness relationships. Response diversity might then be one
mechanism that explains why so much positive relationship between stability and
species richness are found in empirical settings while food-web theoretical
model that do not often include environmental stochasticity and response
diversity finds negative complexity stability relationships
[@ives_stability_2007].

# Conclusion and perspectives

Our study presents the effects of food-web structure on temporal community
stability and its partition into population stability and asynchrony in complex
food-webs, added to two pressing stressors in the context of global change, the
strength of environmental variability and the level of response diversity of
species facing this environmental variability. Our results confirm the results
of previous theoretical models modelling small food-web modules or food-chain,
as we report that average trophic level, omnivory, predator-prey size ratio, and
functional response of type III increase temporal community stability by
stabilising population fluctuations. In turn, our study confirm that species
richness increase temporal community stability by increasing asynchrony due to
portfolio effects and thanks to response diversity. Our study reports that
connectance and average interaction strength have little effects on overall
community stability while they have a strong effect on the sign of the community
stability-richness relationship, in accordance with the complexity-stability
paradox, but only in the absence of response diversity. Overall, it suggests that
high response diversity of species facing environmental stochasticity, on top of
being the essential ingredient driving positive stability-richness relationship,
also cancel the negative effects of food-web complexity on stability-richness
relationship.

Our results highlight that population stability is the primary driver of
community stability. In the context of global changes in which environmental
variability is expected to increase tremendously, we expect to see a strong
decrease in community stability, despite high response diversity. But ongoing
biotic homogenization [@olden_homogocene_2018; @muthukrishnan_invasive_2020;
@wang_biotic_2021] is also suggested to decrease response diversity
[@wang_biotic_2021; @su_human_2021], which in turn is expected to further
decrease stability despite high species richness, even more in complex food-webs
having high interaction strength, high connectance and high average trophic
level. Then, the results of our study reinforces that the restoration of
ecosystems along with a restoration of landscape complexity that provides
refuges from extreme environmental changes will be key to maintain temporal stability of
ecological communities in the coming decades [@ryser_landscape_2021].

# References {-}