### 2024-04-04 paper/supplementary_figures.Rmd

---
title: "Supplementary information"
subtitle: ""
author: ""
date: ""
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    keep_tex: true
    spacing: double
    number_sections: false
  officedown::rdocx_document:
bibliography: "bibliography.bib"
link-citations: true
toc: false
linestretch: 2.0
header-includes:
  - \usepackage{float}
  - \renewcommand{\thefigure}{S\arabic{figure}}
  - \renewcommand{\thetable}{S\arabic{table}}
---

```{r setup, include=FALSE}
library(tidyverse)
library(magrittr)
library(rmarkdown)
library(officedown)
knitr::opts_chunk$set(echo = FALSE,
                      collapse = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      comment = "#>",
                      fig.path = "graphics/knitr-",
                      fig.retina = 1.618, # Control using dpi
                      fig.width = 6,  # generated images
                      fig.pos = "H",  # pdf mode
                      #fig.align = "center",
                      dpi = if (knitr::is_latex_output()) 72 else 300,
                      out.width = "100%",
                      dev = "png",
                      dev.args = list(png = list(type = "cairo-png")),
                      optipng = "-o1 -quiet")

library(here)
library(cowplot)
library(targets)
library(kableExtra)
library(coda)
library(ggbreak)

sapply(list.files(here("R"), full.names = TRUE), source)
if (Sys.info()["login"] == "alain") {
  file.copy(
    from = "~/Documents/post-these/references.bib",
    to = here("paper", "bibliography.bib"),
    overwrite = TRUE
  )
}
```

# Figures

(ref:sem-sens-cap) Total standardised effects for the sem displayed in main text
according to different simulation set-ups. Default: simulations used in main
text; Fixed death rate: non-allometric death rate, meaning that all species have
the same death rate, here $d = 0.1$; K not standardised: carrying capacity not
standardised; rebuild FW after removing disconnected species: instead of setting
starting biomass of disconnected species to 0, we removed them from the
food-webs, and then recompute species trophic level, species bodymass and then
their metabolic rates, prey preferences, producer carrying capacity.

```{r p-sem-sensibility, fig.cap="(ref:sem-sens-cap)"}
knitr::include_graphics(here("report", "figures", "p_sem_main_text_eff_tab_tot.pdf"))
```

(ref:cor-cap) Spearman's correlation coeffients among stability components,
food-web structure, and model parameters.

```{r p-cor, fig.cap="(ref:cor-cap)"}
knitr::include_graphics(here("report", "figures", "p_cor.pdf"))
```

(ref:dharma-lm-cap) Distribution of residuals for the linear models predicting the
relationship between community stability and species richness (Fig. 3). a,b,c
for respectively community stability, population stabilty and asynchrony. Left
panels display the distribution of the quantiles, and indicate that residuals
are quite well distributed. Right panels display the scaled residuals versus the
predicted values. The slight U-shape might indicate the lack of a polynomial
term in the model, but overall the model looks to have appropriate fitting.

```{r dharma-lm, fig.cap="(ref:dharma-lm-cap)"}
knitr::include_graphics(here("report", "figures", "p_diagnotisc_model_fw_stab_comp.jpg"))
```

(ref:coeff-lm-cap) Standardised coefficients for the linear model predicting the
relationship between community stability and species richness (Fig. 3).

```{r coeff-lm, fig.cap="(ref:coeff-lm-cap)"}
knitr::include_graphics(here("report", "figures", "p_model_fw_stab_comp_scaled_coeff.pdf"))
```

```{r}
tar_load(default_variable_values_pred)
pv <- default_variable_values_pred

pv_cap <- paste0(
  "connectance = ", round(pv["ct_alive"], 2),
  ", average interaction strength = ", round(pv["avg_int_strength"], 3),
  ", average trophic level = ", round(pv["w_avg_tlvl"], 2),
  ", predator interference = ", round(pv["c"], 1),
  ", Predator-Prey Mass Ratio = ", round(pv["Z"]),
  ", and environmental stochasticity = ", round(pv["env_stoch"], 1)
)

p_pred_stab_comp_rho_cap <- paste0("Prediction of the statistical model linking
  the temporal stability of community biomass and species richness. Compared to
  Fig.3, we added the prediction of the slope of the relation population
  stability-species richness and asynchrony-species richness. Default
  parameter values: ", pv_cap,
  ".") %>%
str_replace_all("\\s+", " ")
```

```{r stab-comp-pred, fig.cap = p_pred_stab_comp_rho_cap}
knitr::include_graphics(here("report", "figures", "p_pred_sr_resp_div_supp.pdf"))
```

\pagebreak

# Table

```{r}
tar_read(tab_sem_vif) %>%
  kable(booktabs = TRUE,
    format = "latex",
    caption = "Variance Inflation Factors for the Structural Equation Model
    displayed in main text (Fig. 2). All VIF were inferior to 3, indicating low
    multicollinearity.",
    label = "sem-vif"
    ) %>%
  column_spec(1, bold = TRUE) %>%
  collapse_rows(columns = 1, latex_hline = "major", valign = "middle") %>%
  kable_styling(latex_options = c("striped", "hold_position"))
```

```{r}
tar_read(tab_model_fw_stab_comp_wo_int_vif) %>%
  kable(booktabs = TRUE,
    format = "latex",
    caption = "Variance Inflation Factors for the linear model predicting the
    relationship between community stability and species richness (Fig. 3). All
    VIF were inferior to 3, indicating low multicollinearity.",
    label = "lm-vif"
    ) %>%
  kable_styling(latex_options = c("striped", "hold_position"))
```

```{r}
tar_read(tab_sem_coeff) %>%
  kable(booktabs = TRUE,
    format = "latex",
    caption = "Standardised estimates from the Structural Equation Model
    displayed in main text (Fig. 2).",
    label = "sem-coeff"
    ) %>%
collapse_rows(columns = 1, latex_hline = "major", valign = "middle") %>%
kable_styling(latex_options = c("striped", "hold_position"))
```

```{r}
tar_read(tab_sem_simplified_eff) %>%
  kable(booktabs = TRUE,
    format = "latex",
    caption = "Standardised estimates from the supplementary Structural Equation
    Model which includes the partition of compensatory effects rising from
    environmental fluctuations and species interactions.",
    label = "sem-supp-coeff") %>%
collapse_rows(columns = 1, latex_hline = "major", valign = "middle") %>%
kable_styling(latex_options = c("striped", "hold_position"))
```