### 2024-08-29 paper/manuscript.Rmd

---
title: Response diversity is a major driver of temporal stability in complex food webs
author: "Alain Danet (1), Sonia Kéfi (2,3), Thomas F. Johnson (1), Andrew P. Beckerman (1)"
date: ""
output:
  bookdown::pdf_document2:
    latex_engine: xelatex
    keep_tex: true
    spacing: double
    number_sections: false
  officedown::rdocx_document:
    number_sections: false
    base_format: "bookdown::word_document2"
always_allow_html: true
bibliography: references.bib
link-citations: true
csl: nature.csl
#csl: ecology-letters.csl
toc: false
linestretch: 2.0
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{hyperref}
  - \hypersetup{backref=true}
  - \usepackage{setspace}\doublespacing
---

```{r, echo = FALSE}
knitr::opts_chunk$set(
  cache = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.pos = "H",  # pdf mode
  #fig.dim = c(7, 7),
  fig.fullwidth = TRUE,
  fig.show = "hold",
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  results = TRUE,
  fig.path="figures/"
)
```

```{r, results = FALSE}
sapply(c("targets", "tidyverse", "magrittr", "cowplot", "here", "piecewiseSEM",
    "semEff", "ggcorrplot", "kableExtra", "rmarkdown", "officedown"),
  require,
  character.only = TRUE)
mytheme <- theme_cowplot(font_size = 8) +
background_grid() +
theme(legend.position = "bottom")
theme_set(mytheme)
sapply(list.files(here("R"), full.names = TRUE), source)
```

- 1 School of Biosciences, Ecology and Evolutionary Biology, University of Sheffield, Western Bank, Sheffield, S10 2TN, UK
- 2 ISEM, CNRS, Univ. Montpellier, IRD, EPHE, Montpellier, France
- 3 Santa Fe Institute, 1399 Hyde Park Road, Santa Fe, NM 87501, USA

---
## IDEA:
# Talk about about how stochasticity minor interaction effects similar to
# dispersal (Holt, 1985; other papers I guess)
---

# Abstract

Global change constitutes a suite of major threats to biodiversity and ecosystem
functioning. These threats can materialise via changes in the temporal stability
of ecological communities and the services they provide. However, the majority
of research on stability has focused on single trophic level communities and has
not yet integrated classic theory about species richness and food web structure
with more recent theory centred on response diversity and stochasticity. Using a
stochastic bioenergenetic food-web model, we reveal that response diversity is a
major driver of community stability. Moreover, positive stability-richness
relationships emerge only in the presence of response diversity. In contrast to
previous work,  food-web structural properties are only secondary drivers of
overall community stability, but interact with response diversity to determine
the sign of the stability-richness relationship. Our study highlights the
complex pathways by which food-web structure and response diversity drive
community stability, and raises concerns about how the loss of response
diversity may lead to a breakdown of stability and the capacity for these
communities to deliver functions and services to human societies.

# Introduction

Global change, characterised by multiple, simultaneous threats to
biodiversity[@lee_ipcc_2023; @simmons_refocusing_2021], is having impacts on the
temporal stability of complex ecological communities[@olivier_urbanization_2020;
@de_boeck_patterns_2018; @hansen_climate_2013; @bluthgen_land_2016;
@harris_biological_2018]. A breakdown of stability impacts the capacity for
these communities to deliver functions and services to human
societies[@loreau_ecological_2022]. Predicting and managing the stability and
services provided by biodiverse communities requires reconciling multiple
concepts and theories about temporal stability in species diverse and
structurally complex communities. These include classic ideas about the role of
trophic structure and interaction strengths[@may_will_1972;
@angelis_stability_1975; @mccann_weak_1998; @mccann_reevaluating_1997;
@thebault_trophic_2005; @thebault_relationship_2006; @thebault_stability_2010],
newer concepts of response diversity, average population stability, asynchrony
and portfolio effects and integrative theories about the relationship between
stability, structure and diversity.

An abundant literature has established that species diversity can buffer the
effect of environmental perturbations on ecological communities, thereby
increasing the temporal stability (hereafter “stability”) of
communities[@yachi_biodiversity_1999; @doak_statistical_1998;
@tilman_ecological_1999; @de_mazancourt_predicting_2013;
@thibaut_diversity_2012; @zhao_biodiversity_2022]. Because different species
have diverse niches and environmental preferences, they may react differently to
the same environmental change [@elmqvist_response_2003; @ross_how_2023]. This
diversity of species responses to environmental change, coined as “response
diversity”, has been identified as a crucial mechanism that buffers the response
of ecological communities to environmental perturbations
[@yachi_biodiversity_1999; @thibaut_diversity_2012;
@thibaut_understanding_2013]. For example, in a given community, some species
will respond positively and others negatively to changes in temperature,
depending on their thermal niche [@frund_response_2013]. Furthermore, it has
also been shown recently that the stability of communities can be decomposed
into the average population stability and the asynchrony of population
fluctuations [@thibaut_understanding_2013; @zhao_biodiversity_2022]. Average
population stability quantifies the average variability in species abundances,
while asynchrony refers to how much the temporal fluctuations of the different
species of the community are buffering each other. Recent studies have shown
that asynchrony can be further resolved into “portfolio effects”, the
statistical averaging of independent species fluctuations in response to
environmental perturbations, and “compensatory dynamics”, where one species’
decrease can be compensated for by the increase in another
species[@doak_statistical_1998; @thibaut_understanding_2013;
@zhao_biodiversity_2022].

These mechanisms have been well-established in data from single trophic level
(e.g. competitive) communities, such as grasslands, butterflies, birds, and bats
communities[@de_mazancourt_predicting_2013; @olivier_urbanization_2020]. In
these studies, response diversity increases asynchrony, and is a more important
driver of community stability than population
stability[@olivier_urbanization_2020]. In these same studies, the portfolio
effect was found to play a key role in the emergence of positive
diversity-stability relationships in experimental
grasslands[@zhao_biodiversity_2022].

Despite these data, there is limited evidence and understanding for how these
collections of theories integrate to define stability in food webs with
substantial diversity and vertical complexity (i.e. trophic levels)
[@yang_predictability_2019; @danet_species_2021; @eschenbrenner_diversity_2023;
@srednick_understanding_2024]. Classic theory and concepts about
stability-complexity [@may_will_1972; @angelis_stability_1975],
stability-structure [@mccann_weak_1998; @mccann_reevaluating_1997;
@thebault_trophic_2005; @eschenbrenner_diversity_2023] and stability-richness [@ives_stability_2000;
@thebault_trophic_2005; @thebault_relationship_2006] offer some expectation.
They suggest that population stability is expected to decrease with increasing
interaction strength[@may_will_1972; @mccann_reevaluating_1997;
@mccann_weak_1998] and food web connectance[@thebault_trophic_2005] but to
increase with average trophic level [@shanafelt_stability_2018;
@danet_species_2021]. As a result, the distribution and strength of species
interactions, as well as the average trophic level, can have strong effects on
community stability as well as on stability-richness
relationships[@thebault_trophic_2005; @thebault_relationship_2006]. These
relationships indicate that the structure of food-webs strongly influences
population stability and stability-richness relationships.

The more recent theory and concepts centred on response diversity and asynchrony
add further predictions[@thibaut_understanding_2013; @zhao_biodiversity_2022].
With vertical structure and trophic interactions, consumers can synchronise the
dynamics of their prey[@raimondo_interspecific_2004;
@korpimaki_predatorinduced_2005], can offset the effects of response diversity
on asynchrony and can generate substantial compensatory dynamics via prey
switching and trophic cascades[@mccann_weak_1998; @mccann_diversitystability_2000;
@fahimipour_compensation_2017].

Despite this collection of historical and more modern theory, we don’t know what
determines stability in vertically structured, species rich communities.
Against this gap in understanding, we specifically investigate the effects of
response diversity and food-web structure on community stability and on
stability-richness relationships in a variable  environment. The key question is
whether the new concepts of environmental stochasticity, response diversity and
asynchrony are the main mechanisms driving community stability versus more
classic concepts tied to food-web structure and diversity.

We explore this key question by extending  the  bioenergetic food web
model[@brose_allometric_2006; @yodzis_body_1992], to incorporate environmental
stochasticity and response diversity[@vasseur_environmental_2007;
@ives_stability_2007; @ripa_food_2003]. We introduced environmental
stochasticity by making species death rates
stochastic[@vasseur_environmental_2007]. Response diversity is
introduced as the correlation of stochasticity among species, so that species
that experience environmental changes in the same way show positively correlated
variations in their death rates.

We generated over 45000 in-silico communities with diverse richness, trophic
levels and connectance and experimentally simulated variable degrees of
interaction strength, response diversity and environmental stochasticity.  For
each community, we evaluated the effects on stability of species richness,
response diversity, community structure and stochasticity.  Stability was itself
partitioned in our analyses into population stability and asynchrony (i.e.
portfolio and compensatory effects). Finally, we also evaluated the occurrence
of negative and positive relationships between community stability and species
richness.

```{r}
tar_load(sim_param_tab)
p_mm <- function(
  v = "S",
  data = sim_param_tab
) {
  out <- data[data$Parameter %in% v, ]
  paste0("", signif(out[["min"]], 1), "-", signif(out[["max"]], 1), "")
}
```

```{r, eval = FALSE}
p_mm("ct")
```

# Results

```{r}

sem_cap <- paste0(
  "Structural Equation Model linking food-web structure and stressors to the
  temporal stability of community biomass. a, Continuous blue arrows display
  positive and dashed red arrows display negative standardised effects, width of
  the arrows being proportional to the absolute values of the standardised
  effects. Only the standardised coefficients whose absolute values are superior
  to 0.05 are displayed. The full table of coefficients including control
  variables is displayed in Table S2. b, Total effects of food-web structure on
  community stability, asynchrony, and population stability derived from the
  structural equation model. N = ", nrow(tar_read(sim_fw)), " simulations.") %>%
  str_replace_all("\\s+", " ")
```

```{r sem, fig.cap=sem_cap, fig.id = "sem"}
knitr::include_graphics(here::here("report", "figures", "p_sem_fig.pdf"))
```

```{r}
tar_load(c(sem_doak_eff_tab, sem_simplified_eff_tab))

```

```{r fig2-coeff, eval = FALSE}
tar_read(sem_fig_coeff)
tar_read(sem_doak_rsquared_csc_rerun)
```

```{r}
tar_load(sim_summary_stat)
m_sim <- function(
  v = "stab_com",
  data = sim_summary_stat,
  col = "Median (5%, 95%)"
  ) {
  data[data$var %in% v, ][[col]]
}
```
```{r, eval = FALSE}
m_sim(v = c("stab_com", "async"))
m_sim(c("stab_com", "async"))
m_sim(col = "n")
```

```{r}
get_eff <- function(
  y = "pop_stab", x = "richness",
  d = 2, effect_type = "direct",
  tab = sem_doak_eff_tab) {
  mask <- tab$response == y & tab$predictor == x &
    tab$effect_type == effect_type
  round(tab[mask, ]$effect, d)
}

rpt <- get_eff(y = "pop_stab", x = "w_avg_tlvl")
rpc <- get_eff(y = "pop_stab", x = "c")
rpe <- get_eff(y = "pop_stab", x = "env_stoch")
rpd <- get_eff(y = "pop_stab", x = "resp_div")

rsr <- get_eff(y = "sae_doak", x = "richness")
rst <- get_eff(y = "sae_doak", x = "w_avg_tlvl")
rsi <- get_eff(y = "sae_doak", x = "avg_int_strength")
rsc <- get_eff(y = "sae_doak", x = "ct_alive")

rcr <- get_eff(y = "cpe_int", x = "richness")
rci <- get_eff(y = "cpe_int", x = "avg_int_strength")
rct <- get_eff(y = "cpe_int", x = "w_avg_tlvl")
rcd <- get_eff(y = "cpe_int", x = "resp_div")
rcc <- get_eff(y = "cpe_int", x = "ct_alive")

# Detailled SEM coefficients
rsr <- get_eff(y = "sae_doak", x = "richness")
rst <- get_eff(y = "sae_doak", x = "w_avg_tlvl")
rsr <- get_eff(y = "sae_doak", x = "richness")
rsa <- get_eff(y = "sae_doak", x = "avg_int_strength")
rsct <- get_eff(y = "sae_doak", x = "ct_alive")
rsc <- get_eff(y = "sae_doak", x = "c")
rsd <- get_eff(y = "sae_doak", x = "resp_div")

```

```{r}
rscp <- get_eff(y = "stab_com", x = "pop_stab")
rsca <- get_eff(y = "stab_com", x = "async")

rac <- get_eff(y = "async", x = "cpe_int")
ras <- get_eff(y = "async", x = "sae_doak")
```

Applying a structural equation model on the outcomes of food web simulations
(see Methods) revealed that the temporal stability of community biomass is
positively influenced by both population stability and asynchrony, and that
population stability has twice the effect of asynchrony (resp. $r_\delta$=
`r rscp` and `r rsca`, $r_\delta$ being the standardised coefficient, Fig.
\@ref(fig:sem)a).

We further found that asynchrony is driven more by portfolio effects than by
compensatory dynamics (resp. $r_\delta$= `r ras` and `r rac`). In more details, response
diversity has strong positive effects on asynchrony, mainly through portfolio
effects ($r_\delta$= `r rsd`), but also through a weaker positive effect on compensatory
dynamics ($r_\delta$= `r rcd`). All other studied properties were found to have relatively
much weaker effects on asynchrony. Food web structural properties – namely
average trophic level, connectance and average interaction strength – have
consistent negative effects on asynchrony, through both portfolio and
compensatory effects (Fig. \@ref(fig:sem)a). Indeed, as average trophic level and connectance
increase compensatory effects are reduced (resp. $r_\delta$= `r rct` and `r rcc`, Fig.
\@ref(fig:sem)a), while average interaction strength has a small negative effect
on portfolio effects ($r_\delta$= `r rsi`). Species richness has a positive
effect on both portfolio and compensatory effects (resp. $r_\delta$=`r rsr` and
`r rcr`). Finally, as expected, environmental stochasticity has little effect on
asynchrony.

Focusing on population stability, environmental stochasticity was found to have
a strong negative effect on it ($r_\delta$= `r rpe`). All other studied properties were
found to have little effect on population stability, except for the average
trophic level which has a small positive effect ($r_\delta$= `r rpt`, Fig.
\@ref(fig:sem)a, but see Fig. S1) and response diversity a small negative effect
($r_\delta$= `r rpd`).

```{r}
rtsdiv <- get_eff(y = "stab_com", x = "resp_div", effect_type = "total")
rtsenv <- get_eff(y = "stab_com", x = "env_stoch", effect_type = "total")

rtsr <- get_eff(y = "stab_com", x = "richness", effect_type = "total")
rtst <- get_eff(y = "stab_com", x = "w_avg_tlvl", effect_type = "total")
rtsz <- get_eff(y = "stab_com", x = "Z", effect_type = "total")
rtsc <- get_eff(y = "stab_com", x = "c", effect_type = "total")
rtsct <- get_eff(y = "stab_com", x = "ct_alive", effect_type = "total")
rtsi <- get_eff(y = "stab_com", x = "avg_int_strength", effect_type = "total")

rtpr <- get_eff(y = "pop_stab", x = "richness", effect_type = "total")
rtpt <- get_eff(y = "pop_stab", x = "w_avg_tlvl", effect_type = "total")
rtpi <- get_eff(y = "pop_stab", x = "avg_int_strength", effect_type = "total")
rtpz <- get_eff(y = "pop_stab", x = "Z", effect_type = "total")
rtpc <- get_eff(y = "pop_stab", x = "c", effect_type = "total")
rtpct <- get_eff(y = "pop_stab", x = "ct_alive", effect_type = "total")

rtar <- get_eff(y = "async", x = "richness", effect_type = "total")
rtat <- get_eff(y = "async", x = "w_avg_tlvl", effect_type = "total")
rtai <- get_eff(y = "async", x = "avg_int_strength", effect_type = "total")
rtaz <- get_eff(y = "async", x = "Z", effect_type = "total")
rtac <- get_eff(y = "async", x = "c", effect_type = "total")
rtact <- get_eff(y = "async", x = "ct_alive", effect_type = "total")
```

Deriving total effects from the sum of direct and indirect effects in the
structural equation model (See Method), we showed that both environmental
stochasticity and species response diversity have by far the strongest total
effects on community stability (resp. $r_\delta$= `r rtsenv` and `r rtsdiv`,
Fig. \@ref(fig:sem)b). Species richness and average trophic level have similar
small but positive total effects on community stability ($r_\delta$=`r rtsr`),
but through opposite pathways. While species richness increases community
stability through a strong positive effect on asynchrony ($r_\delta$=`r rtar`),
it also weakly decreases population stability ($r_\delta$=`r rtpr`). In
contrast, average trophic level increases community stability through strong
positive effect on population stability ($r_\delta$=`r rtpt`), but is dampened
by a negative total effect on asynchrony ($r_\delta$=`r rtat`). Interestingly,
average interaction strength also has a positive effect on population stability
and a negative one on asynchrony (resp. $r_\delta$= `r rtpi` and `r rtai`). This
results in a total negative effect on community stability ($r_\delta$=
`r rtsi`). Connectance, instead, has a negative effect on community stability
($r_\delta$= `r rtsct`) through negative effects on both asynchrony and
population stability (resp. $r_\delta$=`r rtact` and `r rtpct`).

```{r}
tar_load(default_variable_values_pred)
pv <- default_variable_values_pred

pv_cap <- paste0(
  "connectance = ", round(pv["ct_alive"], 2),
  ", average interaction strength = ", round(pv["avg_int_strength"], 3),
  ", average trophic level = ", round(pv["w_avg_tlvl"], 2),
  ", predator interference = ", round(pv["c"], 1),
  ", Predator-Prey Mass Ratio = ", round(pv["Z"]),
  ", and environmental stochasticity = ", round(pv["env_stoch"], 1)

)
pred_cap <- paste0("Slope of the community stability - species
  richness relationship based on food-web structure and response diversity. a,
  Diagrams displaying the slope coefficients of the effect of species richness
  on community stability. The average values of the control variables were used
  to generate the predictions: ", pv_cap, ". b, c: Examples of predictions
  from the model for two combinations of average trophic level and response
  diversity (values displayed in panel a). Lines display the mean predictions of
  the model. The coefficients of the linear model are displayed in Fig. S3.
  Relationships between population stability, asynchrony and species richness
  are displayed in Fig. S2.
  "
  ) %>%
  str_replace_all("\\s+", " ")
```

```{r pred, fig.cap = pred_cap, fig.id = "pred"}
knitr::include_graphics(here("report", "figures", "p_fig3.pdf"))
```

We further assessed how environmental stochasticity and response diversity
interacted with food-web structural properties to determine the sign of the
stability-richness relationship, using a linear model including main effects and
all interactions among food-web structural properties, response diversity and
environmental stochasticity (see Methods). We found that there are strong
interactions between response diversity and food-web structure, which jointly
determine the sign of the stability-richness relationship.

In the absence of response diversity, we found only negative stability-richness
relationships, regardless of the food-web structure (Fig. \@ref(fig:pred)a).
Response diversity enables the rise of positive stability-richness
relationships, enhanced by average interaction strength and dampened by average
trophic level and connectance. Higher interaction strength leads to positive
stability-richness relationships for lower values of response diversity, and
interaction strength enhances asynchrony-richness relationships (Fig. S2), so
that the stronger positive stability-richness relationships are found at both
high interaction strength and high response diversity levels. In contrast,
higher average trophic level and connectance both lead to negative
stability-richness relationships, because they drive more negative population
stability-richness relationships and weaker positive asynchrony-richness
relationships (Fig. S2). Overall, our results highlight a strong contrast
between the relatively small effects of food-web structural properties on
overall community stability (Fig. \@ref(fig:sem)b) but their strong effects on
the slope of the stability-richness relationship (Fig. \@ref(fig:pred)a).

```{r, eval = FALSE}
tar_read(p_int_fw_str)
```

# Discussion

With the projected changes in climate and land use, and the expected resulting
changes in community composition, it is important to understand the mechanisms
which allow ecological communities to remain stable despite perturbations. The
two components of temporal community stability are the average stability of the
populations composing the community on the one hand, and the asynchrony of the
population fluctuations on the other hand[@thibaut_understanding_2013;
@zhao_biodiversity_2022]. In simplified assemblages, such as single trophic
level communities, community stability has been shown to be driven by asynchrony
in species fluctuations, rather than by population
stability[@olivier_urbanization_2020]. However, whether this holds for more
complex ecological communities, such as food webs, remains largely unknown.
Here, we tackled this question using simulations of complex food-web dynamics
with a stochastic bioenergetic model. Our model results show that, in complex
food-webs, community stability is more driven by population stability than by
asynchrony, in contrast with what was observed in single trophic level
communities but in agreement with a previous food-web data
study[@danet_species_2021]. Our results thereby contribute to reconcile models
and data analyses. The study further reveals that the main drivers of community
stability are environmental stochasticity and response diversity, respectively
acting on population stability and asynchrony. Surprisingly, food-web structural
properties have in turn much weaker effects, a result consistent across
methodological choices (Fig. S1). Structural properties were however found to
play an important role in the sign of the stability-richness relationship, which
is strongly mediated by the interaction between food-web structure and response
diversity. These results highlight the importance of response diversity to
improve our understanding of the stability of ecological communities in variable
environments.

We found environmental stochasticity and response diversity to be the main
overall drivers of community stability. While environmental stochasticity
decreases population stability, response diversity generates asynchrony, thereby
buffering environmental stochasticity. This result is in accordance with
previous theoretical and empirical findings in competitive
communities[@thibaut_diversity_2012; @thibaut_understanding_2013;
@thebault_trophic_2005; @loreau_species_2008; @de_mazancourt_predicting_2013;
@ives_stability_2000]. In turn, response diversity was found to operate on
asynchrony mainly through portfolio effects in agreement with empirical findings
in competitive assemblages[@thibaut_diversity_2012; @zhao_biodiversity_2022].
This suggests that response diversity generates asynchrony mainly by the
statistical averaging of random independent species fluctuations[@doak_statistical_1998].

A consequence of the large effect of response diversity is that asynchrony in
species fluctuations is driven more by the statistical averaging of independent
species fluctuations (portfolio effects) than by compensatory dynamics emerging
from species interactions. This finding is also coherent with previous studies
in competitive communities[@thibaut_diversity_2012; @zhao_biodiversity_2022],
but it was less expected to happen in food-webs. Trophic interactions are indeed
expected to result in more compensatory dynamics resulting from oscillations
between predators and prey[@vandermeer_coupled_2004] or prey switching from
predators[@mccann_weak_1998; @mccann_diversitystability_2000]. However,
pioneering theoretical studies have shown that environmental stochasticity can
dampen prey switching[@vasseur_environmental_2007] and that trophic cascades
between trophic levels in species rich food-webs are weak compared to
food-chains and species poor food-web modules[@fahimipour_compensation_2017].
So, the question of the importance of compensatory dynamics in complex food webs
remains an open and interesting avenue for future research.

In contrast to empirical evidence in competitive
communities[@olivier_urbanization_2020], population stability (not asynchrony)
was found to be  the dominant driver of community stability in our food web
simulations. These results are in agreement with previous findings in empirical
food-webs[@danet_species_2021] and food-web
models[@eschenbrenner_diversity_2023], which suggests fundamental differences
between competitive communities and food webs in the relative contributions of
population stability and asynchrony to community stability. Our findings suggest
that the presence of trophic links contributes to synchronise the species of the
community because we found that the food-web metrics studied – namely average
trophic level, connectance, and average interaction strength – all decrease
asynchrony. These results are in line with empirical evidence that predators can
synchronise their prey[@raimondo_interspecific_2004;
@korpimaki_predatorinduced_2005], and that species pairs involved in trophic
interactions are more synchronous than those that are not[@danet_species_2021].
A complementary explanation for the relatively lower importance of asynchrony in
food webs (compared to single trophic level communities) could be related to the
biomass structure of food-webs. Food-webs typically include a larger range of
body masses than single trophic level assemblages[@brose_predator_2019]. Such
highly uneven biomass distribution can translate in uneven species temporal
variability, thereby creating uneven species contributions to
asynchrony[@thibaut_understanding_2013] and limiting portfolio effects and
compensatory dynamics in species-rich communities[@zhao_biodiversity_2022].

Investigating in more detail the drivers of community stability, our results
suggest that species richness and the network structural properties investigated
only have a weak effect on overall community stability. We found that higher
average trophic level leads to higher community stability by increasing average
population stability, confirming previous theoretical[@shanafelt_stability_2018;
@eschenbrenner_diversity_2023; @brose_allometric_2006] and
empirical[@danet_species_2021] results. However, this effect disappeared when we
simulated community dynamics with equal mortality rates for all species instead
of allometric ones (Fig. S1). This suggests that the higher stability of
food-webs with higher average trophic levels stems more from their lower
mortality rates rather than from their trophic role directly. The overall effect
of connectance on overall community stability was found to be quite low, in
agreement with previous empirical and theoretical findings[@danet_species_2021;
@eschenbrenner_diversity_2023]. Similarly, we found very low effects of the
average interaction strength on community stability, which is in apparent
disagreement with the predominant role of interaction strength and connectance
on stability[@may_will_1972].

Our results also allow us to revisit the old question of the relationship
between the complexity of ecological communities and their
stability[@may_will_1972; @angelis_stability_1975]. We found that response
diversity enables the rise of positive stability-richness relationships,
enhanced by average interaction strength and dampened by average trophic level
and connectance. In the absence of response diversity, only negative
relationships between community stability and species richness are observed.
This reinforces the idea that increases in species richness increase community
stability only if it adds species that respond differently to environmental
perturbations, i.e. if there is response
diversity[@de_mazancourt_predicting_2013; @thebault_trophic_2005]. Indeed, a
community with numerous species which respond differently to environmental
change is more likely to buffer a wide array of perturbations
[@yachi_biodiversity_1999; @ives_stability_2000]. Despite its relatively weaker
effect on community stability (than response diversity and environmental
stochasticity), food-web structure interacts strongly with response diversity to
determine the sign of stability-richness relationships. Surprisingly, we find
that higher average interaction strength enhances positive stability-richness
relationships by enhancing asynchrony-richness relationships. Conversely, higher
average trophic level and higher connectance both led to negative
stability-richness relationships by dampening asynchrony-richness relationships.
Our results resonate with findings that some community structure can lead to
negative stability-richness relationships, as previously reported in freshwater
food-webs[@danet_species_2021]. Response diversity might therefore be one of the
mechanisms that explains why positive relationships between stability and
species richness are so prevalent in empirical settings[@elmqvist_response_2003;
@thibaut_understanding_2013], while food-web theoretical models that do not
often include environmental stochasticity and response diversity typically find
negative complexity-stability relationships[@ives_stability_2007].

In conclusion, our study highlights the complex but consistent effects of
response diversity and food-web structure on the temporal stability of species
rich communities. At the core of the insurance
hypothesis[@yachi_biodiversity_1999; @loreau_biodiversity_2021], our results
suggest that environmental stochasticity and response diversity are the main
mechanisms driving community stability, while food-web structural properties are
only mediating them. This aligns with the idea that, in species-rich
communities, community structure might be secondary in understanding the generic
effects of biodiversity on ecosystem functioning[@barbier_generic_2018;
@ives_stability_2000]. Our results further add to mounting evidence that
population stability is a stronger driver of community stability than asynchrony
in food-webs, in contrast with what was observed in competitive communities. The
fact that environmental stochasticity was found to be the main driver of overall
community stability raises concerns about the consequences of the current rise
in the frequency and severity of environmental extreme events for ecological
communities[@ghosh_temperature_2024; @lee_ipcc_2023]; indeed our results suggest
that these changes are unlikely to be compensated by biotic mechanisms alone.
Documenting how rapid biodiversity changes[@dornelas_assemblage_2014;
@blowes_geography_2019] driven by human pressures[@danet_past_2024] – and in
particular the reported homogenisation[@olden_homogocene_2018;
@wang_biotic_2021] in community composition – affect response diversity will be
key to accurately predict how it will affect, in turn, the stability of
communities.

# Methods

## Bioenergenetic model

We simulated the dynamics of complex food-webs using the allometric bioenergetic
model [@yodzis_body_1992; @brose_allometric_2006; @delmas_simulations_2017]. This model
is widely used to simulate the dynamics of complex ecological communities
because of the simplicity  and because it can be parametrized using  the
metabolic theory of ecology [@brown_toward_2004]. The model describes species
biomass dynamics over time with two master equations, one for the primary
producers (eq. \@ref(eq:prod)) and one for the consumers (eq.\@ref(eq:cons)):

\begin{equation}
    (\#eq:cons)
        \frac{dB_i}{dt} =
        \sum_{j \in \{\mathrm{res}\}_i} x_i y_i B_k F_{ij}
        - \sum_{j \in \{\mathrm{cons}\}_i} \frac{x_j y_j B_j F_{ji}}{e_{ij}}
        - x_i B_i - d_i B_i
\end{equation}

\begin{equation}
    (\#eq:prod)
	\frac{dB_i}{dt} =
    r_i B_i G_i
    - \sum_{j \in \{\mathrm{cons}\}_i} \frac{x_j y_j B_j F_{ji}}{e_{ij}}
    - d_i B_i
\end{equation}

The consumers gain biomass by consuming resources (i.e. their preys) at a rate
that depends on their mass specific metabolic rate ($x_i$),  maximum consumption
rate ($y_i$) and on the functional response ($F_ij$) describing how the rate of
consumption of a consumer $i$ on a resource $j$ vary with the biomass of this resource.
Consumers then lose biomass by being consumed, $e_ij$ being the assimilation
efficiency of the consumer $j$ on the resource $i$. Consumers also lose biomass
over time through metabolic losses ($-x_iB_i$) and by natural mortality due to
environment ($-d_iB_i$). The primary producers gain biomass over time with a
growth rate ($r_i$, eq. \@ref(eq:prod)) and a functional response ($G_i$)
describing how the growth of the producers vary with their biomass. They lose
biomass by being consumed by consumers and by natural mortality.

\begin{equation}
    (\#eq:growth)
    G_i = (1 - \dfrac{\sum_{j \in \{\mathrm{prod}\}} \alpha_{ij} B_j}{K_i})
\end{equation}

\begin{equation}
    (\#eq:holling)
    F_{ij} = \frac{\omega_{ij} B_j^h}
    {B_0^h + c_i B_i
    	+ \sum_{k \in \{\mathrm{res}\}_i} \omega_{ik} B_k^h}
\end{equation}

The growth of the primary producers (eq. \@ref(eq:growth)) is logistic
where the growth rate is maximal and null when
$\sum_{j \in \{\mathrm{prod}\}} \alpha_{ij} B_j$ are respectively close to 0 and 1. $\alpha_{ij}$ is the
per capita effect of the producer $j$ on the producer $i$, $K_i$ being the
carrying capacity for the producer $j$. The functional response of a consumer
feeding on a resource (eq. \@ref(eq:holling)) depends on the relative preference
of the consumer on the resource ($\omega_ij$) is limited by a half-saturation
rate ($B_0$), intraspecific interference coefficient ($c_i$), and on the
availability of its other resources
($\sum_{k \in \{\mathrm{res}\}_i} \omega_{ik} B_k^h$). Finally, the $h$ exponent
controls the shape of the functional response from a saturating function
($h = 1$, Holling type II) to a sigmoid ($h = 2$, Holling type III).

## Environmental stochasticity and response diversity

We added a stochastic natural mortality rate to the species dynamics,
representing a fluctuating environment. Based on the literature and previous
models [@vasseur_environmental_2007], we assumed that natural mortality rates
scales inversely with the species body mass ($d_i = d_0M_i^{-1/4}$), as all
other physiological parameters. We used a basal natural mortality rate of $d_0 =
.4$, as done in previous studies[@vasseur_environmental_2007].

We introduced environmental stochasticity by considering stochastic variation of
mortality rates. Doing so allows to consider stochasticity without modifying
species interactions themselves
[@vasseur_environmental_2007]. The stochastic part of the mortality rates
($\epsilon_{i, t}$) follow a normal distribution 0 centered
($\epsilon_{i, t} \sim N(0,\sigma^2_e)$) with a variance ($\sigma^2_e$). The
mortality rates of each species at the time $t$ equals to $d_{i,t} = d_i
e^{\epsilon_{i, t}}$, which ensured that the mortality rates never become
negative [@vasseur_environmental_2007]. We
simulated $\epsilon$ using an Ornstein-Uhlenbeck process, which is a modified
Brownian motion where the stochastic values tend to come back to the central
value (i.e. $\epsilon = 0$) such as $\epsilon_t$ varies according to this
stochastic differential equation: $d\epsilon_t = (0 - \epsilon_t)d_t + \sigma_e
dW_t$, where $dW_t$ is a Brownian motion. The strength of environmental
stochasticity was then controlled by $\sigma_e$.

Response diversity was controlled by the correlation among species stochastic
mortality rates ($\rho_{ij, i \neq j}$) such as response diversity is maximal
when species mortality rates are uncorrelated ($\rho = 0$) and null when species
mortality rates are perfectly correlated ($\rho = 1$) [@ripa_food_2003;
@vasseur_environmental_2007; @gouhier_synchrony_2010]. So the species mortality
rates were a product of the stochastic mortality rates generated by the
Ornstein-Uhlenbeck process and a variance-covariance matrix such as:

\begin{equation}
\begin{aligned}
\epsilon_{i, t} =
  \begin{bmatrix}
     \rho_{11}\sigma^2_e & \rho_{12}\sigma^2_e & \cdots & \rho_{1n}\sigma^2_e \\
     \rho_{21}\sigma^2_e & \rho_{22}\sigma^2_e & \cdots & \rho_{2n}\sigma^2_e \\
     \vdots  & \vdots  & \ddots & \vdots  \\
     \rho_{n1}\sigma^2_e & \rho_{n2}\sigma^2_e & \cdots & \rho_{nn}\sigma^2_e
  \end{bmatrix}
  \begin{bmatrix}
     a_{1,t} \\
     a_{2,t} \\
     \vdots  \\
     a_{n,t}
  \end{bmatrix}
(\#eq:rho)
\end{aligned}
\end{equation}

Where $a_{i,t}$ the stochastic value generated by the Ornstein-Uhlenbeck process
at time $t$, $\rho_{ii} = 1.0$ while all $\rho_{ij, i \neq j}$ were equal and
control the level of response diversity.

## Simulation

```{r}
tar_load(sim_param_tab)
spt <- sim_param_tab
get_sim_par_val <- function(x = "S", data = sim_param_tab) {
  x <- data[data$Parameter == x,][c("min", "max", "n")]
  setNames(unlist(x), c("min", "max", "n"))

}
pS <- get_sim_par_val(x = "S")
pct <- round(get_sim_par_val(x = "ct"), 2)
pZ <- get_sim_par_val(x = "Z")
pc <- get_sim_par_val(x = "c")
pe <- get_sim_par_val(x = "env_stoch")
prd <- get_sim_par_val(x = "resp_div")
pr <- get_sim_par_val(x = "rho")
```

We generated food-web structure using the niche model[@williams_simple_2000]
with an initial species richness from $`r pS["min"]`$ to $`r pS["max"]`$, and
connectance from $`r pct["min"]`$ to $`r pct["max"]`$ (Table S1). We generated
food-webs without cannibalistic links and discarded food-webs that contained
disconnected species.

We varied the strength of environmental stochasticity by varying $\sigma_e$ from
$`r pe["min"]`$ to $`r pe["max"]`$, the range of mortality rates observed in
protists[@vasseur_environmental_2007]. We decreased response diversity by
increasing $\rho$ from $`r pr["min"]`$ to $`r pr["max"]`$ (i.e. response
diversity being $1 - \rho$) (Fig. S4).

To vary interaction strength and mortality rates (i.e. allometrically
scaled mortality rate $d$) in the community, we varied the Predator-Prey
Mass Ratio (PPMR) from $`r pZ["min"]`$ to $`r pZ["max"]`$ (Fig. S5b).
We set $h = 2$, i.e. a functional response of type III (eq. \@ref(eq:holling)),
and varied predator interference (i.e. from $c = `r pc["min"]`$ to
$c = `r pc["max"]`$, eq. \@ref(eq:holling)), as predator interference was shown
to have a tremendous role on population stability[@brose_allometric_2006]. We
set a global carrying capacity of $K' = 10$ to ensure that consumers were not
limited by the biomass input from primary producers. We standardised the
carrying capacity by the number of primary producers and by the interspecific
competition among producers to ensure that the effects of species richness on
temporal stability are not driven by the increase in the number of primary
producers, thereby trivially increase biomass input for consumers. We then
standardised $K'$ such as $K_i = \frac{1 + \alpha_{ij} (S - 1)}{S}K'$
[@ives_stability_2000]. However, we found that our results were not affected
when removing the standardisation of carrying capacity (Fig. S1).

We numerically solved the stochastic differential equation systems during 2000
timesteps and collected the last 500 timesteps, similarly to previous studies
[@brose_allometric_2006]. We set $d_t = .1$, and $d_t = .05$ when instability
was detected during the simulations due to stiff changes in biomass. We set
extinction threshold to $10^{-6}$ and if an extinction happened in the last 500
timesteps, we continued to run the simulation for another 1000 timesteps until
there was no extinction events during the last 500 timesteps. When we found
disconnected species at the end of the simulation (i.e. a primary producer
without consumer or a consumer without prey), we set their biomass to 0 and
re-run the simulation for another 1000 timesteps. Except in the case of
disconnected species, we re-ran simulations with species starting biomass equal
to their biomass at the last timestep of the previous simulation. Supplementary
analysis showed that alternative choices of simulation processes, such as longer
simulations, rebuilding food-webs (i.e. resetting consumer preferences,
recomputing species bodymass according to the new food-web) after removing
disconnected species did not affect our results (Fig. S1). The food-web
generation and simulations were done using the `Julia` package
`EcologicalNetworkDynamics.jl`[@lajaaiti_ecologicalnetworksdynamicsjl_2024].

In sum, we generated `r format(m_sim(col = "n"), scientific = FALSE)`
simulations generated a wide range of food-web structure (Median (5%, 95%),
species richness: `r m_sim("richness")`, connectance: `r m_sim("ct_alive")`,
average interaction strength: `r m_sim("avg_int_strength")`, maximum trophic
level: `r m_sim("max_tlvl")`) (Table S3).

## Simulation analysis

We measured food-web properties that have been linked with stability in the
literature.  We measured total biomass, species richness, connectance, average
trophic level weighted by biomass, average omnivory and average interaction
strength. Food-web properties were measured at each of the 500 timesteps and
then averaged, except connectance and trophic level that were static.
Connectance was computed as $C = L / S(S-1)$, $L$ being the number of trophic
links and $S$ being the number of species in the community. Species trophic
level was computed recursively from the bottom of the food-web, such as the
trophic level of a consumer was equal to the average trophic level of its
resources added to 1, the trophic level of primary producers being set to 1
[@christensen_ecopath_1992]. Average trophic level was then equal to:
$1/\hat{B}_i\sum_{i}\mathrm{TL}_i$, $\mathrm{TL}_i$ being the trophic level and
$\hat{B}_i$ the average biomass of the species $i$. The degree of omnivory of a
consumer was computed  such as the sum of squares of its resource trophic levels
weighted by the relative preference of the consumer for each resource: $\sum_i
(\mathrm{TL}_i - \hat{\mathrm{TL}})^2 \omega_i/ \sum \omega_i$
[@christensen_ecopath_1992]. Interaction strength was computed was quantified as
the biomass fluxes going from a resource to a consumer, such as: $I = x_i y_i
B_k F_{ij}$ (eq. \@ref(eq:cons)), and averaged over time. Our way of measuring
interaction strength is very similar to previous metrics used in previous
studies, at the difference here that we average fluxes over time and not the
maximal interaction strength [@mccann_weak_1998]. Finally, the average
interaction strength of a community was computed as the average interaction
strength at the exclusion of null interaction strengths (i.e. excluding absence
of trophic interactions).

We assessed the effects of food-web structure, environmental stochasticity and
response diversity on temporal stability of biomass. We measured temporal stability of biomass
as the inverse of the coefficient of variation. We further partitioned stability
into population stability ($S_{pop}$) and asynchrony ($\phi$) such as:

\begin{equation}
\begin{aligned}
S_{com} = S_{pop} \times \phi
(\#eq:scom)
\end{aligned}
\end{equation}

with $S_{com} = \mu_{tot} / \sigma_{tot}$, $S_{pop} = \mu_{tot} / \sum_i \sigma_i$,
$\phi = \sum_i \sigma_i / \sigma_{tot}$.

$\mu_{tot}$ and $\sigma_{tot}$ being respectively average total biomass and standard
deviation of total biomass.

To further get a mechanistic insight about the effects of stressors and food-web
structure on stability, we partitioned asynchrony into portfolio effects and
compensatory dynamics stemming from species interaction[@zhao_biodiversity_2022]
($CPE_{int}$). The measurement of portfolio effects has been a topic of
debate[@loreau_biodiversity_2021;@zhao_biodiversity_2022]. Doak's
definition[@doak_statistical_1998] defines portfolio effects as the product of
statistical averaging effects ($SAE$) and compensatory effects arising from
response diversity[@zhao_biodiversity_2022; @doak_statistical_1998]
($CPE_{env}$). It then quantities the portfolio effect ($PFE = SAE \times
CPE_{env}$) emerging from independent species fluctuations, i.e. the statistical
averaging effect ($SAE$), weighted by the effects of response diversity of
species to environmental fluctuations, such as portfolio effects are dampened if
species have correlated responses.

\begin{equation}
\begin{aligned}
S_{com} = S_{pop} \times SAE \times CPE
(\#eq:scom2)
\end{aligned}
\end{equation}

Statistical averaging effect is the share of asynchrony assuming that species are
independent, meaning that the variance total of the community is equal to the
sum of species variances only (i.e. species covariances are null). We can
define community stability ($S_{com, IP}$) in that scenario and then derive $SAE$.

\begin{equation}
\begin{aligned}
S_{com, IP} &= \dfrac{\mu_{tot}}{\sqrt{\sum_i \sigma^2_i}} = SAE \times S_{pop} \\
SAE &= \dfrac{S_{com, IP}}{S_{pop}} = \dfrac{\sum_i \sigma_i}{\sqrt{\sum_i \sigma^2_i}}\\
\end{aligned}
(\#eq:sae2)
\end{equation}

Compensatory dynamics in the food-webs is then the remaining part of asynchrony:

\begin{equation}
\begin{aligned}
CPE &= \dfrac{S_{com}}{SAE \times S_{pop}} = \dfrac{\sqrt{\sum_i \sigma^2_i}}{\sigma_{tot}}\\
\end{aligned}
(\#eq:cpe)
\end{equation}

Compensatory dynamics can arise from predator-prey interactions
($CPE_{int}$) or from differential response of species to environmental
stochasticity ($CPE_{env}$). From the deterministic and stochastic parts of
mortality rates, we computed the compensatory dynamics linked to the environment
such as:

\begin{equation}
\begin{aligned}
CPE_{env} &= \mathbf{B} \times \mathrm{diag}(d_i) \times e^\mathbf{E}
\end{aligned}
(\#eq:cpenv)
\end{equation}

With $\mathbf{B}$ and $\mathbf{E}$ being the matrices time x species for species biomass
and stochastic mortality rates respectively. Then compensatory dynamics raising
from species interactions were computed as: $CPE_{int} = CPE / CPE_{env}$.
Finally, portfolio effects was computed as: $PFE = SAE \times CPE_{env}$ and
compensatory dynamics raising from species interactions as $CPE_{int}$. The
final stability decomposition reads as follows:

\begin{equation}
\begin{aligned}
S_{com} = S_{pop} \times PFE \times CPE_{int}
(\#eq:scom3)
\end{aligned}
\end{equation}

## Statistical analysis

Using a structural equation model, we disentangled the complex effects of
food-web structure, environmental stochasticity and response diversity on the
different facets of temporal stability of biomass (i.e. population stability,
asynchrony, portfolio effects and compensatory dynamic effects). We tested the
effects of food-web structure (average interaction strength, average trophic
level, connectance) on population stability and asynchrony partitions (portfolio
and compensatory effects). Grounded in theoretical ecology, we expected that
higher average trophic level, lower interaction strength, lower
connectance result in higher population stability and lower asynchrony[@mccann_weak_1998; @mccann_reevaluating_1997;
@mccann_diversitystability_2000; @fahimipour_compensation_2017]. We further
added simulation parameters as control
variables because higher Predator-Prey Mass Ratio and predator interference were
also expected to increase population stability [@brose_allometric_2006].
Finally, we tested the effects of environmental stochasticity and response
diversity on population stability and asynchrony partitions, although there were
predicted to mainly affect to population stability (i.e. by increasing amplitude
of population fluctuations) and asynchrony (i.e. by generating independent
population fluctuations) respectively. We completed the structural equation
model with the mathematical relationship linking compensatory dynamics due to
species interactions ($CPE_{int}$) and statistical averaging effects ($SAE
\times CPE_{env}$) to asynchrony, and from asynchrony and population stability
to community stability.

Prior to test the structural equation model, we logged all the stability
components to transform their relation from multiplicative to additive (eq.
\@ref(eq:scom2)). We ensured that all the linear models composing the structural
equation model presented low multicollinearity (Variance Inflation Factor < 3,
Table S4), although interaction strength and species richness were highly
correlated (Fig. S5). The structural equation model using the R package
`PiecewiseSEM`[@lefcheck_piecewisesem_2016]. We then bootstrapped (N = 100) the
SEM to obtain confidence intervals on the estimated effects and computed the sum
of the direct and indirect effects using `semEff` R
package[@murphy_semeff_2021]. In the main text, we reported standardised
coefficients which were obtained by scaling the coefficients by the standard
deviation of the response and predictor variable. Finally, we reported only the
direct standardised coefficients that had an absolute value above $0.05$,
because almost all linked were statistically significant given the high numbers
of simulations.

In a second analysis, we tested how food-web structure, response diversity, and
environmental stochasticity modulate stability-species richness relationships.
We modelled temporal stability of community biomass according to food-web
structure: species richness, weighted average trophic level, connectance and
average interaction strength, as well as environmental stochasticity and
response diversity. We further included the two-way interactions between species
richness and food-web structure, between species richness and response
diversity, between species richness and environmental stochasticity. Finally, we
added the three-way interactions between species richness, food-web structure
and response diversity, between species richness, food-web structure and
environmental stochasticity. We added predator interference, predator-prey mass
ratio as control variables. We further added the ID of the food-web as a random
intercept. We then used the linear equation of the model and associated slope
coefficients to predict the shape of the relationship between community
stability and species richness according to a gradient of response diversity and
of food-web structure. To do so, we summed all the coefficients involving
species richness in the model, i.e. the one, two, and three-way terms. The
values of the other variables were set to the values indicated in the Fig.
\@ref(fig:pred). The VIF were inferior to 3, indicating low multicollinearity
(Table S5), we checked the distribution of the residuals (Fig. S6) using
`DHARMa` R package. The model was implemented using the R package
`glmmTMB`[@brooks_glmmtmb_2017].

# References {-}